### 2018-11-21 paper.tex
Paper outline

\documentclass{IEEEtran}

\usepackage{graphicx}
\usepackage{color}
\usepackage[bookmarks=false]{hyperref}

\newcommand{\todo}[1]{\marginpar{\parbox{18mm}{\flushleft\tiny\color{red}\textbf{TODO}:
      #1}}}

\begin{document}

\title{Performance Evaluation of Big Data Management Strategies\\ for Neuroinformatics}

\author{
  \IEEEauthorblockN{
    Val\'erie Hayot-Sasson and 
    Tristan Glatard
  }\\
  \IEEEauthorblockA{
    Department of Computer Science and Software Engineering, Concordia University, Montreal, Qu\'ebec, Canada
  }
}

\maketitle

%
% Abstract
%
\begin{abstract}

\end{abstract}

%
% Section
%
\section{Introduction} % 1 page with abstract

% Big Data Engines have been developed mostly for text applications, 
%  in particular Web. Mention Spark and Dask

% Scientific applications, including neuroinformatics are different: CLI, binary data, shared HPC clusters (although ref to cloud).
% In neuroinformatics, nipype, psom, etc
% This follows the current trend to big data science, data sets *and derived data* are larger (give examples).

% Our question: is it relevant to use Big Data engines for neuroinformatics? 
% Implications: migrate existing pipelines, adopt a new framework (costly)
% Some papers said yes, for instance Ariel's and the Spark provenance one (https://github.com/UFFeScience/SAMbA).
% Perhaps this too: https://www.sciencedirect.com/science/article/pii/S0925231218311470 Spark is being used to implement neuroinformatics pipelines, but is it
% the right tool?
% But a quantitative evaluation is lacking.

% More precisely, we intend to quantify the effect of in-memory computing, data locality, lazy evaluation
% on various use cases. We intentionally leave scheduling out here.

% The use of a Big Data engine is not obvious when tmpfs and asynchronous writes
% are available and could be used to emulate in-memory computing. Our goal is to quantify when and how Big Data engines will be useful.
% Say that we don't use HDFS because we focus on clusters and HDFS on multi-tenant clusters is difficult (refer to hadoop on demand).

% Talk about pilot jobs and say we don't consider them here.

% "```Disk and network I/O, of course, play a part in Spark performance 
% as well, but neither Spark nor YARN currently do anything to actively 
% manage them.```"
% http://blog.cloudera.com/blog/2015/03/how-to-tune-your-apache-spark-jobs-part-2/

% Here we focus on performance but we are aware of other desirable properties of workflow engines, including provenacne.

% Outline

% Talk about containers somewhere.

\section{Materials and Methods} % 4 pages

\subsection{Engines} % 1.5 pages

\subsubsection{Apache Spark}

%% Main concepts
% detailed description relevant to the experiment
% how shuffle works
% how data locality is implemented
% spills to disk when dataset (and derived data) is too large for memory
% Lazy evaluation
% overlay cluster: not considered here as this would be scheduling

%% Technical details
% data serialization to Java (we use Python as most neuroinformatics do)
% limitation in size of RDD element, due to java implementation of binary data
% memory management: don't give too much memory to executors or GC
%  many places recommend limiting to 5 cores/executor, no one explains why
% more on tuning?

\subsubsection{Nipype}

% general description relevant to the experiment
% two plugins
% provenance
% memory management (bad?)

\subsubsection{Execution modes} % perhaps find a better title, it's not informative

% in-memory (Spark only)

% tmpfs: two extra in-memory copies. memfs could address that but we don't have it yet.
% Spark: when just passing filenames. turns out i need to set `preservePartitioning` to True
% local disk: write buffer issue

% shared fs: Lustre

% Make a table to explain where data locality and in-memory computing are
% available.

% What happens when dataset and derived data is too large for disk?

\subsection{Simulation} % is it included? 0.5 page

\subsection{Models: Write Buffer} % perhaps other models if relevant. 0.5 page

% Perhaps refer to this: https://ieeexplore.ieee.org/abstract/document/5496998

% 15 cores per executor can lead to bad I/O throughput
% http://blog.cloudera.com/blog/2015/03/how-to-tune-your-apache-spark-jobs-part-2/

% measured disk bandwidth with random I/O, very different from sequential

\subsection{Experiments} % 1.5 pages

\subsubsection{Infrastructure} % 1/4 a page (half a col)

% Dedicated Compute nodes with SLURM
% Lustre server

\subsubsection{Datasets} % 1/4 of a page (half a col)

% BigBrain
% Some BIDS dataset for fmriprep

\subsubsection{Applications} % 1 page

% Incrementation (avoid cache effects in binarization)

% average (1 shuffle)
% Nipype: can't use SLURM plugin so had to actively poll.

% kmeans (many shuffles)

% Example BIDS app

% fmriprep?

% Explain how we tweaked scheduling in Nipype to make it similar to Spark, 
% otherwise we're only measuring scheduling differences. Needs a discussion on scheduling somewhere.

\section{Results} % 2 pages

% We checked that scheduling was similar in both cases (show a two Gantt charts to illustrate)

% interactive gantt charts?

\section{Discussion} % 2 pages

\subsection{General concepts}

% Shared fs vs local disk vs in-memory: what do we gain?

% Can we emulate in-memory computing using write buffers or tmpfs?

% Scheduling: load balance as much as you can.
% (which isn't what nipype multiproc does)

% tmpfs and local disks fill up: need cleanup.

% fault-tolerance: do we have anything to discuss here?

\subsection{Engine-specific observations}

\section{Conclusion} % 1 page with refs

% When is it useful to use a Big Data engine?

% Future work: scheduling in a shared environment.

\section{Acknowledgments}

\bibliographystyle{IEEEtran} 
\bibliography{IEEEfull,biblio}

\end{document}

### 2018-11-21 tests/test_inc.py
Fixed pipeline paths

@@ -8,7 +8,7 @@ from os import path as op
 
 def test_increment_spark():
 
-    p = subprocess.Popen(['python', 'spark_inc.py', 'sample_data', 'inc_out',
+    p = subprocess.Popen(['python', 'pipelines/spark_inc.py', 'sample_data', 'inc_out',
                           '1'])
     p.communicate()
 
@@ -20,7 +20,7 @@ def test_increment_spark():
 
     assert h_prog_1 == h_exp_1
 
-    p = subprocess.Popen(['python', 'spark_inc.py', 'sample_data', 'inc_out',
+    p = subprocess.Popen(['python', 'pipelines/spark_inc.py', 'sample_data', 'inc_out',
                           '10'])
     p.communicate()
 
@@ -38,7 +38,7 @@ def test_increment_spark_cli():
     shutil.rmtree('inc_out', ignore_errors=True)
     shutil.rmtree('inc_work', ignore_errors=True)
 
-    p = subprocess.Popen(['python', 'spark_inc.py', 'sample_data', 'inc_out',
+    p = subprocess.Popen(['python', 'pipelines/spark_inc.py', 'sample_data', 'inc_out',
                           '1', '--cli', '--work_dir', 'inc_work'])
     p.communicate()
 
@@ -50,7 +50,7 @@ def test_increment_spark_cli():
 
     assert h_prog_1 == h_exp_1
 
-    p = subprocess.Popen(['python', 'spark_inc.py', 'sample_data', 'inc_out',
+    p = subprocess.Popen(['python', 'pipelines/spark_inc.py', 'sample_data', 'inc_out',
                           '10', '--cli', '--work_dir', 'inc_work'])
     p.communicate()
 
@@ -67,7 +67,7 @@ def test_increment_spark_cli():
 def test_increment_nipype():
 
     shutil.rmtree('nipinc_out', ignore_errors=True)
-    p = subprocess.Popen(['python', 'nipype_inc.py', 'sample_data',
+    p = subprocess.Popen(['python', 'pipelines/nipype_inc.py', 'sample_data',
                           'nipinc_out', '1'])
     (out, err) = p.communicate()
 
@@ -80,7 +80,7 @@ def test_increment_nipype():
     assert h_prog_1 == h_exp_1
 
     shutil.rmtree('nipinc_out')
-    p = subprocess.Popen(['python', 'nipype_inc.py', 'sample_data',
+    p = subprocess.Popen(['python', 'pipelines/nipype_inc.py', 'sample_data',
[…]

### 2018-11-21 paper.tex
A few more comments from @ValHayot's course project report
Typo
Wrote the page cache model

@@
 % Implications: migrate existing pipelines, adopt a new framework (costly)
-% Some papers said yes, for instance Ariel's and the Spark provenance one (https://github.com/UFFeScience/SAMbA).
+% Some papers said yes, for instance Ariel's, Thunder and the Spark provenance one (https://github.com/UFFeScience/SAMbA).
 % Perhaps this too: https://www.sciencedirect.com/science/article/pii/S0925231218311470 Spark is being used to implement neuroinformatics pipelines, but is it
@@
 
+% define pipeline and workflows
+
 % Here we focus on performance but we are aware of other desirable properties of workflow engines, including provenacne.
@@
 %% Main concepts
+% pipeline description
 % detailed description relevant to the experiment
@@
 % Lazy evaluation
+% persistence (refer to Thunder)
+% multiple schedulers: we focus on Spark Standalone
 % overlay cluster: not considered here as this would be scheduling
@@
 
+% pipeline description: map nodes and their limitations
 % general description relevant to the experiment
@@
 % provenance
+% file system caching
 % memory management (bad?)
@@
 
-\subsection{Models: Write Buffer} % perhaps other models if relevant. 0.5 page
+\subsection{Page Cache Model} % perhaps other models if relevant. 0.5 page
 
@@
 % measured disk bandwidth with random I/O, very different from sequential
+
+The Linux kernel uses in-memory page caching to speed-up disk writes. 
+When a process calls the \texttt{write()} system call, data is copied 
+to a memory cache that is asynchronously written to disk by flusher 
+threads when memory shrinks, ``dirty" (unwritten) data grows, or a 
+process invokes the \texttt{sync()} system call~\cite{love2010linux}. 
+The flushing of page cache to disk is also called the writeback 
+process.
+
+Page caching has a dramatic effect on I/O, reducing the duration of 
+disk writes by several orders of magnitude. It is essentially a way to 
+emulate in-memory computing at the kernel level, without requiring a 
+dedicated engine. The size of the page cache, however, becomes a 
+limitation when processes write faster than the disk bandwidth. When 
+this happens, the page cache rapidly fills up and writes are limited by 
+the disk write bandwidth -- typically 500~MB/s for random writes to 
+contemporary SSDs  -- as if no page cache was involved.
+
+We introduce the following basic model to describe the filling and 
+flushing of the page cache by an application:
+$$
+d(t) = \left( \frac{D}{C} - \frac{\delta}{\gamma} \right)t + d_0,
+$$
+where:
+\begin{itemize}
+\item $d(t)$ is the amount of data in the page cache at time t
+\item $D$ is the total amount of data written by the application
+\item $C$ is the total CPU time of the application
+\item $\delta$ is the disk bandwidth
+\item $\gamma$ is the max number of concurrent processes
+\item $d_0$ is the amount of data in the page cache at time $t_0$
+\end{itemize}
+
+This model applies to parallel applications assuming that (1) 
+concurrent processes all write the same amount of data, (2) 
+concurrent processes all consume the same CPU time, (3) data is written 
+uniformly along task execution. Due these assumptions, all the 
+processes will write at the same rate, which explains why the model 
+does not depend on the total number of concurrent processes in the 
+application, but only on the max number of concurrent processes 
+executing on the same node ($\gamma$). We realize that these 
+assumptions would usually be violated in practice, but this simple 
+model already provides interesting insights on the performance of disk 
+writes, as shown later. Naturally, the model also ignores other 
+processes that might be writing to disk concurrently to the 
+application, which we assume negligible here.
+
+In general, an application should ensure that $\dot d$ remains negative 
+or null, leading to the following inequality:
+\begin{equation}
+\frac{D}{C} \leq \frac{\delta}{\gamma} \label{eq:page-cache-inequality}
+\end{equation}
+This defines a data-compute ratio beyond which the page cache becomes 
+asymptotically useless, which we will use to define our benchmarks.
+
+% Cache eviction: LRU/n.
 
@@
 
+% Do we see the effect of Java serialization?
+
 \section{Conclusion} % 1 page with refs
@@
 
-% Future work: scheduling in a shared environment.
+% Future work:
+% - scheduling in a shared environment.
+% - workflow-aware cache eviction strategies instead of LRU/n
 
@@
 \bibliographystyle{IEEEtran} 
-\bibliography{IEEEfull,biblio}
+\bibliography{biblio}
 

### 2018-11-21 tests/test_inc.py
Fixed tests

@@ -8,7 +8,8 @@ from os import path as op
 
 def test_increment_spark():
 
-    p = subprocess.Popen(['python', 'pipelines/spark_inc.py', 'sample_data', 'inc_out',
+    p = subprocess.Popen(['python', 'pipelines/spark_inc.py',
+                          'sample_data', 'inc_out',
                           '1'])
     p.communicate()
 
@@ -20,7 +21,8 @@ def test_increment_spark():
 
     assert h_prog_1 == h_exp_1
 
-    p = subprocess.Popen(['python', 'pipelines/spark_inc.py', 'sample_data', 'inc_out',
+    p = subprocess.Popen(['python', 'pipelines/spark_inc.py',
+                          'sample_data', 'inc_out',
                           '10'])
     p.communicate()
 
@@ -38,7 +40,8 @@ def test_increment_spark_cli():
     shutil.rmtree('inc_out', ignore_errors=True)
     shutil.rmtree('inc_work', ignore_errors=True)
 
-    p = subprocess.Popen(['python', 'pipelines/spark_inc.py', 'sample_data', 'inc_out',
+    p = subprocess.Popen(['python', 'pipelines/spark_inc.py',
+                          'sample_data', 'inc_out',
                           '1', '--cli', '--work_dir', 'inc_work'])
     p.communicate()
 
@@ -50,7 +53,8 @@ def test_increment_spark_cli():
 
     assert h_prog_1 == h_exp_1
 
-    p = subprocess.Popen(['python', 'pipelines/spark_inc.py', 'sample_data', 'inc_out',
+    p = subprocess.Popen(['python', 'pipelines/spark_inc.py',
+                          'sample_data', 'inc_out',
                           '10', '--cli', '--work_dir', 'inc_work'])
     p.communicate()
 
@@ -129,7 +133,8 @@ def test_increment_nipype_cli():
 def test_benchmark_spark():
 
     shutil.rmtree('inc_out', ignore_errors=True)
-    p = subprocess.Popen(['python', 'pipelines/spark_inc.py', 'sample_data', 'inc_out',
+    p = subprocess.Popen(['python', 'pipelines/spark_inc.py',
+                          'sample_data', 'inc_out',
                           '1', '--benchmark'], stdin=subprocess.PIPE,
                          stdout=subprocess.PIPE)
     (out, err) = p.communicate()

### 2018-11-22 paper.tex
Some restructuration

@@
 
-\title{Performance Evaluation of Big Data Management Strategies\\ for Neuroinformatics}
+\title{Performance Evaluation of Big Data Management Strategies for Neuroinformatics}
 
@@
 %  in particular Web. Mention Spark and Dask
+Big Data processing engines have been leveraging data locality, 
+in-memory computing and lazy evaluation to reduce the impact of data 
+manipulation on the performance of applications. From early Hadoop 
+implementations~\cite{hadoop} to Apache Spark~\cite{spark} and 
+Dask~\cite{dask}, these frameworks have gained popularity and are now 
+becoming mainstream in data analytics.
+
+In scientific data analysis, however, the adoption of Big Data engines
 
@@
 
-\subsubsection{Execution modes} % perhaps find a better title, it's not informative
+\subsection{Performance Models}
 
-% in-memory (Spark only)
+\subsubsection{In-Memory File System}
 
-% tmpfs: two extra in-memory copies. memfs could address that but we don't have it yet.
-% Spark: when just passing filenames. turns out i need to set `preservePartitioning` to True
-% local disk: write buffer issue
-
-% shared fs: Lustre
-
-% Make a table to explain where data locality and in-memory computing are
-% available.
-
-% What happens when dataset and derived data is too large for disk?
-
-\subsection{Simulation} % is it included? 0.5 page
-
-\subsection{Page Cache Model} % perhaps other models if relevant. 0.5 page
+\subsubsection{Local Disk (Page Caching)} % 0.5 page
 
@@
 
+\subsubsection{Shared File System}
+
+% is there a page cahe when writing to a socket? there must be!
+% anyway the model is in the form of Eq 1, but now delta is the network bandwidth
+% and gamma is the total number of concurrent processes.
+% also, is lustre going to slow down network writes?
+
+% Make a table to explain where data locality and in-memory computing are
+% available.  As a summary of the models.
+
+% What happens when dataset and derived data is too large for disk?
+
+\subsection{Simulation} % is it included? 0.5 page
+
+% Even if we end up not using simulation, here we should explain
+% why the model in simgrid is limited. Refer to Fred's paper at CCGrid 2017.
+% The following sub-sections become a set of models that we could implement
+% in simulation to make it more realistic.
+
+% Talk about ~\cite{lebre2015adding}. Most simulation toolkits have focused on task 
+%  scheduling. And https://dl.acm.org/citation.cfm?id=3041715
+
+% And Wrench.
+
 \subsection{Experiments} % 1.5 pages
@@
 
+\subsubsection{Execution modes} % perhaps find a better title, it's not informative
+
+% separate this in (1) the concepts, put in the model sections below, 
+% (2) the implementation 'details', put in the experiments
+
+% in-memory (Spark only)
+
+% tmpfs: two extra in-memory copies. memfs could address that but we don't have it yet.
+% Spark: when just passing filenames. turns out i need to set `preservePartitioning` to True
+% local disk: write buffer issue
+
+% shared fs: Lustre
+
 \section{Results} % 2 pages
@@
 
-% interactive gantt charts?
+\subsection{Model Evaluations}
+
+\subsection{CPU Time}
+% Comment on the general trend
+
+\subsection{Image Block size}
+% Comment on the general trend
+
+\subsection{Data Size}
+% Comment on the general trend
+
+\subsection{Iterations}
+% Comment on the general trend
+
+% Interactive gantt charts?
 
@@
 
-\subsection{General concepts}
-
+\subsection{Effect of In-Memory Computing}
 % Shared fs vs local disk vs in-memory: what do we gain?
 
+\subsection{Effect of Data Locality}
+
+% Effect of Lazy Evaluation?
+
+\subsection{Can \texttt{tmpfs} and Page Caches Emulate In-Memory Computing?}
+% tmpfs and local disks fill up: need cleanup.
 % Can we emulate in-memory computing using write buffers or tmpfs?
 
+\subsection{Scheduling Remarks}
 % Scheduling: load balance as much as you can.
@@
 
-% tmpfs and local disks fill up: need cleanup.
+\subsection{Other Comments}
 
@@
 
-\subsection{Engine-specific observations}
+\subsection{Engine-Specific Comparisons}
 
+% Spark vs Nipype
 % Do we see the effect of Java serialization?

### 2018-11-22 paper.tex
First cut at an intro

@@
 
-\title{Performance Evaluation of Big Data Management Strategies for Neuroinformatics}
+\title{Performance Evaluation of Big Data Processing Strategies for Neuroimaging}
 
@@
 
-%
-% Abstract
-%
 \begin{abstract}
@@
 
-%
-% Section
-%
+% Talk about containers somewhere?
+
 \section{Introduction} % 1 page with abstract
 
-% Big Data Engines have been developed mostly for text applications, 
-%  in particular Web. Mention Spark and Dask
-Big Data processing engines have been leveraging data locality, 
-in-memory computing and lazy evaluation to reduce the impact of data 
-manipulation on the performance of applications. From early Hadoop 
-implementations~\cite{hadoop} to Apache Spark~\cite{spark} and 
-Dask~\cite{dask}, these frameworks have gained popularity and are now 
-becoming mainstream in data analytics.
+Big Data processing engines leverage data locality, 
+in-memory computing, and lazy evaluation to reduce the impact of data 
+manipulation on the performance of applications. From early MapReduce 
+implementations~\cite{dean2008mapreduce} to Apache Spark~\cite{zaharia2016apache} and now 
+Dask~\cite{rocklin2015dask}, these frameworks have become 
+mainstream tools for data analytics. Meanwhile, several scientific 
+domains including bioinformatics, physics or astronomy, have entered 
+the Big Data era due to increasing data volumes and variety. 
+However, the adoption of Big Data engines for scientific data analysis 
+remains limited, perhaps due to the widespread availability of 
+scientific processing engines such as Pegasus~\cite{deelman2005pegasus}, 
+Taverna~\cite{oinn2004taverna}, and many others.
 
-In scientific data analysis, however, the adoption of Big Data engines
+Scientific applications slightly differ from the typical Big Data use 
+cases, which might explain the remaining gap between Big Data and 
+scientific engines. While Big Data applications mostly target text 
+processing (Web search, frequent pattern mining, recommender systems, 
+etc~\cite{leskovec2014mining}), implemented in consistent software libraries, and 
+running on clouds or dedicated commodity clusters with locality-aware 
+file systems such as HDFS~\cite{shvachko2010hadoop}, scientific applications often 
+involve binary data such as images and signals, processed by 
+command-line tools using a mix of programming languages (C, Fortran, 
+Python, shell scripts) and deployed on large, shared clusters where 
+data is fetched from data nodes to compute nodes using shared file 
+systems such as Lustre~\cite{schwan2003lustre}. Such differences in applications 
+and infrastructures have important consequences. To mention only one, 
+in-memory computing requires instrumentation to be applied to 
+command-line tools.
 
-% Scientific applications, including neuroinformatics are different: CLI, binary data, shared HPC clusters (although ref to cloud).
-% In neuroinformatics, nipype, psom, etc
-% This follows the current trend to big data science, data sets *and derived data* are larger (give examples).
+Technological advances of the past decade, in particular page caching 
+in the Linux kernel~\cite{love2010linux}, in-memory file systems 
+(\texttt{tmpfs}) and memory-mapped files\todo{check that} might also 
+explain the lack of adoption of Big Data engines for scientific 
+applications.
 
-% Our question: is it relevant to use Big Data engines for neuroinformatics? 
-% Implications: migrate existing pipelines, adopt a new framework (costly)
-% Some papers said yes, for instance Ariel's, Thunder and the Spark provenance one (https://github.com/UFFeScience/SAMbA).
-% Perhaps this too: https://www.sciencedirect.com/science/article/pii/S0925231218311470 Spark is being used to implement neuroinformatics pipelines, but is it
-% the right tool?
-% But a quantitative evaluation is lacking.
+In-memory computing would then be a feature provided by 
+the operating system rather than by the engine itself. The frontier 
+between these two components is blurred and needs to be clarified.
 
-% More precisely, we intend to quantify the effect of in-memory computing, data locality, lazy evaluation
-% on various use cases. We intentionally leave scheduling out here.
+% https://ntrs.nasa.gov/archive/nasa/casi.ntrs.nasa.gov/20130001600.pdf
 
-% The use of a Big Data engine is not obvious when tmpfs and asynchronous writes
-% are available and could be used to emulate in-memory computing. Our goal is to quantify when and how Big Data engines will be useful.
-% Say that we don't use HDFS because we focus on clusters and HDFS on multi-tenant clusters is difficult (refer to hadoop on demand).
+Neuroimaging, our primary field of interest, is no exception to the 
+generalized rise of data volumes in science, due to the joint increase 
+of image resolution and subject cohort sizes~\cite{van2014human}. 
+Processing engines have been developed with neuroinformatics 
+applications in mind, for instance Nipype~\cite{gorgolewski2011nipype} 
+or PSOM~\cite{bellec2012pipeline}. Big Data engines have been used for 
+neuroimaging applications too, including in the Thunder 
+project~\cite{freeman2014mapping} and in more specific works such 
+as~\cite{makkie2019fast}. However, no quantitative performance 
+evaluation has been conducted on neuroimaging applications to assess the 
+added-value of Big Data engines compared to traditional processing engines.
 
-% Talk about pilot jobs and say we don't consider them here.
+This paper addresses the following questions:
+\begin{enumerate}
+\item What is the effect of in-memory computing and data locality on neuroimaging applications?
+\item Can in-memory computing be effectively enabled by the operating system rather than the data processing engine?
+\end{enumerate}
 
-% "```Disk and network I/O, of course, play a part in Spark performance 
-% as well, but neither Spark nor YARN currently do anything to actively 
-% manage them.```"
-% http://blog.cloudera.com/blog/2015/03/how-to-tune-your-apache-spark-jobs-part-2/
+Answers to these questions have important implications. 
+In~\cite{mehta2017comparative}, a comparative study of Dask, Spark, 
+TensorFlow, MyriaDB, and SciDB on neuroinformatics use-cases is 
+presented. It  concludes that these systems need to be extended to 
+better address scientific code integration, data partitioning, data 
+formats and system tuning. The work in~\cite{samba} also extended Spark 
+with a better provenance system for scientific applications. We argue 
+that such efforts should only be conducted if substantial performance 
+improvements are expected from in-memory computing, lazy 
+evaluation\todo{are we talking about lazy evaluation at all?} or data 
+locality. On the other hand, scientific data processing engines are 
+still being developed, and it is legitimate to wonder whether these 
+projects should just be migrated to Spark, Dask, or other Big Data 
+engines.
 
-% define pipeline and workflows
+Our study focuses on performance. We intentionally do not intend to 
+compare Big Data and scientific data processing engines on the grounds 
+of workflow language expressivity, provenance capture and 
+representation, portability or reproducibility, which are otherwise 
+critical concerns. Besides, our study of performance focuses on the 
+impact of data writes and transfers. It purposely leaves out task 
+scheduling to computing resources, to focus on the understanding of 
+data writes and movement. Task scheduling will be part of our 
+discussion though.
 
-% Here we focus on performance but we are aware of other desirable properties of workflow engines, including provenacne.
+\todo{Perhaps talk about modeling and simulation here.}
 
-% Outline
-
-% Talk about containers somewhere.
+Our methods, including performance models, processing engines, \todo{Make this paragraph more informative}
+applications and infrastructure used are described in 
+Section~\ref{sec:methods}. Section~\ref{sec:results} presents our 
+results which we discuss in Section~\ref{sec:discussion}. 
+Section~\ref{sec:conclusion} concludes on the answer to our two research questions.
 
 \section{Materials and Methods} % 4 pages
+\label{sec:methods}
 
@@
 This defines a data-compute ratio beyond which the page cache becomes 
-asymptotically useless, which we will use to define our benchmarks.
+asymptotically useless, which we will use to define our benchmarks. It 
+should be noted that leveraging the page cache, and therefore ensuring 
+that Equation~\ref{eq:page-cache-inequality} holds, has important 
+performance implications: with page caching, the application will 
+experience an almost negligible write time, while without page caching 
+the write throughput will be that of the disk.
 
 % Cache eviction: LRU/n.
+% page caching is mentioned in https://ntrs.nasa.gov/archive/nasa/casi.ntrs.nasa.gov/20130001600.pdf too
+% Note: writeback is dangerous as a system failure could result in data loss.
 
@@
 
+%~ In any case, I think we want the following inequality to hold:
+%~ ```
+%~ D/C <= \Delta / \Gamma
+%~ ```
+%~ where Delta is now the apparent bandwidth of the file system and Gamma 
+%~ is the max number of concurrent tasks in the cluster. Whether page 
+%~ caching is used or not will affect what happens when the inequality 
+%~ holds: if page caching is here, write time will be 0, otherwise it will 
+%~ be non-zero. If the inequality doesn't hold then we have congestion and 
+%~ it's a different game: transfer times might tend to infinity and data 
+%~ loss is to be expected.
+
+% Say that we don't use HDFS because we focus on clusters and HDFS on multi-tenant clusters is difficult (refer to hadoop on demand).
+
 % Make a table to explain where data locality and in-memory computing are
@@
 % What happens when dataset and derived data is too large for disk?
+
+In~\cite{saini2012performance} the Lustre file system is benchmarked on NASA applications.
 
@@
 
-% Talk about ~\cite{lebre2015adding}. Most simulation toolkits have focused on task 
+% Talk about ~\cite{lebre2015adding} and \cite{wrench}. Most simulation toolkits have focused on task 
 %  scheduling. And https://dl.acm.org/citation.cfm?id=3041715
 
-% And Wrench.
+\subsection{Task Scheduling}
+
+% Explain how we ensured that both engines lead to the same scheduling.
+% Talk about pilot jobs and say we don't consider them here.
 
@@
 % Lustre server
+
+\begin{table}
+\centering
+\begin{tabular}{c|c}
+$\delta/\gamma$ & $\Delta/\Gamma$\\
+\hline
+50~MB/s         & 5~MB/s
+\end{tabular}
+\label{table:infrastructure}
+\caption{Bandwidth thresholds on the target cluster}
+\end{table}
 
@@
 
+% Talk about our I/O pattern: random I/O, sequential.
+
+\begin{table}
+\centering
+\begin{tabular}{c|c}
+Application & $\frac{D}{C}$ (GB/s)\\
+\hline
+Big Brain            & [0-75] \\
+fMRI pre-processing  & [0-1.5]
+\end{tabular}
+\label{table:applications}
+\caption{Data-compute ratios for typical neuroinformatics applications}
+\end{table}
+% fMRI: about 100M in, 2GB out,2 30 min processing time to 10 hours
+
 \subsubsection{Execution modes} % perhaps find a better title, it's not informative
+
+We tuned job scheduling to balance the load among
+cluster nodes. This is the default behavior of Spark, but required
+specific instrumentation in Nipype.
 
@@
 \section{Results} % 2 pages
+\label{sec:results}
 
@@
 % Comment on the general trend
+% We can't really say that imaging tasks are not small. Even Freesurfer could be decomposed.
+%
 
@@
 \section{Discussion} % 2 pages
-
+\label{sec:discussion}
 \subsection{Effect of In-Memory Computing}
@@
 \subsection{Effect of Data Locality}
+% It should be important, in particular when there is contention and when
+% tasks are small.
 
@@
 \section{Conclusion} % 1 page with refs
+\label{sec:conclusion}
+% When is it useful to use a Big Data engine?
 
-% When is it useful to use a Big Data engine?
+% "```Disk and network I/O, of course, play a part in Spark performance 
+% as well, but neither Spark nor YARN currently do anything to actively 
+% manage them.```"
+% http://blog.cloudera.com/blog/2015/03/how-to-tune-your-apache-spark-jobs-part-2/
 
@@
 
+\todo{Acknowledge Dell and ask them if they want to co-author the paper.}
+
 \bibliographystyle{IEEEtran} 

### 2018-11-25 benchmark_scripts/nipype_sbatch.sh
fixed experimental conditions

@@ -1,7 +1,7 @@
 #!/bin/bash
 
 source /mnt/lustrefs/spark/env/exp0/bin/activate
-export PATH=/mnt/lustrefs/spark/SOEN691-project:$PATH
+export PATH=/mnt/lustrefs/spark/SOEN691-project/pipelines/:$PATH
 
 echo "{\"n_procs\": ${SLURM_CPUS_PER_TASK}}" > /mnt/lustrefs/spark/multiproc_conf.json
 

### 2018-11-25 benchmark_scripts/run_benchmarks.py

@@ -5,12 +5,15 @@ from random import shuffle
 from time import sleep
 import sys
 import subprocess
+import json
 
+
+num_nodes = 14
 slurm_spark = {
                "partition": "requestq",
                "reservation": "spark",
-               "time": "35:00:00",
-               "nodes": "16",
+               "time": "99:00:00",
+               "nodes": num_nodes + 1,
                "mem": "183G",
                "cpus-per-task": "40",
                "ntasks-per-node": "1"
@@ -19,7 +22,7 @@ slurm_spark = {
 slurm_nipype = {
                 "partition": "requestq",
                 "reservation": "spark",
-                "time": "35:00:00",
+                "time": "99:00:00",
                 "nodes": "1",
                 "mem": "183G",
                 "cpus-per-task": "1",
@@ -45,8 +48,8 @@ filesystems = {
                 'lustre': lustre
               }
 
-spark_script='/mnt/lustrefs/spark/SOEN691-project/spark_inc.py'
-nipype_script='/mnt/lustrefs/spark/SOEN691-project/nipype_inc.py'
+spark_script='/mnt/lustrefs/spark/SOEN691-project/pipelines/spark_inc.py'
+nipype_script='/mnt/lustrefs/spark/SOEN691-project/pipelines/nipype_inc.py'
 
 bb_125dir =  op.join(data_dir, '125_splits')
 bb_30dir = op.join(data_dir, '30_splits')
@@ -54,7 +57,9 @@ bb_750dir = op.join(data_dir, '750_splits')
 hbb_125dir = op.join(data_dir, 'half_bb/125_splits')
 mri_125dir = op.join(data_dir, 'mri/125_splits')
 
-num_nodes = 15
+data_names = {"bb_125dir": bb_125dir, "bb_30dir": bb_30dir,
+              "bb_750dir": bb_750dir, "hbb_125dir": hbb_125dir,
+              "mri_125dir": mri_125dir}
 
 legends = {
             bb_125dir: op.join(data_dir, '125_splits_legend.txt'),
@@ -63,412 +68,17 @@ legends = {
             hbb_125dir: op.join(data_dir, 'half_bb/125_splits_legend.txt'),
             mri_125dir: op.join(data_dir, 'mri/125_splits_legend.txt')
           }
-
-conditions = [
-              { 
-                  "framework": "spark",
-                  "filesystem": "mem",
-                  "delay": 55,
-                  "dataset": bb_125dir,
-                  "iterations": 1
-              },
-              {
[…]

### 2018-11-25 benchmark_scripts/spark_sbatch.sh

@@ -13,7 +13,7 @@ export SPARK_LOG_DIR=/mnt/lustrefs/spark/todelete-sparklogs
 export PYTHONIOENCODING=utf8
 
 source /mnt/lustrefs/spark/env/exp0/bin/activate
-export PATH=/mnt/lustrefs/spark/SOEN691-project:$PATH
+export PATH=/mnt/lustrefs/spark/SOEN691-project/pipelines:$PATH
 
 $SPARK_HOME/sbin/start-master.sh
 while [ -z "$MASTER_URL" ]

### 2018-11-25 benchmark_scripts/tmp_cleanup.sh

@@ -1,8 +1,8 @@
 #!/usr/bin/env bash
 #SBATCH --partition requestq       # Queue (Partition in slurm)
 #SBATCH --reservation=spark
-#SBATCH --time=00:05:00
-#SBATCH --nodes=16
+#SBATCH --time=10:00:00
+#SBATCH --nodes=15
 #SBATCH --mem=20G
 #SBATCH --cpus-per-task=1
 #SBATCH --ntasks-per-node=1

### 2018-11-25 pipelines/increment.py

@@ -5,59 +5,29 @@ import nibabel as nib
 from os import path as op, makedirs as md
 import time
 import subprocess
+import socket
+try:
+    from threading import get_ident
+except Exception as e:
+    from thread import get_ident
 
 
-def increment(fn, outdir, delay):
+
+def increment(fn, outdir, delay, benchmark_file, start_time):
     print('Incrementing image: ', fn)
-    '''
-    p = subprocess.Popen("iostat -y 1 1", shell=True, stdout=subprocess.PIPE,
-                      stderr=subprocess.PIPE)
-    (out, err) = p.communicate()
 
-    print(out.encode("utf-8"))
-    print(err.encode("utf-8"))
-    '''
-    start = time.time()
+    start = time.time() - start_time
     im = nib.load(fn)
-    print("read time", time.time() - start)
-    '''p = subprocess.Popen("iostat -y 1 1", shell=True,
-                         stdout=subprocess.PIPE, stderr=subprocess.PIPE)
-    (out, err) = p.communicate()
+    end = time.time() - start_time
+    print("read time", end)
+  
+    if benchmark_file is not None:
+        write_bench(benchmark_file, "Read file", start, end,
+                    socket.gethostname(), op.basename(fn), get_ident())
 
-    print(out.encode("utf-8"))
-    print(err.encode("utf-8"))'''
     inc_data = im.get_data() + 1
 
     im = nib.Nifti1Image(inc_data, affine=im.affine, header=im.header)
-
-    '''p = subprocess.Popen("top -b -n 1 | head -n 10 | tail -n 2", shell=True,
-                         stdout=subprocess.PIPE, stderr=subprocess.PIPE)
-    (out, err) = p.communicate()
-
-    print(out.encode("utf-8"))
-    print(err.encode("utf-8"))
-
-    p = subprocess.Popen("lsof /local", shell=True, stdout=subprocess.PIPE,
-                      stderr=subprocess.PIPE)
-    (out, err) = p.communicate()
-
-    print(out.encode("utf-8"))
-    print(err.encode("utf-8"))
-
-    p = subprocess.Popen("ps -ef", shell=True, stdout=subprocess.PIPE,
-                      stderr=subprocess.PIPE)
-    (out, err) = p.communicate()
-
-    print(out.encode("utf-8"))
-    print(err.encode("utf-8"))
-
-
-    p = subprocess.Popen("free", stdout=subprocess.PIPE,
[…]

### 2018-11-25 pipelines/nipype_inc.py

@@ -55,10 +55,22 @@ def increment_chunk(chunk, delay, benchmark, benchmark_dir=None, cli=False):
 
     else:
         program = 'increment.py'
-        p = subprocess.Popen([program, chunk, os.getcwd(),
-                              '--delay', str(delay)],
-                             stdout=subprocess.PIPE,
-                             stderr=subprocess.PIPE)
+
+        if benchmark:
+            benchmark_file = os.path.join(benchmark_dir,
+                                          "bench-{}.txt".format(
+                                              str(uuid.uuid1())))
+            p = subprocess.Popen([program, chunk, os.getcwd(),
+                                  '--benchmark_file', benchmark_file,
+                                  '--delay', str(delay)],
+                                 stdout=subprocess.PIPE,
+                                 stderr=subprocess.PIPE)
+        else:
+            p = subprocess.Popen([program, chunk, os.getcwd(),
+                                  '--delay', str(delay)],
+                                 stdout=subprocess.PIPE,
+                                 stderr=subprocess.PIPE)
+
         (out, err) = p.communicate()
         print(out)
         print(err)
@@ -131,7 +143,7 @@ def main():
     parser.add_argument('--cli', action='store_true',
                         help='use CLI application')
     parser.add_argument('--work_dir', type=str, help='working directory')
-    parser.add_argument('--delay', type=int, default=0,
+    parser.add_argument('--delay', type=float, default=0,
                         help='task duration time (in s)')
     parser.add_argument('--benchmark', action='store_true',
                         help='benchmark pipeline')
@@ -143,7 +155,7 @@ def main():
     args = parser.parse_args()
 
     start = time()
-    wf = Workflow('nipinc_bb')
+    wf = Workflow('npinc_bb')
 
     if args.work_dir is not None:
         wf.base_dir = os.path.abspath(args.work_dir)

### 2018-11-25 pipelines/spark_inc.py

@@ -45,6 +45,7 @@ def read_img(filename, data, benchmark, start, output_dir, bench_dir=None):
 
     start_time = time() - start
 
+    print('Reading image')
     # load binary data into Nibabel
     fh = nib.FileHolder(fileobj=BytesIO(data))
     im = nib.Nifti1Image.from_file_map({'header': fh, 'image': fh})
@@ -68,9 +69,23 @@ def read_img(filename, data, benchmark, start, output_dir, bench_dir=None):
 def increment_data(idx, filename, data, metadata, delay, benchmark, start,
                    output_dir, iteration=0, work_dir=None, bench_file=None,
                    cli=False):
+
     start_time = time() - start
 
+    if benchmark and os.path.isdir(bench_file):
+        bench_dir = bench_file
+        try:
+            os.makedirs(bench_dir)
+        except Exception as e:
+            pass
+
+        bench_file = os.path.join(
+                        bench_dir,
+                        "bench-{}.txt".format(str(uuid.uuid1()))
+                     )
+
     if not cli:
+        print("Incrementing data in memory")
         data += 1
         sleep(delay)
     else:
@@ -86,8 +101,15 @@ def increment_data(idx, filename, data, metadata, delay, benchmark, start,
 
         fn = filename[5:] if 'file:' in filename else os.path.abspath(filename)
 
-        p = subprocess.Popen(['increment.py', fn,
-                              work_dir, '--delay', str(delay)])
+        if benchmark:
+            p = subprocess.Popen(['increment.py', fn,
+                                  work_dir, '--benchmark_file', bench_file,
+                                  '--start', str(start),
+                                  '--delay', str(delay)])
+        else:
+            p = subprocess.Popen(['increment.py', fn,
+                                  work_dir, '--delay', str(delay)])
+
         (out, err) = p.communicate()
 
         fn = os.path.basename(fn)
@@ -100,19 +122,12 @@ def increment_data(idx, filename, data, metadata, delay, benchmark, start,
 
     bench_dir = None
     if benchmark:
-        if os.path.isdir(bench_file):
-            bench_dir = bench_file
-            bench_file = None
-
         bn = os.path.basename(filename)
         write_bench('increment_data', start_time, end_time,
                     socket.gethostname(), output_dir, bn,
                     get_ident(), bench_dir, bench_file)
 
-    if bench_file is not None:
-        return (idx, filename, data, metadata, bench_file, iteration + 1)
-    else:
[…]

### 2018-11-25 tests/test_inc.py

@@ -70,12 +70,12 @@ def test_increment_spark_cli():
 
 def test_increment_nipype():
 
-    shutil.rmtree('nipinc_out', ignore_errors=True)
+    shutil.rmtree('npinc_out', ignore_errors=True)
     p = subprocess.Popen(['python', 'pipelines/nipype_inc.py', 'sample_data',
-                          'nipinc_out', '1'])
+                          'npinc_out', '1'])
     (out, err) = p.communicate()
 
-    h_prog_1 = hashlib.md5(open('nipinc_out/inc1-dummy_1.nii', 'rb').read()) \
+    h_prog_1 = hashlib.md5(open('npinc_out/inc1-dummy_1.nii', 'rb').read()) \
                       .hexdigest()
     h_exp_1 = hashlib.md5(open('tests/test_outputs/testinc_1_1.nii', 'rb')
                           .read()) \
@@ -83,12 +83,12 @@ def test_increment_nipype():
 
     assert h_prog_1 == h_exp_1
 
-    shutil.rmtree('nipinc_out')
+    shutil.rmtree('npinc_out')
     p = subprocess.Popen(['python', 'pipelines/nipype_inc.py', 'sample_data',
-                          'nipinc_out', '10'])
+                          'npinc_out', '10'])
     p.communicate()
 
-    h_prog_10 = hashlib.md5(open('nipinc_out/inc10-dummy_1.nii', 'rb')
+    h_prog_10 = hashlib.md5(open('npinc_out/inc10-dummy_1.nii', 'rb')
                             .read()) \
                        .hexdigest()
     h_exp_10 = hashlib.md5(open('tests/test_outputs/testinc_1_10.nii', 'rb')
@@ -100,13 +100,13 @@ def test_increment_nipype():
 
 def test_increment_nipype_cli():
 
-    shutil.rmtree('nipinc_out', ignore_errors=True)
+    shutil.rmtree('npinc_out', ignore_errors=True)
     p = subprocess.Popen(['python', 'pipelines/nipype_inc.py', 'sample_data',
-                          'nipinc_out', '1', '--cli', '--work_dir',
-                          'nipinc_tmp'])
+                          'npinc_out', '1', '--cli', '--work_dir',
+                          'npinc_tmp'])
     (out, err) = p.communicate()
 
-    h_prog_1 = hashlib.md5(open('nipinc_out/inc1-dummy_1.nii', 'rb').read()) \
[…]

### 2018-11-26 benchmark_scripts/run_benchmarks.py
bug fixes and benchmark changes

@@ -8,7 +8,7 @@ import subprocess
 import json
 
 
-num_nodes = 14
+num_nodes = 15
 slurm_spark = {
                "partition": "requestq",
                "reservation": "spark",
@@ -107,10 +107,10 @@ for cdn in conditions:
         slurm_conf["cpus-per-task"] = 9
     elif cdn["dataset"] == bb_30dir:
         cdn_ident += "30BB"
-        slurm_conf["cpus-per-task"] = 3
+        slurm_conf["cpus-per-task"] = 2
     elif cdn["dataset"] == bb_750dir:
         cdn_ident += "750BB"
-        slurm_conf["cpus-per-task"] = 27
+        slurm_conf["cpus-per-task"] = 25
     elif cdn["dataset"] == mri_125dir:
         cdn_ident += "125MRI"
         slurm_conf["cpus-per-task"] = 9

### 2018-11-26 benchmark_scripts/tmp_cleanup.sh

@@ -2,15 +2,14 @@
 #SBATCH --partition requestq       # Queue (Partition in slurm)
 #SBATCH --reservation=spark
 #SBATCH --time=10:00:00
-#SBATCH --nodes=15
+#SBATCH --nodes=16
 #SBATCH --mem=20G
 #SBATCH --cpus-per-task=1
 #SBATCH --ntasks-per-node=1
 
-#srun ls /local/
 srun sync
 srun echo 1 > /proc/sys/vm/drop_caches
 srun rm -rf /dev/shm/inc_exp/work
 srun rm -rf /local/inc_exp/work
 srun rm -rf /mnt/lustrefs/spark/inc_exp/work
-
+srun ls /local/

### 2018-11-26 pipelines/increment.py

@@ -66,7 +66,7 @@ def main():
                         help='the file to write benchmarks to')
     parser.add_argument('--start', type=float, default=0,
                         help='start time of the application')
-    parser.add_argument('--delay', type=int, default=0,
+    parser.add_argument('--delay', type=float, default=0,
                         help='task duration time (in s)')
 
     args = parser.parse_args()

### 2018-11-26 paper.tex

@@
 \usepackage{color}
+\usepackage{colortbl} % for \rowcolor
+
 \usepackage[bookmarks=false]{hyperref}
@@
       #1}}}
+\definecolor{headcolor}{gray}{0.9}
 
@@
 in-memory computing, and lazy evaluation to reduce the impact of data 
-manipulation on the performance of applications. From early MapReduce 
-implementations~\cite{dean2008mapreduce} to Apache Spark~\cite{zaharia2016apache} and now 
-Dask~\cite{rocklin2015dask}, these frameworks have become 
+manipulation on the performance of applications. \todo{We need to 
+define data locality, in-memory computing and lazy evaluation somewhere 
+in the intro.} From early MapReduce 
+implementations~\cite{dean2008mapreduce} to Apache 
+Spark~\cite{zaharia2016apache} and now Dask~\cite{rocklin2015dask}, 
+these frameworks have become 
 mainstream tools for data analytics. Meanwhile, several scientific 
@@
 data is fetched from data nodes to compute nodes using shared file 
-systems such as Lustre~\cite{schwan2003lustre}. Such differences in applications 
-and infrastructures have important consequences. To mention only one, 
-in-memory computing requires instrumentation to be applied to 
-command-line tools.
+systems such as Lustre~\cite{schwan2003lustre}. Such differences in 
+applications and infrastructures have important consequences. To 
+mention only one, in-memory computing requires instrumentation to be 
+applied to command-line tools. \todo{Also mention that performance is 
+only one of the concerns of scientific processing engines, which might
+explain why Big Data engines are overlooked.}
 
@@
 applications. 
-
 In-memory computing would then be a feature provided by 
@@
 compare Big Data and scientific data processing engines on the grounds 
-of workflow language expressivity, provenance capture and 
+of workflow language expressivity, fault-tolerance, provenance capture and 
 representation, portability or reproducibility, which are otherwise 
@@
 Section~\ref{sec:conclusion} concludes on the answer to our two research questions.
+
+\newpage % Remove this when abstract is written
 
@@
 % memory management (bad?)
+% Nipype only has data locality becausae we use multiproc. Data locality is emulated by manual decisions on scheduling.
 
-\subsection{Performance Models}
+\subsection{Data Storage Locations}
 
-\subsubsection{In-Memory File System}
+Data storage location is critical to the performance of Big Data 
+applications on HPC clusters. Data may reside in the engine memory, on 
+a file system whose content reside in virtual memory (for instance 
+\texttt{tmpfs}), on disks local to the processing node, or on a shared 
+file system. Table~\ref{table:features} summarizes the Big Data 
+strategies that can be used depending on the data location. The 
+remainder of this Section explains this Table and provides related 
+performance models.
+\begin{table}
+\centering
+\begin{tabular}{c|ccc}
+   \rowcolor{headcolor}
+    Data Location                 & In-Memory     & Data Locality & Lazy \\
+    \rowcolor{headcolor}
+                                  & Computing     &               & Evaluation      \\
+                                  \hline          
+In-memory                         &   Y           & Y             & Spark               \\
+\texttt{tmpfs}                    &   Y           & Y             & Spark           \\
+Local Disk                        & Page Caching  & Y             & Spark           \\
+Shared File System                & Page Caching  & N            & Spark          
+% data locality is somehow present in shared fs due to page caching again.
+\end{tabular}
+\caption{Big Data strategies on a shared HPC cluster.}
+\label{table:features}
+\end{table}
+
+\subsubsection{In-Engine-Memory and In-Memory File System}
 
@@
 
-% 15 cores per executor can lead to bad I/O throughput
-% http://blog.cloudera.com/blog/2015/03/how-to-tune-your-apache-spark-jobs-part-2/
-
 % measured disk bandwidth with random I/O, very different from sequential
 
-The Linux kernel uses in-memory page caching to speed-up disk writes. 
-When a process calls the \texttt{write()} system call, data is copied 
-to a memory cache that is asynchronously written to disk by flusher 
-threads when memory shrinks, ``dirty" (unwritten) data grows, or a 
-process invokes the \texttt{sync()} system call~\cite{love2010linux}. 
-The flushing of page cache to disk is also called the writeback 
-process.
+Storing data on local disks obviously enables data locality as tasks 
+executed on the node where their input files reside do not require to 
+transfer input or output data. However, in absence of a more specific 
+file system such as the Hadoop Distributed File System (HDFS) to handle 
+file replication across computing nodes, data locality comes at the price
+of stringent scheduling restrictions, as tasks can only be scheduled to the
+single node that contains their input data.
+% this brings data locality but not at the level of HDFS because replication isn't there and 
+% there is only 1 single option to schedule a task while preserving data locality.
 
-Page caching has a dramatic effect on I/O, reducing the duration of 
-disk writes by several orders of magnitude. It is essentially a way to 
+The performance of local disk accesses is strongly dependent on the 
+page caching mechanism provided by the Linux kernel, described in great 
+details in~\cite{love2010linux}. To summarize, data read from disk 
+remains cached in memory until evicted by an LRU (Least Recently Used) 
+strategy. When a process invokes the \texttt{read()} system call, the 
+kernel will return the data directly from memory if the requested data 
+lies in the page cache. Such cache hits drastically speed-up data 
+reads, by masking the disk latency and limited bandwidth behind a 
+memory buffer. In effect, page caching provides in-memory computing 
+transparently to the processing engine. However, page cache eviction 
+strategies currently cannot be controlled by the application 
+\todo{check that}, which prevents processing engines to anticipate 
+reads by preloading the cache. Scheduling strategies might be designed 
+to maximize cache hits though. For instance, we hypothesize that lazy 
+evaluation would result in more cache hits by scheduling data-dependent 
+tasks on the same node.
+
+Page caching has a more dramatic effect on disk writes, reducing their 
+duration by several orders of magnitude. When a process calls the 
+\texttt{write()} system call, data is copied to a memory cache that is 
+asynchronously written to disk by flusher threads when memory shrinks, 
+``dirty" (unwritten) data grows, or a 
+process invokes the \texttt{sync()} system call. 
+This asynchronous flushing of the page cache is called 
+\emph{writeback}.
+%~ It should be noted that writeback may lead to data 
+%~ loss in case of system failure.
+
+Page caching is essentially a way to 
 emulate in-memory computing at the kernel level, without requiring a 
@@
 this happens, the page cache rapidly fills up and writes are limited by 
-the disk write bandwidth -- typically 500~MB/s for random writes to 
+the disk write bandwidth -- about 500~MB/s for random writes to 
 contemporary SSDs  -- as if no page cache was involved.
@@
 This defines a data-compute ratio beyond which the page cache becomes 
-asymptotically useless, which we will use to define our benchmarks. It 
+asymptotically useless. We will use this ratio to define our benchmarks. It 
 should be noted that leveraging the page cache, and therefore ensuring 
@@
 performance implications: with page caching, the application will 
-experience an almost negligible write time, while without page caching 
+experience negligible write times, while without page caching 
 the write throughput will be that of the disk.
@@
 % Cache eviction: LRU/n.
-% page caching is mentioned in https://ntrs.nasa.gov/archive/nasa/casi.ntrs.nasa.gov/20130001600.pdf too
-% Note: writeback is dangerous as a system failure could result in data loss.
+
+\todo{How about reads?}
 
 \subsubsection{Shared File System}
+
+We model a shared file system using its global apparent bandwidth 
+$\Delta$, shared by all concurrent processes in the system. We are 
+aware that such a simplistic model does not describe at all the 
+intricacies of systems such as Lustre. In particular, metadata 
+management, RPC protocol optimizations and storage optimizations are 
+all covered under the apparent bandwidth. We do, however, consider the 
+effect of page caching in shared file systems too, since in Linux 
+writes to network-mounted volumes benefit from this feature too.
+
+As in the local disk model, we note that page caching will only be 
+useful when the flush bandwidth is greater than the write throughput of 
+the application, that is:
+\begin{equation}
+\frac{D}{C} \leq \frac{\Delta}{\Gamma}, \label{eq:page-cache-sharedfs}
+\end{equation}
+where $\Gamma$ is the max number of concurrent jobs in the cluster. 
+Note that $\frac{\Delta}{\Gamma}$ will be much lower than 
+$\frac{\delta}{\gamma}$ in Equation~\ref{eq:page-cache-inequality}.
 
@@
 
-% Say that we don't use HDFS because we focus on clusters and HDFS on multi-tenant clusters is difficult (refer to hadoop on demand).
+% Say that we don't use HDFS because we focus on multi-tenant clusters and HDFS on multi-tenant clusters is difficult (refer to hadoop on demand).
 
@@
 
+A common recommendation in Spark is to limit the number of cores per 
+executors to 5, to preserve a good I/O 
+throughput\footnote{\url{http://blog.cloudera.com/blog/2015/03/how-to-tune-your-apache-spark-jobs-part-2}}. 
+We believe that throughput degradation observed with more than 5 cores 
+per executor might be coming from full page caches.
+
 \subsection{Other Comments}
 
-% fault-tolerance: do we have anything to discuss here?
+% fault-tolerance: if you loose a node, you loose the data when it's written on disk or in memory.
+%                  then the only solution is to recompute, which spark would do automatically.
+%                  when nipype is run on lustre, it would avoid recomputing existing files, which Spark doesn't do.
+%                   this niype feature is lost when not using lustre.
+
+% Comment on the practical implications of using tmpfs or local disk: we might not get all the intermediary files.
+
+% mention burst buffers or heterogeneous storage managers as potential solutions to write to local disk, 
+% i.e., to benefit from data locality while getting your results on Lustre.
 

### 2018-11-27 results/code/bench_to_json.py
added benchmarks and figure generation files

@@ -1,0 +1,38 @@
+import json
+import sys
+
+start_time = None
+
+if len(sys.argv) > 3 and sys.argv[3] == '--nipype':
+    with open(sys.argv[1], 'r') as f:
+        for line in f:
+            l = line.split(' ')
+
+            if l[0] == 'driver_program' and (start_time is None 
+                                             or float(l[1]) < start_time):
+                start_time = float(l[1])
+
+if start_time is None:
+    start_time = 0
+
+with open(sys.argv[1], 'r') as f:
+    bench = {}
+    bench['tasks'] = []
+    count = 0
+
+    driver_start = None
+    for line in f:
+        l = line.split(' ')
+        node = {}
+        node['name'] = l[0]
+        node['id'] = count
+        node['start-time'] = float(l[1]) - start_time
+        node['end-time'] = float(l[2]) - start_time
+        node['node'] = l[3]
+
+        bench['tasks'].append(node)
+        count += 1
+
+    with open(sys.argv[2], 'w') as j:
+        json.dump(bench, j)
+

### 2018-11-27 results/code/gen_figures.py

@@ -1,0 +1,108 @@
+#!/usr/bin/env python
+import matplotlib.pyplot as plt
+import sys
+import argparse
+
+
+def genfig(exp, makef, outf, cond):
+    bench_res = {('sp', 'mem'): ([], []),
+                 ('sp', 'tmpfs'): ([], []),
+                 ('sp', 'local'): ([], []),
+                 ('sp', 'lustre'): ([], []),
+                 ('np', 'tmpfs'): ([], []),
+                 ('np', 'local'): ([], []),
+                 ('np', 'lustre'): ([], [])
+                }
+
+    with open(makef, 'r') as f:
+        for line in f:
+            filename, bench = line.split(':')
+            filename = filename[:-4]
+            conditions = filename.split('_')
+            delay = float(conditions[3][:-5])
+            engine = conditions[0]
+            fs = conditions[1]
+            iterations, chunks = conditions[2].split('it')
+
+            if 'HBB' in chunks or 'MRI' in chunks:
+                im = chunks[3:]
+                chunks = 125
+            else:
+                im = 'BB'
+                chunks = int(chunks[:-2])
+       
+            iterations = int(iterations)
+
+            if exp == 1:
+                if cond is not None and delay == cond:
+                    bench_res[(engine, fs)][0].append(iterations)
+                    bench_res[(engine, fs)][1].append(float(
+                        bench.split(' ')[2]))
+                elif cond is None and delay == 3.44:
+                    bench_res[(engine, fs)][0].append(iterations)
+                    bench_res[(engine, fs)][1].append(float(
+                        bench.split(' ')[2]))
+            elif exp == 2:
+                bench_res[(engine, fs)][0].append(int(chunks))
+                bench_res[(engine, fs)][1].append(float(bench.split(' ')[2]))
+            elif exp == 3:
+                bench_res[(engine, fs)][0].append(im)
+                bench_res[(engine, fs)][1].append(float(bench.split(' ')[2]))
+            else:
+                if cond is not None and iterations == cond:
+                    bench_res[(engine, fs)][0].append(delay)
+                    bench_res[(engine, fs)][1].append(float(
+                        bench.split(' ')[2]))
+                elif cond is None and iterations == 10:
+                    bench_res[(engine, fs)][0].append(delay)
+                    bench_res[(engine, fs)][1].append(float(
+                        bench.split(' ')[2]))
+
[…]

### 2018-11-27 results/code/get_makespan.py

@@ -1,0 +1,37 @@
+import sys
+import os
+
+# clear contents of file
+with open(sys.argv[2], 'w+') as f:
+    pass
+
+for bench in os.listdir(sys.argv[1]):
+    start = None
+    end = None
+    driver_line = None
+    if 'np' in bench:
+        with open(os.path.join(sys.argv[1], bench), 'r') as f:
+            for line in f:
+                if 'driver_program' in line:
+                    row = line.split(' ')
+                    start_time = float(row[1])
+                    end_time = float(row[2])
+                    if start is None or start_time < start:
+                        driver_line = row
+                        start = start_time
+                    if end is None or end_time < end:
+                        end = end_time
+                        driver_line[2] = str(end - start)
+                        driver_line[1] = "0"
+
+            driver_line[0] = os.path.basename(bench) + ":" + driver_line[0]
+
+    else:
+        with open(os.path.join(sys.argv[1], bench), 'r') as f:
+            driver_line = f.readline().split(" ")
+            driver_line[0] = os.path.basename(bench) + ":" + driver_line[0]
+            
+
+    with open(sys.argv[2], 'a') as f:
+        f.write(" ".join(driver_line))
+

### 2018-11-27 results/code/makespan_to_json.py

@@ -1,0 +1,29 @@
+import json
+import sys
+
+
+with open(sys.argv[1], 'r') as f:
+    bench = {}
+    bench['tasks'] = []
+    count = 0
+
+    driver_start = None
+    for line in f:
+
+        experiment = line.split(':')[0]
+        l = line.split(' ')
+        l[0] = l[0].replace(experiment + ':', '')
+        print(l)
+        node = {}
+        node['name'] = l[0]
+        node['id'] = count
+        node['start-time'] = float(l[1])
+        node['end-time'] = float(l[2]) - node['start-time']
+        node['node'] = experiment
+
+        bench['tasks'].append(node)
+        count += 1
+
+    with open(sys.argv[2], 'w') as j:
+        json.dump(bench, j)
+

### 2018-11-27 results/code/gen_figures.py
some cleanup

@@ -36,27 +36,19 @@ def genfig(exp, makef, outf, cond):
             if exp == 1:
                 if cond is not None and delay == cond:
                     bench_res[(engine, fs)][0].append(iterations)
-                    bench_res[(engine, fs)][1].append(float(
-                        bench.split(' ')[2]))
                 elif cond is None and delay == 3.44:
                     bench_res[(engine, fs)][0].append(iterations)
-                    bench_res[(engine, fs)][1].append(float(
-                        bench.split(' ')[2]))
             elif exp == 2:
                 bench_res[(engine, fs)][0].append(int(chunks))
-                bench_res[(engine, fs)][1].append(float(bench.split(' ')[2]))
             elif exp == 3:
                 bench_res[(engine, fs)][0].append(im)
-                bench_res[(engine, fs)][1].append(float(bench.split(' ')[2]))
             else:
                 if cond is not None and iterations == cond:
                     bench_res[(engine, fs)][0].append(delay)
-                    bench_res[(engine, fs)][1].append(float(
-                        bench.split(' ')[2]))
                 elif cond is None and iterations == 10:
                     bench_res[(engine, fs)][0].append(delay)
-                    bench_res[(engine, fs)][1].append(float(
-                        bench.split(' ')[2]))
+
+            bench_res[(engine, fs)][1].append(float(bench.split(' ')[2]))
 
         plt.plot(bench_res[('sp', 'mem')][0],
                  bench_res[('sp', 'mem')][1], 

### 2018-11-27 results/code/d_c_graph.py
Code to make a D-C graph

@@ -1,0 +1,150 @@
+#!/usr/bin/env python
+
+import utils
+import sys  
+import os
+import argparse
+
+def parse_bench_file_name(file_name):
+    # Assumes name looks like sp_lustre_100it125BB_2.4delay.out
+    strings = file_name.split('_')
+    wf = strings[0]
+    fs = strings[1]
+    delay = float(strings[3].replace('delay.out', ''))
+    stringsit = strings[2].split('it')
+    iterations = int(stringsit[0])
+    blocks = 30
+    if '125' in stringsit[1]:
+        blocks = 125
+    if '750' in stringsit[1]:
+        blocks = 750
+    data = 'MRI'
+    if 'HB' in stringsit[1]:
+        data = 'HB'
+    if 'BB' in stringsit[1]:
+        data = 'BB'
+    
+    return iterations, data, fs, wf, delay, blocks
+
+def d_c_fig(bench_dir, makespan_file, out_file):
+    sizes = {  # in MiB
+        'MRI': 13016/1024,
+        'HBB': 39284504/1024,
+        'BB': 78568516/1024
+    }
+    
+    bandwidths = { # MiB/s
+        'lustre': 401.49,
+        'local': 600, # TODO: get the actual value
+        'tmpfs': 1793.75
+    }
+    
+    gammas = { # by number of    blocks
+        30: 2,
+        125: 9,
+        750: 25
+    }
+    
+    Gammas = { # by number of blocks
+        30: 30,
+        125: 125,
+        750: 350
+    }
+    
+    benches = {}
+    for file_name in os.listdir(bench_dir):
+        it, data, fs, wf, delay, blocks = parse_bench_file_name(file_name)
+        benches[file_name] = {
+            'iterations': it,
+            'data_file': data,
+            'file_system': fs,
+            'blocks': blocks,
+            'wf': wf,
+            'task_duration': delay,
+            'C': delay*it*blocks,
+            'D': sizes[data]*it,
+            'g': gammas[blocks],
+            'G': Gammas[blocks]
+        }
+        benches[file_name]
+    
+    # Set makespans
+    with open(makespan_file, 'r') as f:
+        for line in f:
+            splits = line.split(' ')
+            file_name = splits[0].split(':')[0]
+            makespan = float(splits[2])
+            benches[file_name]['makespan'] = makespan
+    
+    
+    # Set memory speed-ups
+    for file_name in os.listdir(bench_dir):
[…]

### 2018-11-27 results/code/d_c_graph.py
Fixed HBB bug

@@ -19,9 +19,9 @@ def parse_bench_file_name(file_name):
     if '750' in stringsit[1]:
         blocks = 750
     data = 'MRI'
-    if 'HB' in stringsit[1]:
-        data = 'HB'
-    if 'BB' in stringsit[1]:
+    if 'HBB' in stringsit[1]:
+        data = 'HBB'
+    elif 'BB' in stringsit[1]:
         data = 'BB'
     
     return iterations, data, fs, wf, delay, blocks

### 2018-11-27 pipelines/increment.py
minor fixes and bandwidth benchmarking

@@ -22,7 +22,7 @@ def increment(fn, outdir, delay, benchmark_file, start_time):
     print("read time", end)
   
     if benchmark_file is not None:
-        write_bench(benchmark_file, "Read file", start, end,
+        write_bench(benchmark_file, "read_file", start, end,
                     socket.gethostname(), op.basename(fn), get_ident())
 
     inc_data = im.get_data() + 1
@@ -41,7 +41,7 @@ def increment(fn, outdir, delay, benchmark_file, start_time):
     print("write time", time.time() - start)
 
     if benchmark_file is not None:
-        write_bench(benchmark_file, "Write file", start, end,
+        write_bench(benchmark_file, "write_file", start, end,
                     socket.gethostname(), op.basename(out_fn), get_ident())
 
     time.sleep(delay)

### 2018-11-27 results/code/avg_io_times.py

@@ -1,0 +1,71 @@
+#!/usr/bin/env python
+from os import listdir
+from os.path import join
+import sys
+import matplotlib.pyplot as plt
+
+
+read_bandwidths = write_bandwidths = {
+                                      'tmpfs': [],
+                                      'local': [],
+                                      'lustre': [],
+                                     }
+
+for bench_f in listdir(sys.argv[1]):
+    filename = bench_f[:-4]
+    conditions = filename.split('_')
+    # delay = float(conditions[3][:-5])
+    # engine = conditions[0]
+    fs = conditions[1]
+    # iterations, chunks = conditions[2].split('it')
+    chunks = conditions[2].split('it')[1]
+
+    im_size = 13
+    im = 'MRI'
+
+    if fs == 'mem': continue
+    if '125' not in chunks: continue
+
+    if 'HB' in chunks:
+        continue
+        im_size = 38364
+        im = 'HBB'
+    elif 'BB' in chunks:
+        im_size = 76727
+        im = 'BB'
+    else:
+        continue
+           
+    # iterations = int(iterations)
+
+    fullpath = join(sys.argv[1], bench_f)
+
+    with open(fullpath, 'r') as f:
+        for line in f:
+            data = line.split(' ')
+           
+            io_time = float(data[2]) - float(data[1])
+            if data[0] == 'read_file':
+                if io_time == 0:
+                    read_bandwidths[fs].append(im_size/0.001)
+                else:
+                    read_bandwidths[fs].append(im_size/io_time)
+
+            if data[1] == 'write_file':
+                if io_time == 0:
+                    write_bandwidths[fs].append(im_size/0.001)
+                else:
+                    write_bandwidths[fs].append(im_size/io_time)
+
+plt.boxplot([read_bandwidths['tmpfs'], read_bandwidths['local'],
+             read_bandwidths['lustre']], 0, '')
+plt.ylim(7.66*10**7, 7.68*10**7)
+plt.xlabel(['tmpfs', 'local', 'lustre'])
+plt.savefig('read_bandwidths')
+
+plt.figure()
+plt.boxplot([write_bandwidths['tmpfs'], write_bandwidths['local'],
+             write_bandwidths['lustre']], 0, '')
+plt.ylim(7.66*10**7, 7.68*10**7)
+plt.xlabel(['tmpfs', 'local', 'lustre'])
+plt.savefig('write_bandwidths')

### 2018-11-28 pipelines/spark_inc.py
minor fixes, additional benchmarks  and figure generating scripts improvements

@@ -230,6 +230,7 @@ def main():
         files = glob(os.path.join(args.bb_dir, '*'))
         fidx = [i for i in range(0, len(files))]
         imRDD = sc.parallelize(zip(fidx, files), len(files))
+        if args.work_dir is None: args.work_dir = output_dir
         work_dir = os.path.abspath(os.path.join(args.work_dir,
                                                 'app-{}'.format(app_uuid)))
         print(work_dir)

### 2018-11-28 results/code/avg_io_times.py

@@ -5,40 +5,66 @@ import sys
 import matplotlib.pyplot as plt
 
 
-read_bandwidths = write_bandwidths = {
-                                      'tmpfs': [],
-                                      'local': [],
-                                      'lustre': [],
-                                     }
+read_bandwidths = {
+                   'tmpfs': [],
+                   'local': [],
+                   'lustre': [],
+                  }
+write_bandwidths = {
+                    'tmpfs': [],
+                    'local': [],
+                    'lustre': [],
+                   }
+spr_bandwidths = {
+                  'tmpfs': [],
+                  'local': [],
+                  'lustre' : []
+                 }
+npr_bandwidths = {
+                  'tmpfs': [],
+                  'local': [],
+                  'lustre' : []
+                 }
+spw_bandwidths = {
+                  'tmpfs': [],
+                  'local': [],
+                  'lustre' : []
+                 }
+npw_bandwidths = {
+                  'tmpfs': [],
+                  'local': [],
+                  'lustre' : []
+                 }
+
+
+rb_outf = 'read_bandwidth.pdf'
+wb_outf = 'write_bandwidth.pdf'
+snrb_out = 'nsp-' + rb_outf
+snwb_out = 'nsp-' + wb_outf
 
 for bench_f in listdir(sys.argv[1]):
     filename = bench_f[:-4]
+    fullpath = join(sys.argv[1], bench_f)
+
     conditions = filename.split('_')
-    # delay = float(conditions[3][:-5])
-    # engine = conditions[0]
+    engine = conditions[0]
+
     fs = conditions[1]
-    # iterations, chunks = conditions[2].split('it')
     chunks = conditions[2].split('it')[1]
 
-    im_size = 13
+    chunk_size = 102752/(1024**2)
     im = 'MRI'
 
     if fs == 'mem': continue
-    if '125' not in chunks: continue
 
     if 'HB' in chunks:
-        continue
-        im_size = 38364
+        chunk_size = 321813010/(1024**2)
         im = 'HBB'
     elif 'BB' in chunks:
         im_size = 76727
+        num_chunks = int(chunks[:-2])
+        chunk_size = im_size/num_chunks
         im = 'BB'
-    else:
-        continue
[…]

### 2018-11-28 results/code/d_c_graph.py

@@ -35,7 +35,7 @@ def d_c_fig(bench_dir, makespan_file, out_file):
     
     bandwidths = { # MiB/s
         'lustre': 401.49,
-        'local': 600, # TODO: get the actual value
+        'local': 1480.2, # TODO: get the actual value
         'tmpfs': 1793.75
     }
     
@@ -84,7 +84,7 @@ def d_c_fig(bench_dir, makespan_file, out_file):
                                                              b['blocks'],
                                                              b['data_file'],
                                                              b['task_duration'])
-        if benches.get(in_mem_file) == None:
+        if benches.get(in_mem_file) == None or 'MRI' in in_mem_file:
             continue # file isn't in benchmark
         b['memory-speed-up'] = (
             b['makespan'] / benches[in_mem_file]['makespan']
@@ -120,8 +120,7 @@ def d_c_fig(bench_dir, makespan_file, out_file):
     from matplotlib import pyplot as plt
     #plt.plot(x_mem, y_mem, 'o', label="In memory")
     #plt.plot(x_tmpfs, y_tmpfs, 'o', label="tmpfs")
-    plt.plot(x_disk, y_disk, '+', label="Local Disk")
-    plt.plot(x_sfs, y_sfs, '+', label="Lustre")
+    plt.plot(x_disk, y_disk, 'b+', label="Local Disk")
     rect = plt.Rectangle([1, 0], 150, 1, color='gray', edgecolor=None)
     plt.gca().add_patch(rect)
     rect = plt.Rectangle([0, 1], 1, 5, color='gray', edgecolor=None)
@@ -132,7 +131,22 @@ def d_c_fig(bench_dir, makespan_file, out_file):
     plt.ylim(0)
     plt.xlim(0)
     #plt.show()
-    plt.savefig(out_file)
+    plt.savefig(os.path.join(os.path.dirname(out_file), 
+                'local-{}'.format(os.path.basename(out_file))))
+    plt.figure()
+    plt.plot(x_sfs, y_sfs, 'r+', label="Lustre")
+    rect = plt.Rectangle([1, 0], 150, 1, color='gray', edgecolor=None)
+    plt.gca().add_patch(rect)
+    rect = plt.Rectangle([0, 1], 1, 5, color='gray', edgecolor=None)
+    plt.gca().add_patch(rect)
+    plt.xlabel("(D/C) / (d(D)elta/g(G)amma)")
[…]

### 2018-11-28 results/code/gen_figures.py

@@ -3,16 +3,22 @@ import matplotlib.pyplot as plt
 import sys
 import argparse
 
+exp14_delays = [2.4, 3.44, 7.68, 320]
+exp2_delays = [0.59, 3.52, 14.67]
+exp3_delays = 1.76
 
 def genfig(exp, makef, outf, cond):
-    bench_res = {('sp', 'mem'): ([], []),
-                 ('sp', 'tmpfs'): ([], []),
-                 ('sp', 'local'): ([], []),
-                 ('sp', 'lustre'): ([], []),
-                 ('np', 'tmpfs'): ([], []),
-                 ('np', 'local'): ([], []),
-                 ('np', 'lustre'): ([], [])
+    bench_res = {('sp', 'mem'): [],
+                 ('sp', 'tmpfs'): [],
+                 ('sp', 'local'): [],
+                 ('sp', 'lustre'): [],
+                 ('np', 'tmpfs'): [],
+                 ('np', 'local'): [],
+                 ('np', 'lustre'): []
                 }
+
+    if exp == 1 and cond is None: cond = 3.44
+    elif exp == 4 and cond is None: cond = 10
 
     with open(makef, 'r') as f:
         for line in f:
@@ -34,43 +40,54 @@ def genfig(exp, makef, outf, cond):
             iterations = int(iterations)
 
             if exp == 1:
-                if cond is not None and delay == cond:
-                    bench_res[(engine, fs)][0].append(iterations)
-                elif cond is None and delay == 3.44:
-                    bench_res[(engine, fs)][0].append(iterations)
+                if delay != cond: continue
+                bench_res[(engine, fs)].append((iterations,
+                                                float(bench.split(' ')[2])))
             elif exp == 2:
-                bench_res[(engine, fs)][0].append(int(chunks))
+                if delay not in exp2_delays: continue
+                bench_res[(engine, fs)].append((chunks,
+                                                float(bench.split(' ')[2])))
             elif exp == 3:
-                bench_res[(engine, fs)][0].append(im)
+                if delay != exp3_delays: continue
+                bench_res[(engine, fs)].append((im,
+                                                float(bench.split(' ')[2])))
             else:
-                if cond is not None and iterations == cond:
-                    bench_res[(engine, fs)][0].append(delay)
-                elif cond is None and iterations == 10:
-                    bench_res[(engine, fs)][0].append(delay)
[…]

### 2018-11-28 pipelines/spark_inc.py
fixed pycodestyle

@@ -230,7 +230,10 @@ def main():
         files = glob(os.path.join(args.bb_dir, '*'))
         fidx = [i for i in range(0, len(files))]
         imRDD = sc.parallelize(zip(fidx, files), len(files))
-        if args.work_dir is None: args.work_dir = output_dir
+
+        if args.work_dir is None:
+            args.work_dir = output_dir
+
         work_dir = os.path.abspath(os.path.join(args.work_dir,
                                                 'app-{}'.format(app_uuid)))
         print(work_dir)

### 2018-11-28 results/code/d_c_graph.py
Fixed import and local plot

@@ -1,6 +1,5 @@
 #!/usr/bin/env python
 
-import utils
 import sys  
 import os
 import argparse
@@ -35,7 +34,7 @@ def d_c_fig(bench_dir, makespan_file, out_file):
     
     bandwidths = { # MiB/s
         'lustre': 401.49,
-        'local': 1480.2, 
+        'local': 292, 
         'tmpfs': 1793.75
     }
     
@@ -135,6 +134,7 @@ def d_c_fig(bench_dir, makespan_file, out_file):
     plt.savefig(os.path.join(os.path.dirname(out_file), 
                 'tmpfs-{}'.format(os.path.basename(out_file))))
     
+    plt.figure()
     plt.plot(x_disk, y_disk, 'b+', label="Local Disk")
     rect = plt.Rectangle([1, 0], 150, 1, color='gray', edgecolor=None)
     plt.gca().add_patch(rect)

### 2018-11-30 pipelines/nipype_kmeans.py
fixed kmean functions

@@ -6,6 +6,7 @@ from shutil import rmtree
 import argparse
 import os
 import glob
+import uuid
 
 
 
@@ -21,6 +22,7 @@ def get_nearest_centroid(img, centroids):
     for vox in data:
         distance = None
         nearest_c = None
+        nearest_cv = None
 
         for c in centroids:
             c_dist = abs(vox - c[1])
@@ -28,6 +30,16 @@ def get_nearest_centroid(img, centroids):
             if distance is None or c_dist < distance:
                 distance = c_dist
                 nearest_c = c[0]
+                nearest_cv = c[1]
+            '''elif c_dist == distance:
+                if vox % 2 == 1:
+                    if nearest_cv < c[1]:
+                        nearest_c = c[0]
+                        nearest_cv = c[1]
+                else:
+                    if nearest_cv > c[1]:
+                        nearest_c = c[0]
+                        nearest_cv = c[1]'''
 
         if nearest_c not in assignments:
             assignments[nearest_c] = [vox]
@@ -80,23 +92,28 @@ def classify_chunks(img, assignments):
     import nibabel as nib
     import pickle
     from os import path as op
+    from numpy import where, isin
 
+    # assume all assignment files fit in memory
     a_files = [t for l in assignments for t in l]
-
     i = nib.load(img)
     data = i.get_fdata()
     shape = i.shape
 
-    for z in range(0, shape[2]):
-        for y in range(0, shape[1]):
-            for x in range(0, shape[0]):
-                for t in a_files:
-                    with open(t[1], 'rb') as f:
-                        elements = set(pickle.load(f))
+    assignments = {}
 
-                        if data[x][y][z] in elements:
-                            data[x][y][z] = t[0]
-                            break
+    for t in a_files:
+        if t[0] not in assignments:
+            with open(t[1], 'rb') as f:
+                assignments[t[0]] = set(pickle.load(f))
+        else:
+            with open(t[1], 'rb') as f:
+                assignments[t[0]].update(pickle.load(f))
+
+
+    for k in assignments:
+        assigned_voxels = list(assignments[k])
+        data[where(isin(data, assigned_voxels))] = k
 
     i_out = nib.Nifti1Image(data, i.affine, i.header)
     i_name = op.basename(img)
@@ -121,6 +138,8 @@ def main():
     parser.add_argument('bb_dir', type=str, help='The folder containing '
[…]

### 2018-11-30 pipelines/spark_kmeans.py

@@ -18,7 +18,10 @@ def get_nearest_centroid(d, c):
     for centroid in c:
         c_dist = abs(d-centroid)
 
-        if distance is None or c_dist < distance:
+        if (distance is None or c_dist < distance
+                or (c_dist == distance
+                    and ((d % 2 == 1 and nearest_c < centroid)
+                         or nearest_c > centroid))):
             distance = c_dist
             nearest_c = centroid
 
@@ -27,7 +30,7 @@ def get_nearest_centroid(d, c):
 
 def update_centroids(d):
 
-    updated = sum(d)/len(d)
+    updated = float(sum(d))/len(d)
 
     return updated
 
@@ -66,6 +69,7 @@ def save_segmented(d, assignments, out):
 
 
 def main():
+    # mri centroids: 0.0, 125.8, 251.6, 377.4
     conf = SparkConf().setAppName("Spark kmeans")
     sc = SparkContext.getOrCreate(conf=conf)
 
@@ -76,17 +80,14 @@ def main():
                         " fs only)")
     parser.add_argument('iters', type=int, help="maximum number of kmean "
                                                 "iterations")
+    parser.add_argument('centroids', type=float, nargs='+',
+                        help="cluster centroids")
     parser.add_argument('output_dir', type=str, help="the folder to save "
                                                      "segmented images to "
                                                      "(local fs only)")
     args = parser.parse_args()
 
-    # fixed centroids to be able to compare with nipype implementation
-    centroids = [50314.747730447438, 5.4095965052460562, 29218.970083958127,
-                 60767.571375735897]
-    # [60741.945981249322, 28998.276891703455, 5.3324138454658172,
-    #   49808.209711495881]
-    # [0, 26214, 45874.5, 65535]
+    centroids = args.centroids
 
     output_dir = op.abspath(args.output_dir)
 
@@ -130,7 +131,7 @@ def main():
                                                  output_dir)
                         ).collect()
 
-    print("***FINAL CENTROIDS***:", centroids)
+    print("***FINAL CENTROIDS***:", count, centroids)
     print(results)
 
 

### 2018-11-30 pipelines/nipype_kmeans.py
minor fix to kmeans functions

@@ -27,19 +27,13 @@ def get_nearest_centroid(img, centroids):
         for c in centroids:
             c_dist = abs(vox - c[1])
 
-            if distance is None or c_dist < distance:
+            if (distance is None or c_dist < distance
+                    or (c_dist == distance
+                        and ((vox % 2 == 1 and nearest_cv < c[1])
+                             or (vox % 2 == 0 and nearest_cv > c[1]))):
                 distance = c_dist
                 nearest_c = c[0]
                 nearest_cv = c[1]
-            '''elif c_dist == distance:
-                if vox % 2 == 1:
-                    if nearest_cv < c[1]:
-                        nearest_c = c[0]
-                        nearest_cv = c[1]
-                else:
-                    if nearest_cv > c[1]:
-                        nearest_c = c[0]
-                        nearest_cv = c[1]'''
 
         if nearest_c not in assignments:
             assignments[nearest_c] = [vox]

### 2018-11-30 pipelines/spark_kmeans.py

@@ -21,7 +21,7 @@ def get_nearest_centroid(d, c):
         if (distance is None or c_dist < distance
                 or (c_dist == distance
                     and ((d % 2 == 1 and nearest_c < centroid)
-                         or nearest_c > centroid))):
+                         or (d % 2 == 0 and nearest_c > centroid)))):
             distance = c_dist
             nearest_c = centroid
 

### 2018-12-01 pipelines/nipype_kmeans.py
fixe nipype kmeans for slurm

@@ -7,7 +7,7 @@ import argparse
 import os
 import glob
 import uuid
-
+import json
 
 
 def get_nearest_centroid(img, centroids):
@@ -30,7 +30,7 @@ def get_nearest_centroid(img, centroids):
             if (distance is None or c_dist < distance
                     or (c_dist == distance
                         and ((vox % 2 == 1 and nearest_cv < c[1])
-                             or (vox % 2 == 0 and nearest_cv > c[1]))):
+                             or (vox % 2 == 0 and nearest_cv > c[1])))):
                 distance = c_dist
                 nearest_c = c[0]
                 nearest_cv = c[1]
@@ -61,6 +61,70 @@ def get_nearest_centroid(img, centroids):
 
     return outfiles
 
+def nearest_centroid_wf(partition, centroids, work_dir):
+    from nipype import Workflow, MapNode, Function
+    import nipype_kmeans as nk
+    import uuid
+
+    wf = Workflow('km_bb{}'.format(uuid.uuid1()))
+    wf.base_dir = work_dir
+
+    gc_nname = 'gc'
+    gc = MapNode(Function(input_names=['img', 'centroids'],
+                          output_names=['assignment_files'],
+                          function=nk.get_nearest_centroid),
+                 name=gc_nname,
+                 iterfield=['img'])
+
+    gc.inputs.img = partition
+    gc.inputs.centroids = centroids
+
+    wf.add_nodes([gc])
+    wf_out = wf.run('MultiProc')
+
+    node_names = [i.name for i in wf_out.nodes()]
+    result_dict = dict(zip(node_names, wf_out.nodes()))
+
+    assignments = (result_dict[gc_nname].result
+                                        .outputs
+                                        .assignment_files)
+
+    assignments = [t for l in assignments for t in l]
+    return assignments
+
+def save_classified_wf(partition, assignments, work_dir, output_dir,
+                       iteration):
+
+    from nipype import Workflow, Node, Function
+    import nipype_kmeans as nk
+    
+    res_wf = Workflow('km_classify')
+    res_wf.base_dir = work_dir
+    c_idx = 0
+    for chunk in partition:
+        cc = Node(Function(input_names=['img', 'assignments'],
+                           output_names=['out_file'],
+                           function=nk.classify_chunks),
+                  name='{0}cc_{1}'.format(iteration, c_idx))
+
+        cc.inputs.img = chunk
+        cc.inputs.assignments = assignments 
+        res_wf.add_nodes([cc])
+
[…]

### 2018-12-01 pipelines/nipype_kmeans.py
fixed bugs in kmeans

@@ -15,7 +15,7 @@ def get_nearest_centroid(img, centroids):
     import pickle
     from os import path as op
 
-    data = nib.load(img).get_fdata().flatten()
+    data = nib.load(img).get_data().flatten()
 
     assignments = {}
 
@@ -156,7 +156,7 @@ def classify_chunks(img, assignments):
     # assume all assignment files fit in memory
     a_files = [t for l in assignments for t in l]
     i = nib.load(img)
-    data = i.get_fdata()
+    data = i.get_data()
     shape = i.shape
 
     assignments = {}
@@ -169,12 +169,11 @@ def classify_chunks(img, assignments):
             with open(t[1], 'rb') as f:
                 assignments[t[0]].update(pickle.load(f))
 
-
     for k in assignments:
         assigned_voxels = list(assignments[k])
         data[where(isin(data, assigned_voxels))] = k
 
-    i_out = nib.Nifti1Image(data, i.affine, i.header)
+    i_out = nib.Nifti1Image(data, i.affine)
     i_name = op.basename(img)
     nib.save(i_out, i_name)
 
@@ -185,7 +184,7 @@ def save_classified(img, output_dir):
     import shutil
     from os import path as op
 
-    out_name = 'classified_{}'.format(op.basename(img))
+    out_name = 'classified-{}'.format(op.basename(img))
 
     out_file = op.join(output_dir, out_name)
 

### 2018-12-01 pipelines/spark_kmeans.py

@@ -47,6 +47,7 @@ def get_voxels(d):
 
 
 def save_segmented(d, assignments, out):
+
     # read data into nibabel
     fh = nib.FileHolder(fileobj=BytesIO(d[1]))
     im = nib.Nifti1Image.from_file_map({'header': fh, 'image': fh})
@@ -62,7 +63,7 @@ def save_segmented(d, assignments, out):
     im_seg = nib.Nifti1Image(data, im.affine)
 
     # save segmented image
-    output_file = op.join(out, 'seg-' + op.basename(d[0]))
+    output_file = op.join(out, 'classified-' + op.basename(d[0]))
     nib.save(im_seg, output_file)
 
     return (output_file, "SAVED")
@@ -103,13 +104,13 @@ def main():
     voxelRDD = imRDD.flatMap(get_voxels).cache()
 
     c_changed = True
-    count = 1
+    count = 0
     assignments = None
 
-    while c_changed or count > args.iters:
+    while c_changed or count < args.iters:
         assignments = voxelRDD.map(lambda x: get_nearest_centroid(x,
                                                                   centroids)) \
-                              .groupByKey()
+                              .groupByKey().sortByKey()
 
         updated_centroids = sc.parallelize(centroids) \
                               .zipWithIndex() \
@@ -119,7 +120,7 @@ def main():
 
         c_changed = not bool(set(centroids).intersection(updated_centroids))
 
-        centroids = updated_centroids
+        centroids = sorted(updated_centroids)
 
         if c_changed:
             print("it", count, centroids)

### 2018-12-03 pipelines/nipype_kmeans.py
minor fixes; made spark implementation more efficient

@@ -3,11 +3,16 @@ from numpy import iinfo
 from random import seed, sample
 from time import time
 from shutil import rmtree
+from socket import gethostname
 import argparse
 import os
 import glob
 import uuid
-import json
+from benchmark import write_bech
+try:
+    from threading import get_ident
+except Exception as e:
+    from thread import get_ident
 
 
 def get_nearest_centroid(img, centroids):
@@ -61,10 +66,19 @@ def get_nearest_centroid(img, centroids):
 
     return outfiles
 
-def nearest_centroid_wf(partition, centroids, work_dir):
+def nearest_centroid_wf(partition, centroids, work_dir, benchmark_dir=None):
     from nipype import Workflow, MapNode, Function
     import nipype_kmeans as nk
     import uuid
+    from time import time
+    from benchmark import write_bench
+    from socket import gethostname
+    try:
+        from threading import get_ident
+    except Exception as e:
+        from thread import get_ident
+
+    start = time()
 
     wf = Workflow('km_bb{}'.format(uuid.uuid1()))
     wf.base_dir = work_dir
@@ -90,14 +104,29 @@ def nearest_centroid_wf(partition, centroids, work_dir):
                                         .assignment_files)
 
     assignments = [t for l in assignments for t in l]
+
+    end = time()
+
+    if benchmark_dir is not None:
+        bench.write_bench('get_nearest_centroid', start, end, gethostname(),          
+                          'partition', get_ident(), benchmark_dir)
     return assignments
 
 def save_classified_wf(partition, assignments, work_dir, output_dir,
-                       iteration):
+                       iteration, benchmark_dir=None):
 
     from nipype import Workflow, Node, Function
     import nipype_kmeans as nk
-    
+    from time import time
+    from benchmark import write_bench
+    from socket import gethostname
+    try:
+        from threading import get_ident
+    except Exception as e:
+        from thread import get_ident
+   
+    start = time()
+
     res_wf = Workflow('km_classify')
     res_wf.base_dir = work_dir
     c_idx = 0
@@ -124,11 +153,24 @@ def save_classified_wf(partition, assignments, work_dir, output_dir,
 
     res_wf.run(plugin='MultiProc')
 
+    end = time()
+    if benchmark_dir is not None:
+        write_bench('save_classified', start, os.gethostname(),          
+                    'partition', get_ident(), benchmark_dir)
+
     return ('Success', partition)
 
-def update_centroids(centroid, assignments):
[…]

### 2018-12-03 pipelines/spark_kmeans.py

@@ -2,13 +2,18 @@
 
 from pyspark import SparkContext, SparkConf
 from io import BytesIO
-from os import path as op
-from os import makedirs
+from os import path as op, makedirs
+from time import time
 import argparse
 import sys
 import nibabel as nib
 import numpy as np
-
+from socket import gethostname
+import benchmark as bench
+try:
+    from threading import get_ident
+except Exception as e:
+    from thread import get_ident
 
 def get_nearest_centroid(d, c):
 
@@ -16,21 +21,24 @@ def get_nearest_centroid(d, c):
     nearest_c = None
 
     for centroid in c:
-        c_dist = abs(d-centroid)
+        c_dist = abs(d[0]-centroid)
 
         if (distance is None or c_dist < distance
                 or (c_dist == distance
-                    and ((d % 2 == 1 and nearest_c < centroid)
-                         or (d % 2 == 0 and nearest_c > centroid)))):
+                    and ((d[0] % 2 == 1 and nearest_c < centroid)
+                         or (d[0] % 2 == 0 and nearest_c > centroid)))):
             distance = c_dist
             nearest_c = centroid
 
-    return (nearest_c, d)
+    return (nearest_c, (d[0], d[1]))
 
 
 def update_centroids(d):
-
-    updated = float(sum(d))/len(d)
+    # format = list of tuples
+    # e.g. [(0, 1), (2, 4), (3, 2)]
+    sum_els = float(sum([i[0]*i[1] for i in d]))
+    num_els = sum(i[1] for i in d)
+    updated = sum_els/num_els
 
     return updated
 
@@ -55,11 +63,10 @@ def save_segmented(d, assignments, out):
     data = im.get_data()
 
     assigned_class = [c[0] for c in assignments]
-
-    for i in range(0, len(assignments)):
-        assigned_voxels = list(set(assignments[i][1]))
+    
+    for i in range(0, len(assigned_class)):
+        assigned_voxels = [l[0] for c in assignments if c[0] == assigned_class[i] for l in c[1]]#list(set(assignments[i][1]))
         data[np.where(np.isin(data, assigned_voxels))] = assigned_class[i]
-
     im_seg = nib.Nifti1Image(data, im.affine)
 
     # save segmented image
@@ -71,6 +78,8 @@ def save_segmented(d, assignments, out):
 
 def main():
     # mri centroids: 0.0, 125.8, 251.6, 377.4
+    start = time()
+
     conf = SparkConf().setAppName("Spark kmeans")
[…]

### 2018-12-03 paper.tex
added spark section

@@
 
-Big Data processing engines leverage data locality, 
-in-memory computing, and lazy evaluation to reduce the impact of data 
-manipulation on the performance of applications. \todo{We need to 
-define data locality, in-memory computing and lazy evaluation somewhere 
-in the intro.} From early MapReduce 
-implementations~\cite{dean2008mapreduce} to Apache 
-Spark~\cite{zaharia2016apache} and now Dask~\cite{rocklin2015dask}, 
-these frameworks have become 
-mainstream tools for data analytics. Meanwhile, several scientific 
+Big Data processing engines specialize in minimizing data transfers, through 
+use of concepts such as data locality, in-memory computing and lazy evaluation,
+to speed up the processing time of such large datasets.
+Data locality was popularized by the 
+MapReduce~\cite{dean2008mapreduce} framework. Rather than performing costly 
+data transfers across the network, tasks were scheduled to the nodes where data
+was located. Apache Spark~\cite{zaharia2016apache} succeeded MapReduce with its
+incorporation of in-memory computing and lazy-evaluation which brought 
+significant speed-ups in Big Data applications. In-memory computing
+enabled data to be cached in memory whenever possible, reducing more costly
+disk I/Os between pipeline steps. Introducing lazy
+evaluation meant that results were only
+computed only when necessary, enabling further performace optimizations. 
+Frameworks such as
+MapReduce and Spark have since become mainstream tools for data analytics, 
+although many others, such as language-specific 
+Dask~\cite{rocklin2015dask}, are emerging. 
+Meanwhile, several scientific 
 domains including bioinformatics, physics or astronomy, have entered 
@@
 scientific processing engines such as Pegasus~\cite{deelman2005pegasus}, 
-Taverna~\cite{oinn2004taverna}, and many others.
+Taverna~\cite{oinn2004taverna} and the necessary adaptations that would need to
+be made to Big Data processing engines for scientific computing.
 
-Scientific applications slightly differ from the typical Big Data use 
+Scientific applications differ from the typical Big Data use 
 cases, which might explain the remaining gap between Big Data and 
 scientific engines. While Big Data applications mostly target text 
-processing (Web search, frequent pattern mining, recommender systems, 
-etc~\cite{leskovec2014mining}), implemented in consistent software libraries, and 
-running on clouds or dedicated commodity clusters with locality-aware 
-file systems such as HDFS~\cite{shvachko2010hadoop}, scientific applications often 
-involve binary data such as images and signals, processed by 
+processing (e.g. Web search, frequent pattern mining, recommender 
+systems~\cite{leskovec2014mining}), implemented in consistent software 
+libraries and running on clouds or dedicated commodity clusters with locality-aware 
+file systems such as HDFS~\cite{shvachko2010hadoop}, scientific applications 
+often involve binary data such as images and signals, processed by 
 command-line tools using a mix of programming languages (C, Fortran, 
-Python, shell scripts) and deployed on large, shared clusters where 
-data is fetched from data nodes to compute nodes using shared file 
+Python, shell scripts) and are deployed on large, shared clusters where 
+data is fetched from data nodes to compute nodes using centralized shared file 
 systems such as Lustre~\cite{schwan2003lustre}. Such differences in 
@@
 mention only one, in-memory computing requires instrumentation to be 
-applied to command-line tools. \todo{Also mention that performance is 
-only one of the concerns of scientific processing engines, which might
-explain why Big Data engines are overlooked.}
+applied to command-line tools. Moreover, performance is only one of the important
+aspects to scientific computing. Other important aspects include reproducibility 
+and provenance tracking. While these are minimally addressed by Big Data
+processing engines, it is currently insufficient for scientific computing. 
+Incorporation of provenance techniques into frameworks such are Spark are
+currently being investigated~\cite{samba}.
 
@@
 better address scientific code integration, data partitioning, data 
-formats and system tuning. The work in~\cite{samba} also extended Spark 
-with a better provenance system for scientific applications. We argue 
+formats and system tuning. We argue 
 that such efforts should only be conducted if substantial performance 
@@
 \subsubsection{Apache Spark}
+
+Apache Spark is a popular Scala-based processing framework for Big Data, with
+APIs in Java, Python and R. Its 
+generalized nature allows Spark to not only be applied to batch workflows,
+but also SQL queries, iterative machine learning applications, 
+data streaming and graph processing. Spark's
+main features include data-locality, in-memory processing and lazy evaluation,
+which it achieves through its principle abstraction, the Resilient Distributed 
+Dataset (RDD).
+
+An RDD is an immutable parallel data structure that achieve fault-tolerance 
+through the concept of lineage~\cite{RDDpaper}. Rather than permitting
+fine-grained transformations, only coarse-grained transformations (applying to
+many elements in the RDD), can be applied, thereby making it simple to keep a 
+log of how the data was modified. This log, known as the lineage, is used
+to reproduce any lost data.
+
+Two types of operations can be performed on RDDs:
+transformations and actions. Applying a transformation to an RDD produces a new,
+child, RDD through a narrow or wide dependency. A narrow dependency signifies 
+that the child is only dependent on a single parent partition, whereas a child 
+RDD is dependent on all parent partitions in a wide dependency. Examples of 
+spark transformations include map, filter and join. To materialize an RDD, an
+action must be performed, such as a reduce or a collect.
+
+All actions and wide dependencies require a shuffle -- Spark's most costly
+operation. Every shuffle begins with each map task saving its data to local
+files to be fetched by the shuffle operation. The shuffle operation then 
+redistributes the data across the partitions as requested. A shuffle marks a 
+stage boundary in Spark. In other words, a shuffle will not begin until all 
+dependent map tasks have completed.
+
+Spark will spill RDD data to disk whenever the complete dataset does not fit in
+memory. Moreover, as Spark transformations generate new RDDs and numerous 
+transformations may occur in a single application, Spark implements a LRU eviction
+policy. If an evicted RDD needs to be reused, Spark will recomputed it using the
+lineage data collected. It therefore becomes of great importance to persist to
+memory and/or 
+disk RDDs which will be reused throughout the application, particularly if the
+recomputation is costly. 
+%mention results in thunder paper
+
+Data locality in Spark is achieved through the  scheduling task to partitions which have the
+data loaded in memory. If the data is instead stored on HDFS, the scheduler will
+assign it to one of the preferred locations specified by HDFS. Spark's scheduler
+utilizes delay scheduling to ensure optimal fairness and locality for all tasks.
+%improve paragraph
+
+Three different types of schedulers are compatible with Spark: 1) Spark 
+Standalone, 2) YARN~\cite{yarn} and 3) Mesos~\cite{mesos}. The Spark Standalone
+scheduler is the default scheduler. YARN was a scheduler designed for Hadoop~\cite{hadoop}
+clusters. It is prepackaged with Hadoop installations. In contrast to YARN, 
+Mesos was designed to be used in multitenant cluster environments. In our experiments,
+we focus on the Spark Standalone cluster.
+
+Executing Spark applications on HPC systems with Spark-unaware schedulers may
+be inefficient. The amount of resources requested by Spark may impede Spark-cluster
+scheduling time. Using pilot-scheduling strategies to add nodes to the Spark cluster
+as they are allocated by the underlying schedulers may speedup allocation and 
+overall processing time. This, however, is not studied in the current paper.
+
+As Spark is frequently used by the scientific community, we designed our 
+experiments using the PySpark API. This is at a cost to performance as the PySpark
+code must undergo Python to Java serialization.
 

### 2018-12-03 paper.tex
some text on nipype

@@
 \subsubsection{Nipype}
+
+Nipype is a popular Python neuroimaging processing engine. It aims at being a 
+solution for easily creating reproducible neuroimaging workflows. Although Nipype
+does not employ any Big Data processing strategies, it provides access to numerous
+plugins to schedulers found in most clusters readily available to researchers, 
+such as SGE, PBS, SLURM and HTCondor. It also includes its own scheduler, MultiProc,
+for parallel processing on single nodes. Nipype also includes many built-in 
+interfaces to commonly used neuroimaging tools to be incorporated within the 
+workflows. Nipype's ability to easily parallelize workflows in researcher-available 
+cluster setups, capture detailed provenance information necessary for reproducibility
+and allow users to easily integrate existing neuroimaging tools, make it preferable
+to existing Big Data solutions, which would necessitate modification to achieve this.
+
+Jobs, or Interfaces, in Nipype, are encapsulated by a Node. A Node dictates that
+the job will execute on a single input. However, there are exists the MapNode, 
+a child variant of the Node, which can execute on multiple inputs. All tasks in
+Nipype execute in their own uniquely named subdirectory which enables the provenance 
+tracking of inputs and outputs. This also enables checkpointing of the workflow.
+In the case of Node failure or application modification, only the nodes who have
+been modified (verified by hash) are re-executed.
+
+In order for Nipype to operate as intended, a filesystem shared by all the nodes 
+is required. However, it is still possible to save to a non-shared local filesystem, 
+but it may come at the expense of fault-tolerance. Moreover, the user will need to
+ensure that the files are appropriately directed to the nodes that require them as
+there is no guarantee of data locality in Nipype.
 

### 2018-12-03 paper.tex

@@
 
-Big Data processing engines specialize in minimizing data transfers, through 
-use of concepts such as data locality, in-memory computing and lazy evaluation,
-to speed up the processing time of such large datasets.
-Data locality was popularized by the 
-MapReduce~\cite{dean2008mapreduce} framework. Rather than performing costly 
-data transfers across the network, tasks were scheduled to the nodes where data
-was located. Apache Spark~\cite{zaharia2016apache} succeeded MapReduce with its
-incorporation of in-memory computing and lazy-evaluation which brought 
-significant speed-ups in Big Data applications. In-memory computing
-enabled data to be cached in memory whenever possible, reducing more costly
-disk I/Os between pipeline steps. Introducing lazy
-evaluation meant that results were only
-computed only when necessary, enabling further performace optimizations. 
-Frameworks such as
-MapReduce and Spark have since become mainstream tools for data analytics, 
-although many others, such as language-specific 
+Big Data processing engines reduce data-related overheads, through 
+use of concepts such as data locality, in-memory computing, and lazy evaluation.
+Data locality-aware scheduling, popularized by the MapReduce~\cite{dean2008mapreduce} 
+framework, runs tasks to the nodes holding the data rather than 
+performing costly data transfers across the network. In-memory 
+computing, adopted by MapReduce's successor Apache 
+Spark~\cite{zaharia2016apache}, caches data in memory whenever 
+possible, reducing costly disk I/Os between pipeline steps. 
+Lazy evaluation, also used in Spark, means that results are computed only 
+when necessary, enabling further performance optimizations. Frameworks 
+such as
+MapReduce and Spark have become mainstream tools for data analytics, 
+although many others, such as 
 Dask~\cite{rocklin2015dask}, are emerging. 
@@
 remains limited, perhaps due to the widespread availability of 
-scientific processing engines such as Pegasus~\cite{deelman2005pegasus}, 
-Taverna~\cite{oinn2004taverna} and the necessary adaptations that would need to
-be made to Big Data processing engines for scientific computing.
+scientific processing engines such as Pegasus~\cite{deelman2005pegasus} or
+Taverna~\cite{oinn2004taverna}, and the adaptations required in Big 
+Data processing engines for scientific computing.
 
@@
 
-\newpage % Remove this when abstract is written
-
 \section{Materials and Methods} % 4 pages
@@
 An RDD is an immutable parallel data structure that achieve fault-tolerance 
-through the concept of lineage~\cite{RDDpaper}. Rather than permitting
+through the concept of lineage~\cite{zaharia2010spark}. Rather than permitting
 fine-grained transformations, only coarse-grained transformations (applying to
@@
 RDD is dependent on all parent partitions in a wide dependency. Examples of 
-spark transformations include map, filter and join. To materialize an RDD, an
+Spark transformations include map, filter and join. To materialize an RDD, an
 action must be performed, such as a reduce or a collect.
@@
 stage boundary in Spark. In other words, a shuffle will not begin until all 
-dependent map tasks have completed.
+dependent map tasks \todo{shouldn't it be 'actions'?} have completed.
 
@@
 transformations may occur in a single application, Spark implements a LRU eviction
-policy. If an evicted RDD needs to be reused, Spark will recomputed it using the
+policy. If an evicted RDD needs to be reused, Spark will recompute it using the
 lineage data collected. It therefore becomes of great importance to persist to
@@
 
-Data locality in Spark is achieved through the  scheduling task to partitions which have the
-data loaded in memory. If the data is instead stored on HDFS, the scheduler will
+Data locality in Spark is achieved through the  scheduling of task to partitions which have the
+data loaded in memory. If the data is instead stored on HDFS\todo{expand acronym if not done in intro}, the scheduler will
 assign it to one of the preferred locations specified by HDFS. Spark's scheduler
-utilizes delay scheduling to ensure optimal fairness and locality for all tasks.
+utilizes delay scheduling to optimize fairness and locality for all tasks.
 %improve paragraph
@@
 Three different types of schedulers are compatible with Spark: 1) Spark 
-Standalone, 2) YARN~\cite{yarn} and 3) Mesos~\cite{mesos}. The Spark Standalone
-scheduler is the default scheduler. YARN was a scheduler designed for Hadoop~\cite{hadoop}
+Standalone, 2) YARN~\cite{vavilapalli2013apache} and 3) Mesos~\cite{hindman2011mesos}. The Spark Standalone
+scheduler is the default scheduler. YARN was a scheduler designed for Hadoop~\cite{white2012hadoop}
 clusters. It is prepackaged with Hadoop installations. In contrast to YARN, 
 Mesos was designed to be used in multitenant cluster environments. In our experiments,
-we focus on the Spark Standalone cluster.
+we focus on the Spark Standalone cluster. \todo{Maybe defer this to experiments}
 
@@
 as they are allocated by the underlying schedulers may speedup allocation and 
-overall processing time. This, however, is not studied in the current paper.
+overall processing time. This, however, is not studied in the current paper. \todo{Explain why}
 
@@
 plugins to schedulers found in most clusters readily available to researchers, 
-such as SGE, PBS, SLURM and HTCondor. It also includes its own scheduler, MultiProc,
+such as SGE, PBS, SLURM and HTCondor\todo{expand acronyms}. It also includes its own scheduler, MultiProc,
 for parallel processing on single nodes. Nipype also includes many built-in 
@@
 workflows. Nipype's ability to easily parallelize workflows in researcher-available 
-cluster setups, capture detailed provenance information necessary for reproducibility
-and allow users to easily integrate existing neuroimaging tools, make it preferable
+cluster setups, capture detailed provenance information necessary for reproducibility,
+and allow users to easily integrate existing neuroimaging tools, make it preferable\todo{not the right word and sentence is difficult to parse.}
 to existing Big Data solutions, which would necessitate modification to achieve this.
@@
 Jobs, or Interfaces, in Nipype, are encapsulated by a Node. A Node dictates that
-the job will execute on a single input. However, there are exists the MapNode, 
-a child variant of the Node, which can execute on multiple inputs. All tasks in
-Nipype execute in their own uniquely named subdirectory which enables the provenance 
+the job will execute on a single input. However, the MapNode, 
+a child variant of the Node, can execute on multiple inputs. All tasks in
+Nipype execute in their own uniquely named subdirectory which facilitates the provenance 
 tracking of inputs and outputs. This also enables checkpointing of the workflow.
@@
 is required. However, it is still possible to save to a non-shared local filesystem, 
-but it may come at the expense of fault-tolerance. Moreover, the user will need to
+but it may come at the expense of fault-tolerance \todo{explain why}. Moreover, the user will need to
 ensure that the files are appropriately directed to the nodes that require them as

### 2018-12-04 pipelines/nipype_kmeans.py
functional nipype kmeans and minor fixes

@@ -2,13 +2,13 @@ from nipype import Workflow, MapNode, Node, Function
 from numpy import iinfo
 from random import seed, sample
 from time import time
-from shutil import rmtree
 from socket import gethostname
+from math import ceil
 import argparse
 import os
 import glob
 import uuid
-from benchmark import write_bech
+from benchmark import write_bench
 try:
     from threading import get_ident
 except Exception as e:
@@ -18,6 +18,7 @@ except Exception as e:
 def get_nearest_centroid(img, centroids):
     import nibabel as nib
     import pickle
+    from json import dump
     from os import path as op
 
     data = nib.load(img).get_data().flatten()
@@ -40,39 +41,68 @@ def get_nearest_centroid(img, centroids):
                 nearest_c = c[0]
                 nearest_cv = c[1]
 
+        vox = str(vox)
+
         if nearest_c not in assignments:
-            assignments[nearest_c] = [vox]
+            assignments[nearest_c] = { vox: 1 }
+        elif vox not in assignments[nearest_c]:
+            assignments[nearest_c][vox] = 1
         else:
-            assignments[nearest_c].append(vox)
+            assignments[nearest_c][vox] += 1
 
     outfiles = []
 
     try:
         for k, v in assignments.items():
-            out_name = 'centroid-{}.out'.format(k)
-            with open(out_name, 'ab+') as f:
-                pickle.dump(v, f)
+            out_name = 'centroid-{}.json'.format(k)
+            with open(out_name, 'a+') as f:
+                dump(v, f)
 
             outfiles.append((k, op.abspath(out_name)))
 
     except Exception as e:
         for k, v in assignments.iteritems():
-            out_name = 'centroid-{}.out'.format(k)
+            out_name = 'centroid-{}.json'.format(k)
 
-            with open('centroid-{}.out'.format(k), 'ab+') as f:
-                pickle.dump(v, f)
+            with open(out_name, 'a+') as f:
+                dump(v, f)
 
             outfiles.append((k, op.abspath(out_name)))
 
     return outfiles
 
-def nearest_centroid_wf(partition, centroids, work_dir, benchmark_dir=None):
-    from nipype import Workflow, MapNode, Function
+def reduceFilesByCentroid(centroid, assignments):
+    from json import load, dump
+    from collections import Counter
+    from os.path import basename, abspath
+
+    try:
+        a_files = [t[1] for l in assignments for t in l if t[0] == centroid[0]]
+    except Exception as e:
[…]

### 2018-12-04 pipelines/spark_kmeans.py

@@ -167,7 +167,7 @@ def main():
     print("***FINAL CENTROIDS***:", count, centroids)
     print(results)
 
-    end = time()
+    end = time() - start
 
 
     if args.benchmark:

### 2018-12-04 pipelines/nipype_kmeans.py
bug fixes

@@ -74,18 +74,20 @@ def get_nearest_centroid(img, centroids):
 def reduceFilesByCentroid(centroid, assignments):
     from json import load, dump
     from collections import Counter
-    from os.path import basename, abspath
+    from os.path import basename, abspath, isfile
 
     try:
         a_files = [t[1] for l in assignments for t in l if t[0] == centroid[0]]
+
     except Exception as e:
         a_files = [l[1] for l in assignments if l[0] == centroid[0]]
 
     c_assignments = Counter({})
     for fn in a_files:
-        with open(fn, 'r') as f:
-            partial_a = Counter(load(f))
-            c_assignments = c_assignments + partial_a
+        if isfile(fn):
+            with open(fn, 'r') as f:
+                partial_a = Counter(load(f))
+                c_assignments = c_assignments + partial_a
 
     out_name = 'centroid-{}.json'.format(centroid[0]) 
     with open(out_name, 'a+') as f:
@@ -108,6 +110,7 @@ def nearest_centroid_wf(partition, centroids, work_dir, benchmark_dir=None,
     except Exception as e:
         from thread import get_ident
 
+    tmpfs = None
     start = time()
 
     exec_id = uuid.uuid1()
@@ -259,7 +262,6 @@ def classify_chunks(img, assignments, benchmark_dir=None):
     from numpy import where, isin
 
     # assume all assignment files fit in memory
-    a_files = [t for l in assignments for t in l]
     i = nib.load(img)
     data = i.get_data()
     shape = i.shape
@@ -267,7 +269,7 @@ def classify_chunks(img, assignments, benchmark_dir=None):
     a = {}
 
     # should only be one file per centroid
-    for t in a_files:
+    for t in assignments:
         with open(t[1], 'r') as f:
             a[t[0]] = [k for k in load(f)]
 
@@ -423,7 +425,7 @@ def main():
                            function=save_classified_wf),
                   name='scf_{}'.format(c_idx))
         cc.inputs.partition = partition
-        cc.inputs.assignments = (result_dict[gc_nname].result
+        cc.inputs.assignments = (result_dict[gr_nname].result
                                  .outputs
                                  .assignment_files)
         cc.inputs.work_dir = work_dir

### 2018-12-05 pipelines/nipype_kmeans.py
removed all map nodes and cp to outdir

@@ -110,7 +110,6 @@ def nearest_centroid_wf(partition, centroids, work_dir, benchmark_dir=None,
     except Exception as e:
         from thread import get_ident
 
-    tmpfs = None
     start = time()
 
     exec_id = uuid.uuid1()
@@ -120,24 +119,31 @@ def nearest_centroid_wf(partition, centroids, work_dir, benchmark_dir=None,
                    else work_dir)
 
     gc_nname = 'gc'
-    gc = MapNode(Function(input_names=['img', 'centroids'],
-                          output_names=['assignment_files'],
-                          function=nk.get_nearest_centroid),
-                 name=gc_nname,
-                 iterfield=['img'])
+    idx = 0
 
-    gc.inputs.img = partition
-    gc.inputs.centroids = centroids
+    for chunk in partition:
+        gc_nname_it = '{0}-{1}'.format(gc_nname, idx)
+        gc = Node(Function(input_names=['img', 'centroids'],
+                           output_names=['assignment_files'],
+                           function=nk.get_nearest_centroid),
+                  name=gc_nname_it)
 
-    wf.add_nodes([gc])
+        gc.inputs.img = chunk
+        gc.inputs.centroids = centroids
+
+        wf.add_nodes([gc])
+        idx += 1
+
     wf_out = wf.run('MultiProc')
 
     node_names = [i.name for i in wf_out.nodes()]
     result_dict = dict(zip(node_names, wf_out.nodes()))
 
-    assignments = (result_dict[gc_nname].result
-                                        .outputs
-                                        .assignment_files)
+    assignments = ([result_dict['{0}-{1}'.format(gc_nname, i)]
+                                         .result
+                                         .outputs
+                                         .assignment_files
+                    for i in range(0, len(partition))])
 
     assignments = [t for l in assignments for t in l]
 
@@ -145,24 +151,28 @@ def nearest_centroid_wf(partition, centroids, work_dir, benchmark_dir=None,
     wf.base_dir = work_dir
 
     lr_nname = 'lr'
-    lr = MapNode(Function(input_names=['centroid', 'assignments'],
-                          output_names=['assignment_files'],
-                          function=nk.reduceFilesByCentroid),
-                 name=lr_nname,
-                 iterfield=['centroid'])
-    lr.inputs.centroid = centroids
-    lr.inputs.assignments = assignments
 
-    wf.add_nodes([lr])
+    for c in centroids:
+        lr_nname_it = '{0}-{1}'.format(lr_nname, c[0])
[…]

### 2018-12-05 pipelines/nipype_kmeans.py
code cleanup

@@ -275,7 +275,7 @@ def classify_chunks(img, assignments, out_dir, benchmark_dir=None):
         data[where(isin(data, a[k]))] = k
 
     i_out = nib.Nifti1Image(data, i.affine)
-    i_name = op.join(out_dir, op.basename(img))
+    i_name = op.join(out_dir, 'classified-{}'.format(op.basename(img)))
     nib.save(i_out, i_name)
 
     return op.abspath(i_name)
@@ -331,14 +331,15 @@ def main():
 
     work_dir = os.path.join(os.getcwd(), 'np_km_work')
 
+    f_per_n = ceil(len(bb_files) / args.nodes)
+    file_partitions = [bb_files[x:x+f_per_n] for x in range(
+                                                      0,
+                                                      len(bb_files),
+                                                      f_per_n)]
+
     while c_changed and idx < args.iters:
         wf = Workflow('km_bb1_slurm_{}'.format(idx))
         wf.base_dir = work_dir
-        f_per_n = ceil(len(bb_files) / args.nodes)
-        file_partitions = [bb_files[x:x+f_per_n] for x in range(
-                                                          0,
-                                                          len(bb_files),
-                                                          f_per_n)]
 
         gc_nname = 'gc_slurm_part{}'.format(idx)
         pidx = 0
@@ -346,7 +347,7 @@ def main():
 
             gc_nname_it = '{0}-{1}'.format(gc_nname, pidx)
             gc = Node(Function(input_names=['partition', 'centroids',
-                                               'work_dir', 'benchmark_dir'],
+                                            'work_dir', 'benchmark_dir'],
                                   output_names=['assignment_files'],
                                   function=nearest_centroid_wf),
                          name=gc_nname_it)
@@ -380,7 +381,7 @@ def main():
         wf = Workflow('km_bb2_slurm_{}'.format(idx))
         wf.base_dir = work_dir
         for c in centroids:
-            gr_nname_it = '{0}-{1}'.format(gr_nname, int(c[0]))
+            gr_nname_it = '{0}-{1}'.format(gr_nname, c[0])
             gr = Node(Function(input_names=['centroid', 'assignments'],
                                output_names=['assignment_files'],
                                function=reduceFilesByCentroid),
@@ -390,7 +391,7 @@ def main():
 
             wf.add_nodes([gr])
 
-            uc_nname_it = '{0}-{1}'.format(uc_nname, int(c[0])) 
[…]

### 2018-12-09 paper.tex
More methods added to paper. fixed bug in makespan extraction code

@@
 
+All experiments were executed on the Dell Zenith cluster~\cite{dellHPC} running 
+SLURM. For the 
+Spark experiments, a Spark cluster was started atop our SLURM allocation. We
+had access to 16 dedicated nodes. Each compute node ran on RHEL 7.4, was 
+equipped with 40 Intel Xeon Gold 6148 2.4GHz CPUs and had
+187GB of memory, 94GB of tmpfs and a Dell 120GB M.2 SATA SSD as local disk. 
+A 698TB Lustre server is accessible on each node through a 100Gb/s Infiniband
+network. The apparent bandwidth of the local disk was found to be 1480MB/S, 
+whereas Lustre's bandwidth was found to be at 401.49MB/S.
 % Dedicated Compute nodes with SLURM
@@
 \hline
-50~MB/s         & 5~MB/s
+164.5~MB/s         & 3.2~MB/s
 \end{tabular}
 \label{table:infrastructure}
-\caption{Bandwidth thresholds on the target cluster}
+\caption{Bandwidth thresholds on the target cluster using 15 worker nodes and a
+         maximum of 9 CPUs on each node.}
 \end{table}
@@
 % Some BIDS dataset for fmriprep
+For our experiments, BigBrain~\cite{bigbrain} a 75GB 40$\mu$m isotropic 
+histological image of a 65-year-old male's brain. The BigBrain was selected due 
+to being the only one of its kind, as there does not yet exist a 
+higher-resolution image of a human brain. Moreover, there currently exists a 
+lack of standardized tools for processing the BigBrain as a consequence 
+of its size. To examine the effects processing BigBrain has on the Big Data 
+strategies, we partitioned the full 3845X3015X3470 voxel image into 30 (5X3X2)
+chunks, 125 (5X5X5) chunks and 750 (5X15X10) chunks. Additionally, the full 
+image was also split in half (769X603X347 voxels) and the half image was 
+partitioned into 125 chunks.
+
+Processing large images are only considered to be part of the BigData problem 
+in neuroscience. The other problem is processing large MRI datasets, that is,
+datasets consisting of many small brain images belonging to various different 
+subjects. Although we have not explored explicitly the processing of large MRI
+datasets, the 75GB BigBrain is within the size ballpark~\cite{VanHorn2014} of 
+what a large MRI dataset would be today.
+
+As it is possible that both small and large data may need to be processed using
+the same pipeline \todo{explain pipelines in intro if not done already}, we 
+wanted to determine the effects of the data management strategies on small 
+data as well. For this, we selected a 12MB T1W image belonging to subject 1 of 
+the ds000001 dataset version 6 ~\cite{openneuro}. In order to be able to split 
+it into 125 chunks, it was necessary to zero-pad the image to the dimensions
+165X200X200 voxels, which subsequently increased the total image size to 13M.
 
@@
 % Incrementation (avoid cache effects in binarization)
+In order to investigate how the different strategies impact processing, we 
+selected a simple incrementation pipeline that consisted exclusively of map 
+stages. To would allow us to create experiments to tweak data locality, 
+in-memory computing and lazy evaluation easily. Incrementation was selected 
+over other applications, such as binarization, as it ensured that a new image
+was created at each step (i.e. no caching effects within the executing 
+application). Each partitioned chunk was incremented by 1, in parallel, by
+the processing engines. The incremented chunks would be either 
+maintained in-memory (Spark only) or saved to either tmpfs, local disk or 
+Lustre (Spark and Nipype). Should more than a single iteration be requested, 
+the increment chunks would be incremented again and saved to the same storage. 
+This would repeat until the number of requested iterations had elapsed. In all 
+conditions, the first input chunks would be reads from Lustre and the final 
+output chunks would be saved to Lustre. We chose to to perform our initial 
+reads and final writes on Lustre as a multi-tenant clusters' compute node's 
+local storage, if 
+available, is generally temporary and only available for the duration of the 
+allocation.
+
+For our first and second experiment\todo{better split into two paragraphs}, we wanted to determine the effects of in-memory 
+processing on increasing data size and task duration. In other words, the 125 
+chunk BigBrain 
+image was increment 1, 10 and 100 times, such that the total processed data 
+size increased from 75GB, at 1 iteration, to 7500GB at 100 iterations. In order
+to validate the model, the D/C was varied to fall in different regions of 
+Equation~\ref{eq:page-cache-inequality}. The D/C selected were 256, 178.6, 80 
+and 1.9MB/s\todo{present this as a table?}. 80 and 1.9MB/s satisfied the 
+inequality for local disk, whereas only 1.9MB/s satisfied the inequality for 
+Lustre. These D/C ratios were applied to each iteration experiment, such that 
+we were able to get both the effect of increasing the total amount of data, in 
+addition to, increasing the task duration for a given data size.
+
+As a third incrementation experiment, we were interested in the effects of 
+chunk size on in-memory computing. Naturally, greater chunk size signifies 
+a decrease in parallelization, however, it also means an increase in sequential
+I/O. For this experiment we incremented the 30, 125 and 750 chunk complete
+BigBrain image. The amount of parallelization for 30 chunks was 2 cores/node, 
+whereas it was 9 cores/node for 125 chunks and 25 cores/node for 750 chunks. 
+While Spark attempted to load-balance the data, it used up only 25 of the 40 
+cores for 750 chunks. In contrast, Nipype tried to use up as many cores as 
+possible. To ensure both engines were processing the same amount of data 
+concurrently, we fixed the amount of cores used to 25.
+Unlike the previous experiment, the D/C ratio was kept static at 174.4MB/s, 
+however, this ratio ensured that different regions of the inequality were 
+reached depending on amount of parallelization. A ratio of 174.4 MB/s satified
+the inequality for local disk at 30 chunks. All other ratios did not satisfy 
+the inequality for Lustre and local disk. Unfortunately, 30 chunks could no be 
+processed using in-memory Spark as its individual chunk size of 3GB exceeded 
+the 2GB partition size limit imposed by Spark.
+
+For our fourth and final incrementation experiment, we investigated the effects 
+of the strategies on different image sizes. We selected the 75GB BigBrain, the
+38GB half BigBrain and the 13M T1W MRI image for this experiment. The number of
+chunks was fixed at 125. Similarly to the previous experiment, the total 
+sequential compute time was fixed, however, due to varying size in total data 
+processed, the D/C ratio varied. Once again, we ensure that the D/C ratio fell
+in multiple different regions of the inequality.
 

### 2018-12-09 results/code/get_makespan.py

@@ -28,8 +28,11 @@ for bench in os.listdir(sys.argv[1]):
 
     else:
         with open(os.path.join(sys.argv[1], bench), 'r') as f:
-            driver_line = f.readline().split(" ")
-            driver_line[0] = os.path.basename(bench) + ":" + driver_line[0]
+            for line in f:
+                if 'driver_program' in line:
+                    driver_line = line.split(" ")
+                    break
+            driver_line[0] = (os.path.basename(bench) + ":" + driver_line[0])
             
 
     with open(sys.argv[2], 'a') as f:

### 2018-12-09 paper.tex
a bit more content in experiments section

@@
 \usepackage[bookmarks=false]{hyperref}
+\usepackage{algpseudocode}
+\usepackage{algorithm}
+\algnewcommand\algorithmicforeach{\textbf{for each}}
+\algdef{S}[FOR]{ForEach}[1]{\algorithmicforeach\ #1\ \algorithmicdo}
 
@@
 \subsubsection{Applications} % 1 page
+\begin{algorithm}\caption{Incrementation}\label{alg:incrementation}
+    \begin{algorithmic}
+    \Require $x$, a sleep delay in float
+    \Require $fs$, filesystem to save image to (i.e. mem, tmpfs, local disk, lustre). 
+    \ForEach{$i \in iterations$}
+    \ForEach{$chunk \in image$}
+        \State $chunk\gets chunk+1$
+        \State sleep $x$
+        \State save $chunk$ to $fs$
+    \EndFor
+    \EndFor
+\end{algorithmic}
+\end{algorithm}
 
@@
 In order to investigate how the different strategies impact processing, we 
-selected a simple incrementation pipeline that consisted exclusively of map 
-stages. To would allow us to create experiments to tweak data locality, 
-in-memory computing and lazy evaluation easily. Incrementation was selected 
+selected a simple incrementation pipeline (Algorithm~\ref{alg:incrementation}) 
+that consisted exclusively of map 
+stages. A series of map-only stages would enable us to evaluate the effects of
+in-memory computing when data locality is preserved. Incrementation was 
+selected 
 over other applications, such as binarization, as it ensured that a new image
@@
 application). Each partitioned chunk was incremented by 1, in parallel, by
-the processing engines. The incremented chunks would be either 
+the processing engines. As incrementing the images is a relatively quick 
+process, we added a sleep delay to the incrementation stage in order to 
+increase task duration. The incremented chunks would be either 
 maintained in-memory (Spark only) or saved to either tmpfs, local disk or 
@@
 
-For our first and second experiment\todo{better split into two paragraphs}, we wanted to determine the effects of in-memory 
-processing on increasing data size and task duration. In other words, the 125 
-chunk BigBrain 
-image was increment 1, 10 and 100 times, such that the total processed data 
-size increased from 75GB, at 1 iteration, to 7500GB at 100 iterations. In order
+For our first incrementation experiment, we investigated the effects of
+in-memory computing on varying total data size. to do this, we increased the 
+number of incrementation iterations from 1, 10 and 100 times. The total data 
+size would then increased from 75GB, at 1 iteration, to 7500GB, at 100 
+iterations. In order
 to validate the model, the D/C was varied to fall in different regions of 
@@
 addition to, increasing the task duration for a given data size.
+
+We combined the first experiment with our second experiment such the we could 
+evaluate the effects of in-memory computing on varying task duration. If the 
+page cache has sufficient time to flush, it should be expected that using 
+in-memory computing should have no increased speed-up over disk. We maintained 
+the D/C ratios as those mentioned above, but rather than applying only one to
+a different number of iterations, we applied all four D/C to all numbers of 
+iterations. That is 1 iteration, 10 iterations and 100 iterations were all 
+processed with using the aforementioned D/C ratios.
 
@@
 sequential compute time was fixed, however, due to varying size in total data 
-processed, the D/C ratio varied. Once again, we ensure that the D/C ratio fell
-in multiple different regions of the inequality.
+processed, the D/C ratio varied. Once again, we ensured that the D/C ratio fell
+in multiple different regions of the inequality. The D/C ratio ranged from 
+348.8 MB/s for BigBrain and 174.4MB/s for half Bigbrain, to 0.1MB/s for the MRI 
+image. Only the 0.1MB/s MRI satisfied the inequality for both Lustre and local 
+disk.
+
+Since data-locality is not normally preserved in Nipype (a new SLURM allocation 
+is requested for each processed task), we instrumented Nipype to ensure data 
+locality. That is, for each chunk partition, we requested a SLURM allocation to
+process the entire pipeline in parallel, using Nipype's MultiProc scheduler, on 
+a given node. This was possible as no communication was required between the 
+processed chunks.
 

### 2018-12-10 paper.tex
Some editing and extra text

@@
 cases, which might explain the remaining gap between Big Data and 
-scientific engines. While Big Data applications mostly target text 
+scientific engines. \todo{This sentence is extremely long} While Big Data applications mostly target text 
 processing (e.g. Web search, frequent pattern mining, recommender 
 systems~\cite{leskovec2014mining}), implemented in consistent software 
-libraries and running on clouds or dedicated commodity clusters with locality-aware 
-file systems such as HDFS~\cite{shvachko2010hadoop}, scientific applications 
-often involve binary data such as images and signals, processed by 
-command-line tools using a mix of programming languages (C, Fortran, 
-Python, shell scripts) and are deployed on large, shared clusters where 
+libraries and running on clouds or dedicated commodity clusters with 
+locality-aware file systems such as the Hadoop Distributed File System 
+(HDFS~\cite{shvachko2010hadoop}), scientific applications often involve 
+binary data such as images and signals, processed by command-line tools 
+using a mix of programming languages (C, Fortran, Python, shell 
+scripts), and are deployed on large, shared clusters where 
 data is fetched from data nodes to compute nodes using centralized shared file 
@@
 in the Linux kernel~\cite{love2010linux}, in-memory file systems 
-(\texttt{tmpfs}) and memory-mapped files\todo{check that} might also 
+(tmpfs) and memory-mapped files\todo{check that} might also 
 explain the lack of adoption of Big Data engines for scientific 
@@
 applications in mind, for instance Nipype~\cite{gorgolewski2011nipype} 
-or PSOM~\cite{bellec2012pipeline}. Big Data engines have been used for 
+or the Pipeline System for Octave and Matlab 
+(PSOM~\cite{bellec2012pipeline}). Big Data engines have been used for 
 neuroimaging applications too, including in the Thunder 
@@
 \begin{enumerate}
-\item What is the effect of in-memory computing and data locality on neuroimaging applications?
+\item What is the effect of in-memory computing, lazy evaluation and data locality on current neuroimaging applications?
 \item Can in-memory computing be effectively enabled by the operating system rather than the data processing engine?
@@
 evaluation\todo{are we talking about lazy evaluation at all?} or data 
-locality. On the other hand, scientific data processing engines are 
+locality. On the other hand, neuroimaging data processing engines are 
 still being developed, and it is legitimate to wonder whether these 
@@
 
+Infrastructure-wise, we focus on the case of High-Performance Computing 
+(HPC) clusters that are typically available through University 
+facilities or national computing infrastructures such as XSEDE, Compute 
+Canada or DEISA \todo{add links}, as this is the type of platform which 
+is typically available to neuroscientists. We assume that HPC systems 
+are multi-tenant, that compute nodes are accessible through a batch 
+scheduler, and that a file system shared among the compute nodes is 
+available, for instance Lustre. We intentionally leave distributed, 
+HDFS-like file system out, as initiatives to deploy them in HPC 
+centers, for instance Hadoop on-demand~\cite{hadoop-on-demand} have not 
+become mainstream yet.
+
 \todo{Perhaps talk about modeling and simulation here.}
@@
 \label{sec:methods}
+
+\todo{Maybe add an intro to this section, to outline it}
 
@@
 memory. Moreover, as Spark transformations generate new RDDs and numerous 
-transformations may occur in a single application, Spark implements a LRU eviction
+transformations may occur in a single application, Spark implements a Least-Recently Used (LRU) eviction
 policy. If an evicted RDD needs to be reused, Spark will recompute it using the
@@
 Data locality in Spark is achieved through the  scheduling of task to partitions which have the
-data loaded in memory. If the data is instead stored on HDFS\todo{expand acronym if not done in intro}, the scheduler will
+data loaded in memory. If the data is instead stored on HDFS, the scheduler will
 assign it to one of the preferred locations specified by HDFS. Spark's scheduler
@@
 Mesos was designed to be used in multitenant cluster environments. In our experiments,
-we focus on the Spark Standalone cluster. \todo{Maybe defer this to experiments}
+we focus on the Spark Standalone cluster.
 
@@
 as they are allocated by the underlying schedulers may speedup allocation and 
-overall processing time. This, however, is not studied in the current paper. \todo{Explain why}
+overall processing time. This, however, is not studied in the current paper.
 
@@
 cluster setups, capture detailed provenance information necessary for reproducibility,
-and allow users to easily integrate existing neuroimaging tools, make it preferable\todo{not the right word and sentence is difficult to parse.}
+and allow users to easily integrate existing neuroimaging tools, make it preferable
 to existing Big Data solutions, which would necessitate modification to achieve this.
@@
 Data storage location is critical to the performance of Big Data 
-applications on HPC clusters. Data may reside in the engine memory, on 
+applications on HPC clusters.
+Data may reside in the engine memory, on 
 a file system whose content reside in virtual memory (for instance 
-\texttt{tmpfs}), on disks local to the processing node, or on a shared 
+tmpfs), on disks local to the processing node, or on a shared 
 file system. Table~\ref{table:features} summarizes the Big Data 
-strategies that can be used depending on the data location. The 
-remainder of this Section explains this Table and provides related 
-performance models.
+strategies that can be used depending on the data location. In 
+addition, lazy evaluation is available in Spark regardless of data 
+location. The remainder of this Section explains this Table and 
+provides related performance models.
 \begin{table}
 \centering
-\begin{tabular}{c|ccc}
+\begin{tabular}{c|cc}
    \rowcolor{headcolor}
-    Data Location                 & In-Memory     & Data Locality & Lazy \\
+    Data Location                 & In-Memory     & Data Locality        \\
     \rowcolor{headcolor}
-                                  & Computing     &               & Evaluation      \\
+                                  & Computing     &                     \\
                                   \hline          
-In-memory                         &   Y           & Y             & Spark               \\
-\texttt{tmpfs}                    &   Y           & Y             & Spark           \\
-Local Disk                        & Page Caching  & Y             & Spark           \\
-Shared File System                & Page Caching  & N            & Spark          
+In-memory                         &   Y           & Y                       \\
+tmpfs                             &   Y           & Y                   \\
+Local Disk                        & Page Caching  & Y                   \\
+Shared File System                & Page Caching  & N                
 % data locality is somehow present in shared fs due to page caching again.
@@
 \subsubsection{In-Engine-Memory and In-Memory File System}
+
+ The main difference between storing the data in the engine memory, as in Spark, 
+ and simply writing to an in-memory file system, such as tmpfs, is 
+ related to what happens when the processed data fills up the available 
+ memory. In case engine memory is used, the engine has to cleanup 
+ unused data with some strategy to avoid crashes. In case an in-memory 
+ file system is used, the kernel starts swapping memory to disk, and 
+ the performance becomes that of local disk writes. In our experiments, 
+ we will explore the region  when the data consumed by the 
+ application approaches the threshold of available memory.
 
@@
 The performance of local disk accesses is strongly dependent on the 
-page caching mechanism provided by the Linux kernel, described in great 
+page caching mechanism provided by the Linux kernel, described in    
 details in~\cite{love2010linux}. To summarize, data read from disk 
@@
 kernel will return the data directly from memory if the requested data 
-lies in the page cache. Such cache hits drastically speed-up data 
+lies in the page cache, realizing a cache \emph{hit}. Cache hits drastically speed-up data 
 reads, by masking the disk latency and limited bandwidth behind a 
@@
 reads by preloading the cache. Scheduling strategies might be designed 
-to maximize cache hits though. For instance, we hypothesize that lazy 
-evaluation would result in more cache hits by scheduling data-dependent 
+to maximize cache hits though. For instance, lazy 
+evaluation could result in more cache hits by scheduling data-dependent 
 tasks on the same node.
@@
 \texttt{write()} system call, data is copied to a memory cache that is 
-asynchronously written to disk by flusher threads when memory shrinks, 
-``dirty" (unwritten) data grows, or a 
+asynchronously written to disk by flusher threads, when memory shrinks, when
+``dirty" (unwritten) data grows, or when a 
 process invokes the \texttt{sync()} system call. 
@@
 concurrent processes all consume the same CPU time, (3) data is written 
-uniformly along task execution. Due these assumptions, all the 
+uniformly along task execution. With these assumptions, all the 
 processes will write at the same rate, which explains why the model 
@@
 
-% is there a page cahe when writing to a socket? there must be!
-% anyway the model is in the form of Eq 1, but now delta is the network bandwidth
-% and gamma is the total number of concurrent processes.
-% also, is lustre going to slow down network writes?
-
-%~ In any case, I think we want the following inequality to hold:
-%~ ```
-%~ D/C <= \Delta / \Gamma
-%~ ```
-%~ where Delta is now the apparent bandwidth of the file system and Gamma 
-%~ is the max number of concurrent tasks in the cluster. Whether page 
-%~ caching is used or not will affect what happens when the inequality 
-%~ holds: if page caching is here, write time will be 0, otherwise it will 
-%~ be non-zero. If the inequality doesn't hold then we have congestion and 
-%~ it's a different game: transfer times might tend to infinity and data 
-%~ loss is to be expected.
-
-% Say that we don't use HDFS because we focus on multi-tenant clusters and HDFS on multi-tenant clusters is difficult (refer to hadoop on demand).
-
-% Make a table to explain where data locality and in-memory computing are
-% available.  As a summary of the models.
-
-% What happens when dataset and derived data is too large for disk?
-
-In~\cite{saini2012performance} the Lustre file system is benchmarked on NASA applications.
+% In~\cite{saini2012performance} the Lustre file system is benchmarked on NASA applications.
 
@@
 %  scheduling. And https://dl.acm.org/citation.cfm?id=3041715
-
-\subsection{Task Scheduling}
-
-% Explain how we ensured that both engines lead to the same scheduling.
-% Talk about pilot jobs and say we don't consider them here.
 
@@
 A 698TB Lustre server is accessible on each node through a 100Gb/s Infiniband
-network. The apparent bandwidth of the local disk was found to be 1480MB/S, 
-whereas Lustre's bandwidth was found to be at 401.49MB/S.
+network. The apparent bandwidth of the local disk was found to be 1480MB/s, 
+whereas Lustre's bandwidth was found to be at 401.49MB/s.
 % Dedicated Compute nodes with SLURM
@@
 
-\begin{table}
-\centering
-\begin{tabular}{c|c}
-$\delta/\gamma$ & $\Delta/\Gamma$\\
-\hline
-164.5~MB/s         & 3.2~MB/s
-\end{tabular}
-\label{table:infrastructure}
-\caption{Bandwidth thresholds on the target cluster using 15 worker nodes and a
-         maximum of 9 CPUs on each node.}
-\end{table}
+%~ \begin{table}
+%~ \centering
+%~ \begin{tabular}{c|c}
+%~ $\delta/\gamma$ & $\Delta/\Gamma$\\
+%~ \hline
+%~ 164.5~MB/s         & 3.2~MB/s
+%~ \end{tabular}
+%~ \label{table:infrastructure}
+%~ \caption{Bandwidth thresholds on the target cluster using 15 worker nodes and a
+         %~ maximum of 9 CPUs on each node.}
+%~ \end{table}
 
@@
 % Some BIDS dataset for fmriprep
-For our experiments, BigBrain~\cite{bigbrain} a 75GB 40$\mu$m isotropic 
+We used BigBrain~\cite{bigbrain}, a 75GB 40$\mu$m isotropic 
 histological image of a 65-year-old male's brain. The BigBrain was selected due 
@@
 of its size. To examine the effects processing BigBrain has on the Big Data 
-strategies, we partitioned the full 3845X3015X3470 voxel image into 30 (5X3X2)
-chunks, 125 (5X5X5) chunks and 750 (5X15X10) chunks. Additionally, the full 
-image was also split in half (769X603X347 voxels) and the half image was 
+strategies, we partitioned the full $3845\times3015\times3470$ voxel image into 30 ($5\times3\times2$)
+chunks, 125 ($5\times5\times5$) chunks and 750 ($5\times15\times10$) chunks. Additionally, the full 
+image was also split in half ($769\times603\times347$ voxels) and the half image was 
 partitioned into 125 chunks.
@@
 it into 125 chunks, it was necessary to zero-pad the image to the dimensions
-165X200X200 voxels, which subsequently increased the total image size to 13M.
+$165\times200\times200$ voxels, which subsequently increased the total image size to 13MB.
 
@@
 conditions, the first input chunks would be reads from Lustre and the final 
-output chunks would be saved to Lustre. We chose to to perform our initial 
+output chunks would be saved to Lustre. We chose to perform our initial 
 reads and final writes on Lustre as a multi-tenant clusters' compute node's 
@@
 
+\subsubsection{Setup}
+
 For our first incrementation experiment, we investigated the effects of
-in-memory computing on varying total data size. to do this, we increased the 
+in-memory computing on varying total data size. To do this, we increased the 
 number of incrementation iterations from 1, 10 and 100 times. The total data 
-size would then increased from 75GB, at 1 iteration, to 7500GB, at 100 
+size would then increase from 75GB, at 1 iteration, to 7500GB, at 100 
 iterations. In order
@@
 Equation~\ref{eq:page-cache-inequality}. The D/C selected were 256, 178.6, 80 
-and 1.9MB/s\todo{present this as a table?}. 80 and 1.9MB/s satisfied the 
+and 1.9MB/s\todo{present this as a table? yes, I think it would be clearer (TG) but not mandatory.}. 80 and 1.9MB/s satisfied the 
 inequality for local disk, whereas only 1.9MB/s satisfied the inequality for 
@@
 
-We combined the first experiment with our second experiment such the we could 
+We combined the first experiment with our second experiment such that we could 
 evaluate the effects of in-memory computing on varying task duration. If the 
@@
 
-\begin{table}
-\centering
-\begin{tabular}{c|c}
-Application & $\frac{D}{C}$ (GB/s)\\
-\hline
-Big Brain            & [0-75] \\
-fMRI pre-processing  & [0-1.5]
-\end{tabular}
-\label{table:applications}
-\caption{Data-compute ratios for typical neuroinformatics applications}
-\end{table}
-% fMRI: about 100M in, 2GB out,2 30 min processing time to 10 hours
-
 \subsubsection{Execution modes} % perhaps find a better title, it's not informative
@@
 specific instrumentation in Nipype.
-
-% separate this in (1) the concepts, put in the model sections below, 
-% (2) the implementation 'details', put in the experiments
-
-% in-memory (Spark only)
-
-% tmpfs: two extra in-memory copies. memfs could address that but we don't have it yet.
-% Spark: when just passing filenames. turns out i need to set `preservePartitioning` to True
-% local disk: write buffer issue
-
-% shared fs: Lustre
 
@@
 
-\subsection{Can \texttt{tmpfs} and Page Caches Emulate In-Memory Computing?}
+\subsection{Can tmpfs and Page Caches Emulate In-Memory Computing?}
 % tmpfs and local disks fill up: need cleanup.

### 2018-12-12 paper.tex
Results section and fixes to figures

--- 2018-12-09 paper.tex
@@
 \usepackage{colortbl} % for \rowcolor
-
+\usepackage{caption}
+\usepackage{subcaption}
 \usepackage[bookmarks=false]{hyperref}
@@
 
+\begin{figure*}
+\begin{subfigure}{0.33\linewidth}
+\centering
+    \includegraphics[width=\textwidth]{results/figures/tmpfs-incrementation.pdf}
+\caption{}
+\label{fig:sub1}
+\end{subfigure}%
+    \begin{subfigure}{0.33\linewidth}
+\centering
+    \includegraphics[width=\textwidth]{results/figures/local-incrementation.pdf}
+\caption{}
+\label{fig:sub2}
+    \end{subfigure}%
+\begin{subfigure}{0.33\linewidth}
+\centering
+    \includegraphics[width=\textwidth]{results/figures/lustre-incrementation.pdf}
+\caption{}
+\label{fig:sub3}
+\end{subfigure}
+    \caption{Model evaluation on (a) tmpfs, (b) local and (c) Lustre. Grey 
+             regions denote areas that violate model predictions.}
+\label{fig:modeleval}
+\end{figure*}
+
 % Incrementation (avoid cache effects in binarization)
@@
 
+In order to evaluate our model, we compared the observed in-memory speedup 
+ratio to our model's expected speed-up ratio for the given 
+D/C / $\delta$/$\gamma$ used (Figure~\ref{fig:modeleval}). Experiments for 
+which there was no in-memory equivalent (i.e. BigBrain split into 30 chunks) 
+were not considered.
+
+%probably confusing to mention values, but they're here for now
+Results show that, overall, the model correctly predicted the effect of page 
+cache on processing times for local disk and Lustre. Points which violated 
+model predictions were found at 1 iteration, where page cache would not have 
+been saturated, due to the amount of data being written, and at a task duration
+of 320 seconds. However, in all cases, the ``1" boundary was never trespassed by 
+more than a factor of 0.19, and is therefore likely a result of system 
+variability.
+
+Lustre's regression slope is smallest of the three filesystems, possibly 
+lending to the fact that Lustre was nearly always in transient state. Out of 
+the thirty four benchmarks taken on Lustre, Lustre was only expected to be in steady 
+state for four of the benchmarks (MRI and all benchmarks with a steady state of
+320). This was validated by our experiments.
+
+The next smallest slope is that of the local disk. The local disk was \todo{continue after disk benchmarks}
+
+Unlike local disk and Lustre, tmpfs exists within the page cache and should not 
+be affected by it in the same way. However, when memory is low, tmpfs will 
+begin to write to the swap space, located on disk. As a result, it is possible 
+to see similar trends to local disk when space is low. 
+Figure~\ref{fig:modeleval} (a) shows that the speed-up between in-memory 
+computing and tmpfs increases as the D/C / $\delta$/$\gamma$ ratio increases. 
+This is unexpected as both tmpfs and in-mem reside in memory and should 
+therefore both have comparable makespans. This was found to occur when the 
+contention to RAM was highest. In other words, when 750 chunks were written to
+disk in 10 iterations with a task duration of 0.59s, Spark in-memory was 
+approximately 2x faster. 
+%why is this? will generate gantts to investigate further.
+
+\subsection{Iterations}
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\linewidth]{results/figures/iterations.pdf}
+    \caption{Makespans of Spark and Nipype writing to memory, tmpfs, local 
+             disk and Lustre for a given number of iterations. All 
+             makespans use 125 chunks of the complete BigBrain and have a fixed
+             task duration of 3.44 seconds}\label{fig:iterations}
+\end{figure}
+
+Figure~\ref{fig:iterations} shows the difference between the different 
+filesystem choices given the number of iterations. At 1 iteration, all 
+filesystems behave the same, as they are all writing to the page cache. As the 
+total amount of data written by the application increases to 750~GB, there is a 
+greater disparity between Lustre and in-memory (2.67~x slower, on average). 
+Local disk performance, however, 
+is still comparable to memory (1.38~x slower, on average). Despite local disk and 
+Lustre both being in transient state, local disk encounters less contention 
+than what would be found on Lustre.
+
+%sp_mem 1it = 34s, 10it = 64 100it = 381
+%3.44 * 1 = 3.44, 3.44*10 = 34, 3.44*100 = 344
+
+%sp_tmpfs 1it = 31, 10it = 69,
+
+At 100 iterations, or 7500~GB, Lustre can be found to be, on average, 3.82x 
+slower than Spark in-memory. The page cache, on Zenith, occupies 20\% of 
+available memory. Assuming all memory (187~GB) was available, only 37.4~GB of 
+data could be held in a node's page cache at any given time. The slowdown 
+experienced can be explained by the much less data could reside in the page 
+cache at a given time, in contrast to 10 iterations, and therefore the effects 
+of disk bandwidth are more significant in this workflow
+
+While there is some variability that can be seen in Figure~\ref{fig:iterations} 
+between the two engines, this believed to be insignificant, and potentially due 
+to SLURM node allocation delays in our launching of Nipype.
+
+% Comment on the general trend
+% note 'on average' = average between spark and nipype. local disk spark is 
+% 1.22x slower and local disk nipype is 1.54x slower
+% spark lustre 2.66x slower; nipype lustre 2.68x slower
+
 \subsection{CPU Time}
@@
 %
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\linewidth]{results/figures/cputime.pdf}
+    \caption{Makespans of Spark and Nipype writing to memory, tmpfs, local 
+             disk and Lustre with varying task durations. All 
+             makespans use 125 chunks of the complete BigBrain and are fixed at 
+             10 iterations}\label{fig:cputime}
+\end{figure}
+
+Increasing CPU time ensured that all filesystems had a comparable performance
+(Figure~\ref{fig:cputime}). Lustre, for instance, is approximately 1.01x slower
+than Spark in-memory at a task duration of 320 seconds, whereas it is 
+approximately 3.25x slower that Spark in-memory with 2.4 second tasks. This 
+pattern corroborated by our model which postulates that 
+data movement costs will have little impact on compute-intensive tasks. The 
+reasoning for this is that longer tasks give the page cache more time to flush 
+between write to disk. The longer the delay, the less likely the page cache is 
+to fill up when the concurrent data size written fits within the page cache.
 
 \subsection{Image Block size}
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\linewidth]{results/figures/numchunks.pdf}
+    \caption{Makespans of Spark and Nipype writing to memory, tmpfs, local 
+             disk and Lustre with varying numbers of chunks. All 
+             makespans have the same overall sequential CPU time and 
+             are fixed at 
+             10 iterations}\label{fig:numchunks}
+\end{figure}
 % Comment on the general trend
+As can be seen in Figure~\ref{fig:numchunks}, makespan decreases when the number 
+of chunks increase. This is due to the fact that parallelism increases. With 
+30 chunks, only 2 CPUs per node are actively working. At 125 chunks, this 
+changes to a maximum of 9 CPUs per node, and at 750 chunks all 40 CPUs can be 
+active. It can be noted that no significant speedup can be seen between 125 
+chunks and 750 chunks. This is because out of the 50 chunks processed by each 
+node, only 40 can be processed concurrently, thus requiring two sequential 
+incrementation batches, and resulting in similar makespans as 125 
+chunks.
+
+In terms of filesystem selection, local and tmpfs perform comparably for all 
+conditions, with Lustre being significantly slower. As with varying the number 
+of iterations, Lustre is slower due to increased filesystem contention, which 
+is, at minimum, 15x greater than contention on local disk. With an increase in 
+number of chunks, local disk and tmpfs makespans begin to converge. A potential 
+explanation for this may be that tmpfs is utilizing swap space. As concurrency 
+increases, the memory footprint of the application also increases. It is 
+possible that at 750 chunks, swapping to disk is required by tmpfs, thus 
+resulting in similar processing times as local disk.
+
+Swapping may also be an explanation for the variance between Spark in-memory 
+and tmpfs performance. While Spark may also spill to disk, it only does so when
+data does not fit in memory. As none of the RDDs generated throughout the 
+pipeline were cached and all data concurrently accessed could be mantained 
+in-memory, even for the 750 chunks which were processed in two batches, it did 
+not need to spill to disk.
 
 \subsection{Data Size}
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\linewidth]{results/figures/datasize.pdf}
+    \caption{Makespans of Spark and Nipype writing to memory, tmpfs, local 
+             disk and Lustre with varying data sizes. All 
+             makespans use 125 chunks and are fixed at 
+             10 iterations with a task duration of 1.76 
+             seconds }\label{fig:datasize}
+\end{figure}
 % Comment on the general trend
+Increasing overall data size decreases performance, as can be seen in 
+Figure~\ref{fig:datasize}. When the data size is very small, as is the case 
+with the MRI image, all filesystem makespans are comparable. This is due to the 
+fact that page cache can be leveraged fully regardless of file system. However, 
+this time, Spark in-memory performed significantly worse than all other 
+filesystems. Upon further inspection, it appeared that Spark in-memory executed
+in a sequential order, on a single worker node. Lack of parallelism for the MRI 
+image may be a result of Spark's max partition size, which is by default 128~MB
+-- significantly larger than the 13~MB MRI image. Due to Spark in-memory's 
+sequential execution, it was omitted from the graph.
 
-\subsection{Iterations}
-% Comment on the general trend
+At half BigBrain, the makespan differences become apparent in both local disk 
+and Lustre, with Lustre, with Lustre becoming 2.4x slower than in-memory. This 
+can be attributed to page cache saturation, as predicted by the model for both 
+half the BigBrain image and the complete BigBrain. Only the MRI image was 
+predicted to fall within the model constraints.
+
+When the complete BigBrain is processed, the disparity between the different 
+filesystems becomes even greater. Lustre becomes 3.68x slower, whereas local 
+disk is now 1.68x slower. The reasoning for this is that the page cache fills 
+up faster than with half of BigBrain.
+
+\todo{change Spark's max partition size and test to see if it fixes the problem}
 

### 2018-12-12 results/code/d_c_graph.py

@@ -15,13 +15,15 @@ def parse_bench_file_name(file_name):
     blocks = 30
     if '125' in stringsit[1]:
         blocks = 125
-    if '750' in stringsit[1]:
+    elif '750' in stringsit[1]:
         blocks = 750
     data = 'MRI'
     if 'HBB' in stringsit[1]:
         data = 'HBB'
     elif 'BB' in stringsit[1]:
         data = 'BB'
+    if delay == 320:
+        delay = int(delay)
     
     return iterations, data, fs, wf, delay, blocks
 
@@ -34,8 +36,8 @@ def d_c_fig(bench_dir, makespan_file, out_file):
     
     bandwidths = { # MiB/s
         'lustre': 401.49,
-        'local': 292, 
-        'tmpfs': 1793.75
+        'local': 173.5, #750, #292, 
+        'tmpfs': 1400#1793.75
     }
     
     gammas = { # by number of    blocks
@@ -80,10 +82,11 @@ def d_c_fig(bench_dir, makespan_file, out_file):
     for file_name in os.listdir(bench_dir):
         b = benches[file_name]
         in_mem_file = "sp_mem_{}it{}{}_{}delay.out".format(b['iterations'],
-                                                             b['blocks'],
-                                                             b['data_file'],
-                                                             b['task_duration'])
-        if benches.get(in_mem_file) == None or 'MRI' in in_mem_file:
+                                                           b['blocks'],
+                                                           b['data_file'],
+                                                           b['task_duration'])
+        if benches.get(in_mem_file) == None:
+            print(in_mem_file)
             continue # file isn't in benchmark
         b['memory-speed-up'] = (
             b['makespan'] / benches[in_mem_file]['makespan']
@@ -115,17 +118,31 @@ def d_c_fig(bench_dir, makespan_file, out_file):
         gamma = gammas[benches[file_name]['blocks']]
         if fs == 'lustre':
             gamma = Gammas[benches[file_name]['blocks']]
[…]

### 2018-12-12 results/code/get_makespan.py

@@ -1,3 +1,4 @@
+#!/usr/bin/env python
 import sys
 import os
 
@@ -19,7 +20,7 @@ for bench in os.listdir(sys.argv[1]):
                     if start is None or start_time < start:
                         driver_line = row
                         start = start_time
-                    if end is None or end_time < end:
+                    if end is None or end_time > end:
                         end = end_time
                         driver_line[2] = str(end - start)
                         driver_line[1] = "0"

### 2018-12-12 results/code/d_c_graph.py
added missing benchmarks. removed figures. added script to automate figure generation

@@ -35,9 +35,9 @@ def d_c_fig(bench_dir, makespan_file, out_file):
     }
     
     bandwidths = { # MiB/s
-        'lustre': 401.49,
-        'local': 173.5, #750, #292, 
-        'tmpfs': 1400#1793.75
+        'lustre': 504.03,
+        'local': 193.64, 
+        'tmpfs': 1377.18
     }
     
     gammas = { # by number of    blocks
@@ -86,7 +86,6 @@ def d_c_fig(bench_dir, makespan_file, out_file):
                                                            b['data_file'],
                                                            b['task_duration'])
         if benches.get(in_mem_file) == None:
-            print(in_mem_file)
             continue # file isn't in benchmark
         b['memory-speed-up'] = (
             b['makespan'] / benches[in_mem_file]['makespan']
@@ -124,17 +123,7 @@ def d_c_fig(bench_dir, makespan_file, out_file):
             x.append((benches[file_name]['D']/benches[file_name]['C']) / (bandwidths[fs]/gamma))
         y.append(benches[file_name]['memory-speed-up'])
 
-        if (benches[file_name]['D']/benches[file_name]['C']) / (bandwidths[fs]/gamma) <= 1 and benches[file_name]['memory-speed-up'] > 1:
-            print(benches[file_name], 'LHS')
-        elif (benches[file_name]['D']/benches[file_name]['C']) / (bandwidths[fs]/gamma) > 1 and benches[file_name]['memory-speed-up'] <= 1:
-            print(benches[file_name], 'RHS')
-        #if (benches[file_name]['D']/benches[file_name]['C']) / (bandwidths[fs]/gamma) <= 1 and benches[file_name]['memory-speed-up'] <= 1:
-        #    print(benches[file_name], 'LHS')
 
-        #if benches[file_name]['file_system'] == 'tmpfs' and benches[file_name]['memory-speed-up'] > 1:
-        #    print(benches[file_name])
-
-    print(len(x_tmpfs), len(x_disk), len(x_sfs))
     from matplotlib import pyplot as plt
     from numpy import polyfit, poly1d
     fit = polyfit(x_tmpfs,y_tmpfs,1)

### 2018-12-12 results/code/gen_all_figures.sh

@@ -1,0 +1,9 @@
+#!/usr/bin/env bash
+
+python get_makespan.py ../benchmarks incrementation_makespan.out
+
+python d_c_graph.py ../benchmarks incrementation_makespan.out ../figures/incrementation.pdf
+python gen_figures.py 1 incrementation_makespan.out ../figures/iterations.pdf
+python gen_figures.py 2 incrementation_makespan.out ../figures/numchunks.pdf
+python gen_figures.py 3 incrementation_makespan.out ../figures/datasize.pdf
+python gen_figures.py 4 incrementation_makespan.out ../figures/cputime.pdf

### 2018-12-12 results/code/gen_all_figures.sh
modified script to be executed from main directory

@@ -1,9 +1,9 @@
 #!/usr/bin/env bash
 
-python get_makespan.py ../benchmarks incrementation_makespan.out
+python results/code/get_makespan.py results/benchmarks results/code/incrementation_makespan.out
 
-python d_c_graph.py ../benchmarks incrementation_makespan.out ../figures/incrementation.pdf
-python gen_figures.py 1 incrementation_makespan.out ../figures/iterations.pdf
-python gen_figures.py 2 incrementation_makespan.out ../figures/numchunks.pdf
-python gen_figures.py 3 incrementation_makespan.out ../figures/datasize.pdf
-python gen_figures.py 4 incrementation_makespan.out ../figures/cputime.pdf
+python results/code/d_c_graph.py results/benchmarks results/code/incrementation_makespan.out results/figures/incrementation.pdf
+python results/code/gen_figures.py 1 results/code/incrementation_makespan.out results/figures/iterations.pdf
+python results/code/gen_figures.py 2 results/code/incrementation_makespan.out results/figures/numchunks.pdf
+python results/code/gen_figures.py 3 results/code/incrementation_makespan.out results/figures/datasize.pdf
+python results/code/gen_figures.py 4 results/code/incrementation_makespan.out results/figures/cputime.pdf

### 2018-12-12 benchmark_scripts/measure_bandwidth.py
added computed average, disk bandwidth measurements

@@ -1,0 +1,68 @@
+#!/usr/bin/env python
+import nibabel as nib
+from random import randint, shuffle
+import sys
+from os import listdir, makedirs
+from os.path import abspath, join, basename
+from json import dump
+from time import time
+from numpy import eye
+from shutil import rmtree
+
+
+def main():
+
+    assert len(sys.argv) == 7, ('data_dir, output_dir, min_num_splits, '
+                                 'max_num_splits, step, bench')
+
+    data_dir = sys.argv[1]
+    output_dir = sys.argv[2]
+    min_num_splits = int(sys.argv[3])
+    max_num_splits = int(sys.argv[4])
+    step = int(sys.argv[5])
+    bench = sys.argv[6]
+
+    
+    data_dict = {}
+    bench_dict = {}
+
+    bb_files = [abspath(join('data/125_splits', i)) for i in listdir(data_dir)]
+    num_chunks = len(bb_files)
+
+    order = [i for i in range(min_num_splits, max_num_splits, step)]
+    shuffle(order)
+
+    for chunk in bb_files:
+        data_dict[chunk] = nib.load(chunk).get_data()
+
+    for i in order:
+
+        try:
+            makedirs(output_dir)
+        except Exception as e:
+            pass
+
+        write = 0
+        print('writing data...', i)
+
+        for j in range(0, i):
+            print(j)
+            pos = j % num_chunks
+            data_dict[bb_files[pos]] = data_dict[bb_files[pos]] + randint(1, 50)
+            im = nib.Nifti1Image(data_dict[bb_files[pos]], eye(4))
+            start = time()
+
+            nib.save(im, join(output_dir, 'bb-{}.nii'.format(j % 54)))
+
+            write += (time() - start)
+
+        print('Num chunks: ', i, ' Write: ', write)
+        bench_dict[i] = write
+        rmtree(output_dir)
+    
+    with open(bench, 'w') as f:
+        dump(bench_dict, f)
+
+
+if __name__=='__main__':
+    main()

### 2018-12-12 benchmark_scripts/nipype_kmeans_template.sh

@@ -1,0 +1,14 @@
+#!/usr/bin/env bash
+#SBATCH --partition requestq       # Queue (Partition in slurm)
+#SBATCH --reservation=spark
+#SBATCH --time=99:00:00
+#SBATCH --nodes=1
+#SBATCH --mem=183G
+#SBATCH --cpus-per-task=40
+#SBATCH --ntasks-per-node=1
+
+source /mnt/lustrefs/spark/env/exp0/bin/activate
+export PATH=/mnt/lustrefs/spark/SOEN691-project/pipelines/:$PATH
+export PYTHONPATH=/mnt/lustrefs/spark/SOEN691-project/pipelines:$PYTHONPATH
+
+

### 2018-12-12 benchmark_scripts/run_benchmarks.py

@@ -38,7 +38,7 @@ data_dir = '/mnt/lustrefs/spark/data'
 
 tmpfs = '/dev/shm/inc_exp'
 local = '/local/inc_exp'
-lustre = '/mnt/lustrefs/spark/inc_exp'
+lustre = '/mnt/lustrefs/spark/inc_exp_rem'
 
 done_file = op.join(lustre, "COMPLETED")
 
@@ -146,7 +146,8 @@ for cdn in conditions:
 
         print("Submitting command: ", cmd)
 
-        s.run("bash " + spark_template, cmd_kwargs={"spscript": cmd},
+        s.run("bash " + spark_template, cmd_kwargs={"spscript": cmd,
+                                                    "parallelism": slurm_conf["cpus-per-task"]},
               _cmd=sys.argv[2])
     else:
         with open(legends[cdn["dataset"]]) as legend:

### 2018-12-12 benchmark_scripts/spark_sbatch.sh

@@ -40,7 +40,9 @@ SPARK_NO_DAEMONIZE=1 srun -n ${NWORKERS} -N ${NNODES} --label --output=$SPARK_LO
 slaves_pid=$!
 
 echo 'Starting execution'
-srun -n 1 -N 1 $SPARK_HOME/bin/spark-submit --master=${MASTER_URL} --num-executors=${NEXECUTORS} --executor-memory=${MEM_PER_EXEC}M --driver-memory=${MEM_PER_EXEC}M --conf spark.locality.wait=600s  $spscript
+srun -n 1 -N 1 $SPARK_HOME/bin/spark-submit --master=${MASTER_URL} --num-executors=${NEXECUTORS} \
+    --executor-memory=${MEM_PER_EXEC}M --driver-memory=${MEM_PER_EXEC}M --conf spark.locality.wait=600s \
+    --conf spark.default.parallelism=$parallelism $spscript
 
 kill $slaves_pid
 $SPARK_HOME/sbin/stop-master.sh

### 2018-12-12 pipelines/increment.py

@@ -13,7 +13,7 @@ except Exception as e:
 
 
 
-def increment(fn, outdir, delay, benchmark_file, start_time):
+def increment(fn, outdir, delay, benchmark_file, start_time, avg):
     print('Incrementing image: ', fn)
 
     start = time.time() - start_time
@@ -26,6 +26,9 @@ def increment(fn, outdir, delay, benchmark_file, start_time):
                     socket.gethostname(), op.basename(fn), get_ident())
 
     inc_data = im.get_data() + 1
+
+    if avg is not None:
+        inc_data += nib.load(avg).get_data().astype(inc_data.dtype, copy=False)
 
     im = nib.Nifti1Image(inc_data, affine=im.affine, header=im.header)
 
@@ -68,6 +71,8 @@ def main():
                         help='start time of the application')
     parser.add_argument('--delay', type=float, default=0,
                         help='task duration time (in s)')
+    parser.add_argument('--avg', type=str,
+                        help='average chunk to be added to incremented chunk')
 
     args = parser.parse_args()
 
@@ -77,7 +82,7 @@ def main():
         pass
 
     increment(args.filename, args.output_dir, args.delay,
-              args.benchmark_file, args.delay)
+              args.benchmark_file, args.delay, args.avg)
 
 
 if __name__ == '__main__':

### 2018-12-12 pipelines/nipype_inc.py

@@ -2,6 +2,7 @@
 
 from nipype import Workflow, MapNode, Node, Function
 from time import time
+from math import ceil
 import argparse
 import os
 import shutil
@@ -15,7 +16,8 @@ except Exception as e:
     from thread import get_ident
 
 
-def increment_chunk(chunk, delay, benchmark, benchmark_dir=None, cli=False):
+def increment_chunk(chunk, delay, benchmark, benchmark_dir=None, cli=False,
+                    avg=None):
     import nibabel as nib
     import os
     import socket
@@ -48,6 +50,11 @@ def increment_chunk(chunk, delay, benchmark, benchmark_dir=None, cli=False):
 
         data += 1
 
+        if avg is not None:
+            avg_chunk = nib.load(avg).get_data().astype(data.dtype, copy=False)
+
+            data += avg_chunk
+
         inc_im = nib.Nifti1Image(data, im.affine)
         nib.save(inc_im, inc_chunk)
 
@@ -56,38 +63,39 @@ def increment_chunk(chunk, delay, benchmark, benchmark_dir=None, cli=False):
     else:
         program = 'increment.py'
 
+        cmd = [program, chunk, os.getcwd(), '--delay', str(delay)]
         if benchmark:
             benchmark_file = os.path.join(benchmark_dir,
                                           "bench-{}.txt".format(
                                               str(uuid.uuid1())))
-            p = subprocess.Popen([program, chunk, os.getcwd(),
-                                  '--benchmark_file', benchmark_file,
-                                  '--delay', str(delay)],
-                                 stdout=subprocess.PIPE,
-                                 stderr=subprocess.PIPE)
-        else:
-            p = subprocess.Popen([program, chunk, os.getcwd(),
-                                  '--delay', str(delay)],
-                                 stdout=subprocess.PIPE,
-                                 stderr=subprocess.PIPE)
+            cmd.extend(['--benchmark', benchmark_file])
+
+        if avg is not None:
+            cmd.extend(['--avg', avg])
+
+        p = subprocess.Popen(cmd, stdout=subprocess.PIPE,
+                             stderr=subprocess.PIPE)
 
         (out, err) = p.communicate()
         print(out)
         print(err)
         if 'inc-' not in inc_chunk:
             inc_chunk = 'inc-{}'.format(inc_chunk)
-        inc_file = os.path.join(os.getcwd(), inc_chunk)
[…]

### 2018-12-12 tests/test_comp_avg.py

@@ -1,0 +1,135 @@
+import pytest
+import subprocess
+import hashlib
+import os
+import shutil
+from os import path as op
+import nipype_inc as ca
+import nibabel as nib
+import numpy as np
+
+
+chunk  = op.abspath('sample_data/dummy_1.nii')
+chunks = [chunk, op.abspath('sample_data/dummy_3.nii')]
+
+def test_increment():
+    delay = 0
+    benchmark = False
+    cli = True
+    im = ca.increment_chunk(chunk, delay, benchmark, cli=True)
+
+    assert op.isfile(im)
+
+    original = nib.load(chunk).get_data()
+    incremented = nib.load(im).get_data()
+
+    assert np.array_equal(incremented, original + 1)
+
+def test_increment_wf():
+    delay = 0
+    benchmark = False
+    benchmark_dir = None
+    cli = True
+    wf_name = 'test_incwf'
+    avg = None
+    work_dir = 'test_incwf_work'
+    inc_chunks = ca.increment_wf([chunks], delay, benchmark, benchmark_dir,
+                                  cli, wf_name, avg, work_dir)
+    for im in inc_chunks:
+        assert op.isfile(im)
+
+def test_compute_avg():
+    benchmark = False
+    benchmark_dir = None
+    avg = ca.compute_avg([chunks], benchmark, benchmark_dir)
+
+    images = [i for c in [chunks] for i in c]
+
+    data = None
+    print(images)
+
+    for im in images:
+        if data is None:
+            data = nib.load(im).get_data()
+        else:
+            data += nib.load(im).get_data().astype(data.dtype, copy=False)
+
+    data = data / len(images)
+
+    assert op.isfile(avg)
+    assert np.array_equal(nib.load(avg).get_data(), data)
+
+def test_compute_avg_wf():
+    from nipype import Workflow
+
+    nnodes = 1
+    work_dir = 'test_ca_wf_work'
+    chunk = chunks
+    delay = 0
+    benchmark_dir = None
+    benchmark = False
+    cli = True
+
+    wf = Workflow('test_ca_wf')
+    wf.base_dir = work_dir
+
+    inc_1, ca_1 = ca.computed_avg_node('ca_bb',
+                                       nnodes, work_dir,
+                                       chunk=chunk,
+                                       delay=delay,
+                                       benchmark_dir=benchmark_dir,
+                                       benchmark=benchmark,
[…]

### 2018-12-13 paper.tex
Complete read and editing

@@
 \usepackage{graphicx}
-\usepackage{color}
+\usepackage{xcolor}
 \usepackage{colortbl} % for \rowcolor
@@
 \usepackage[bookmarks=false]{hyperref}
+\usepackage{algorithm}
+\usepackage{algorithmicx} % Doc is at http://tug.ctan.org/macros/latex/contrib/algorithmicx/algorithmicx.pdf
 \usepackage{algpseudocode}
-\usepackage{algorithm}
 \algnewcommand\algorithmicforeach{\textbf{for each}}
 \algdef{S}[FOR]{ForEach}[1]{\algorithmicforeach\ #1\ \algorithmicdo}
+\usepackage[bookmarks=false]{hyperref}
+\hypersetup{colorlinks=true,linkcolor=black,citecolor=blue,filecolor=black,urlcolor=blue}
 
@@
 \definecolor{headcolor}{gray}{0.9}
+
+\algblock{Input}{EndInput}
+\algnotext{EndInput}
+\algblock{Output}{EndOutput}
+\algnotext{EndOutput}
+\newcommand{\Desc}[2]{\State \makebox[2em][l]{#1}#2}
 
@@
 computing, adopted by MapReduce's successor Apache 
-Spark~\cite{zaharia2016apache}, caches data in memory whenever 
+Spark~\cite{zaharia2016apache}, stores data in memory whenever 
 possible, reducing costly disk I/Os between pipeline steps. 
@@
 cases, which might explain the remaining gap between Big Data and 
-scientific engines. \todo{This sentence is extremely long} While Big Data applications mostly target text 
+scientific engines. While Big Data applications mostly target text 
 processing (e.g. Web search, frequent pattern mining, recommender 
-systems~\cite{leskovec2014mining}), implemented in consistent software 
-libraries and running on clouds or dedicated commodity clusters with 
-locality-aware file systems such as the Hadoop Distributed File System 
-(HDFS~\cite{shvachko2010hadoop}), scientific applications often involve 
+systems~\cite{leskovec2014mining}) implemented in consistent software 
+libraries, scientific applications often involve 
 binary data such as images and signals, processed by command-line tools 
 using a mix of programming languages (C, Fortran, Python, shell 
-scripts), and are deployed on large, shared clusters where 
-data is fetched from data nodes to compute nodes using centralized shared file 
-systems such as Lustre~\cite{schwan2003lustre}. Such differences in 
-applications and infrastructures have important consequences. To 
+scripts). Infrastructure-wise, Big Data applications commonly run on 
+clouds or dedicated commodity clusters with locality-aware file systems 
+such as the Hadoop Distributed File System 
+(HDFS~\cite{shvachko2010hadoop}), whereas scientific applications are 
+usually deployed on large, shared clusters where data is moved between
+data nodes and compute nodes using shared file systems such 
+as Lustre~\cite{schwan2003lustre}. Such differences in applications and 
+infrastructures have important consequences. To 
 mention only one, in-memory computing requires instrumentation to be 
-applied to command-line tools. Moreover, performance is only one of the important
-aspects to scientific computing. Other important aspects include reproducibility 
-and provenance tracking. While these are minimally addressed by Big Data
-processing engines, it is currently insufficient for scientific computing. 
-Incorporation of provenance techniques into frameworks such are Spark are
-currently being investigated~\cite{samba}.
+applied to command-line tools.
 
@@
 in the Linux kernel~\cite{love2010linux}, in-memory file systems 
-(tmpfs) and memory-mapped files\todo{check that} might also 
+(tmpfs) and memory-mapped files might also 
 explain the lack of adoption of Big Data engines for scientific 
@@
 improvements are expected from in-memory computing, lazy 
-evaluation\todo{are we talking about lazy evaluation at all?} or data 
+evaluation or data 
 locality. On the other hand, neuroimaging data processing engines are 
@@
 compare Big Data and scientific data processing engines on the grounds 
-of workflow language expressivity, fault-tolerance, provenance capture and 
-representation, portability or reproducibility, which are otherwise 
-critical concerns. Besides, our study of performance focuses on the 
-impact of data writes and transfers. It purposely leaves out task 
-scheduling to computing resources, to focus on the understanding of 
+of workflow language expressivity, fault-tolerance, provenance capture 
+and representation, portability or reproducibility, which are otherwise 
+critical concerns, addressed for instance in~\cite{samba}. Besides, our 
+study of performance focuses on the impact of data writes and 
+transfers. It purposely leaves out task scheduling to computing 
+resources, to focus on the understanding of 
 data writes and movement. Task scheduling will be part of our 
@@
 (HPC) clusters that are typically available through University 
-facilities or national computing infrastructures such as XSEDE, Compute 
-Canada or DEISA \todo{add links}, as this is the type of platform which 
-is typically available to neuroscientists. We assume that HPC systems 
-are multi-tenant, that compute nodes are accessible through a batch 
+facilities or national computing infrastructures such as 
+\href{xsede.org}{XSEDE}, \href{http://computecanada.ca}{Compute Canada} 
+or \href{http://www.prace-ri.eu}{PRACE}, as neuroscientists typically use such platforms.
+ We assume that HPC systems are 
+multi-tenant, that compute nodes are accessible through a batch 
 scheduler, and that a file system shared among the compute nodes is 
-available, for instance Lustre. We intentionally leave distributed, 
+available. We intentionally leave distributed, 
 HDFS-like file system out, as initiatives to deploy them in HPC 
-centers, for instance Hadoop on-demand~\cite{hadoop-on-demand} have not 
+centers, for instance Hadoop on-demand~\cite{krishnan2011myhadoop}, have not 
 become mainstream yet.
@@
 
-Our methods, including performance models, processing engines, \todo{Make this paragraph more informative}
+Our methods, including performance models, processing engines, 
 applications and infrastructure used are described in 
 Section~\ref{sec:methods}. Section~\ref{sec:results} presents our 
-results which we discuss in Section~\ref{sec:discussion}. 
-Section~\ref{sec:conclusion} concludes on the answer to our two research questions.
+results which we discuss in Section~\ref{sec:discussion} along with the 
+two research questions mentioned previously. 
+Section~\ref{sec:conclusion} concludes on the relevance of Big Data 
+processing strategies for neuroimaging applications.
 
@@
 
-\todo{Maybe add an intro to this section, to outline it}
+The pipelines, benchmarks, performance data, and analysis scripts used 
+to implement the methods described hereafter are all available at 
+\url{https://github.com/big-data-lab-team/paper-in-mem-locality} for 
+further inspection and reproducibility. Links to the pipeline engines 
+and processed data are provided in the text.
 
@@
 
-An RDD is an immutable parallel data structure that achieve fault-tolerance 
+An RDD is an immutable parallel data structure that achieves fault-tolerance 
 through the concept of lineage~\cite{zaharia2010spark}. Rather than permitting
@@
 Spark transformations include map, filter and join. To materialize an RDD, an
-action must be performed, such as a reduce or a collect.
+action must be performed, such as a reduce or a collect. \todo{Lazy evaluation should be mentioned here, or elsewhere in the paragraph.
+It is currently not described.}
 
@@
 Executing Spark applications on HPC systems with Spark-unaware schedulers may
-be inefficient. The amount of resources requested by Spark may impede Spark-cluster
-scheduling time. Using pilot-scheduling strategies to add nodes to the Spark cluster
-as they are allocated by the underlying schedulers may speedup allocation and 
-overall processing time. This, however, is not studied in the current paper.
+be inefficient. The amount of resources requested by Spark may impede 
+Spark-cluster scheduling time. Using pilot-scheduling strategies to add 
+nodes to the Spark cluster as they are allocated by the underlying 
+schedulers may speedup allocation and overall processing time 
+\todo{Cite Shantenu's paper here}. This, however, is not studied in the 
+current paper.
 
-As Spark is frequently used by the scientific community, we designed our 
-experiments using the PySpark API. This is at a cost to performance as the PySpark
-code must undergo Python to Java serialization.
+As Spark is frequently used by the scientific community, we designed 
+our experiments using the PySpark API. This is at a cost to performance 
+as the PySpark code must undergo Python to Java serialization. We used 
+version 2.3.2 installed from \url{https://spark.apache.org}.
 
@@
 
+We used Nipype version 1.1.4 installed through the pip package manager.
+
 % pipeline description: map nodes and their limitations
@@
                                   \hline          
-In-memory                         &   Y           & Y                       \\
-tmpfs                             &   Y           & Y                   \\
-Local Disk                        & Page Caching  & Y                   \\
-Shared File System                & Page Caching  & N                
+In-memory                         & \cellcolor{green!25} Yes           & \cellcolor{green!25}Yes                      \\
+tmpfs                             & \cellcolor{green!25} Yes           & \cellcolor{green!25}Yes                  \\
+Local Disk                        & \cellcolor{orange!25}Page Caching  & \cellcolor{green!25}Yes                  \\
+Shared File System                & \cellcolor{orange!25}Page Caching  & \cellcolor{red!25}No                
 % data locality is somehow present in shared fs due to page caching again.
@@
  related to what happens when the processed data fills up the available 
- memory. In case engine memory is used, the engine has to cleanup 
+ memory. In case engine memory is used, the engine must cleanup 
  unused data with some strategy to avoid crashes. In case an in-memory 
@@
 
-\subsubsection{Local Disk (Page Caching)} % 0.5 page
+\subsubsection{Local Disk} % 0.5 page
 
@@
 transfer input or output data. However, in absence of a more specific 
-file system such as the Hadoop Distributed File System (HDFS) to handle 
+file system such as HDFS to handle 
 file replication across computing nodes, data locality comes at the price
@@
 lies in the page cache, realizing a cache \emph{hit}. Cache hits drastically speed-up data 
-reads, by masking the disk latency and limited bandwidth behind a 
+reads, by masking the disk latency and bandwidth behind a 
 memory buffer. In effect, page caching provides in-memory computing 
 transparently to the processing engine. However, page cache eviction 
-strategies currently cannot be controlled by the application 
-\todo{check that}, which prevents processing engines to anticipate 
-reads by preloading the cache. Scheduling strategies might be designed 
+strategies currently cannot be controlled by the application, which 
+prevents processing engines to anticipate reads by preloading the 
+cache. Scheduling strategies might be designed 
 to maximize cache hits though. For instance, lazy 
@@
 \item $\delta$ is the disk bandwidth
-\item $\gamma$ is the max number of concurrent processes
+\item $\gamma$ is the max number of concurrent processes on a node
 \item $d_0$ is the amount of data in the page cache at time $t_0$
@@
 This defines a data-compute ratio beyond which the page cache becomes 
-asymptotically useless. We will use this ratio to define our benchmarks. It 
+asymptotically useless. It should be noted that the transient phase 
+during the page cache fills up might last a significant amount of time, 
+in particular when $\dot d$ is positive but small. We intentionally do not model the 
+transient phase as it requires detailed knowledge on the size of the 
+page cache and the initial amount 
+of data in it ($d_0$), which are difficult to estimate.
+
+ We will use Equation~\ref{eq:page-cache-inequality} to 
+define our benchmarks and interpret the results. It 
 should be noted that leveraging the page cache, and therefore ensuring 
 that Equation~\ref{eq:page-cache-inequality} holds, has important 
-performance implications: with page caching, the application will 
-experience negligible write times, while without page caching 
-the write throughput will be that of the disk.
+performance implications: with page caching, the write throughput will 
+be that of memory, while without page caching it will be that of the 
+disk.
 
 % Cache eviction: LRU/n.
-
-\todo{How about reads?}
 
@@
 We model a shared file system using its global apparent bandwidth 
-$\Delta$, shared by all concurrent processes in the system. We are 
+$\Delta$, shared by all concurrent processes in the cluster. We are 
 aware that such a simplistic model does not describe at all the 
@@
 
-As in the local disk model, we note that page caching will only be 
+As in the local disk model, we note that page caching will only be asymptotically
 useful when the flush bandwidth is greater than the write throughput of 
@@
 \end{equation}
-where $\Gamma$ is the max number of concurrent jobs in the cluster. 
-Note that $\frac{\Delta}{\Gamma}$ will be much lower than 
-$\frac{\delta}{\gamma}$ in Equation~\ref{eq:page-cache-inequality}.
+where $\Gamma$ is the max number of concurrent processes in the cluster. 
+Note that $\frac{\Delta}{\Gamma}$ will usually be much lower than 
+$\frac{\delta}{\gamma}$.
 
@@
 
-\subsection{Experiments} % 1.5 pages
+\subsection{Infrastructure} % 1/4 a page (half a col)
 
-\subsubsection{Infrastructure} % 1/4 a page (half a col)
-
-All experiments were executed on the Dell Zenith cluster~\cite{dellHPC} running 
-SLURM. For the 
-Spark experiments, a Spark cluster was started atop our SLURM allocation. We
-had access to 16 dedicated nodes. Each compute node ran on RHEL 7.4, was 
+ All experiments were executed on  
+ \href{https://www.dellemc.com/resources/en-us/asset/sales-documents/products/storage/h16221-hpc-lab-brochure.pdf}{Dell EMC's Zenith 
+ cluster}, a Top-500 machine in the Dell EMC HPC Innovation Lab, running 
+ SLURM. For the Spark experiments, a Spark cluster was started atop our 
+ SLURM allocation. We had access to 16 dedicated nodes. Each compute 
+ node ran on RHEL 7.4, was 
 equipped with 40 Intel Xeon Gold 6148 2.4GHz CPUs and had
 187GB of memory, 94GB of tmpfs and a Dell 120GB M.2 SATA SSD as local disk. 
-A 698TB Lustre server is accessible on each node through a 100Gb/s Infiniband
-network. The apparent bandwidth of the local disk was found to be 1480MB/s, 
-whereas Lustre's bandwidth was found to be at 401.49MB/s.
+A 698TB Lustre server was accessible on each node through a 100Gb/s 
+Infiniband network. The apparent write bandwidth of the local disk, 
+Lustre file system and tmpfs were measured by sequentially writing various numbers
+of image blocks containing random intensities, to avoid 
+caching effects (measure\_bandwidth.py script in folder 
+benchmark\_scripts). They are reported in Table~\ref{table:bdwdths}.
+
+\begin{table}
+\centering
+\begin{tabular}{c|c}
+\rowcolor{headcolor}
+Data location & Measured write bandwidths (MB/s)\\
+\hline
+tmpfs                 & 1377.18 \\
+Local disk ($\delta$) & 193.64  \\
+Lustre   ($\Delta$)   & 504.03 \\
+\end{tabular}
+\caption{Measured bandwidths}
+\label{table:bdwdths}
+\end{table}
+
 % Dedicated Compute nodes with SLURM
@@
 
-\subsubsection{Datasets} % 1/4 of a page (half a col)
+\subsection{Datasets} % 1/4 of a page (half a col)
 
@@
 % Some BIDS dataset for fmriprep
-We used BigBrain~\cite{bigbrain}, a 75GB 40$\mu$m isotropic 
+We used BigBrain~\cite{amunts2013bigbrain}, a 75GB 40$\mu$m isotropic 
 histological image of a 65-year-old male's brain. The BigBrain was selected due 
@@
 
-Processing large images are only considered to be part of the BigData problem 
+Processing large images is only considered to be part of the Big Data problem 
 in neuroscience. The other problem is processing large MRI datasets, that is,
@@
 subjects. Although we have not explored explicitly the processing of large MRI
-datasets, the 75GB BigBrain is within the size ballpark~\cite{VanHorn2014} of 
-what a large MRI dataset would be today.
+datasets, the 75GB BigBrain is within the size ballpark~\cite{van2014human} of 
+MRI datasets commonly processed in today's studies.
 
 As it is possible that both small and large data may need to be processed using
-the same pipeline \todo{explain pipelines in intro if not done already}, we 
+the same analysis pipeline, we 
 wanted to determine the effects of the data management strategies on small 
-data as well. For this, we selected a 12MB T1W image belonging to subject 1 of 
-the ds000001 dataset version 6 ~\cite{openneuro}. In order to be able to split 
-it into 125 chunks, it was necessary to zero-pad the image to the dimensions
+data as well. For this, we selected a 12MB T1W image belonging to 
+subject 1 of OpenNeuro's 
+\href{https://openneuro.org/datasets/ds000001/versions/00006/file-display/sub-01:anat:sub-01_T1w.nii.gz}{ds000001 
+dataset version 6}. In order to be able to split it into 125 chunks, it 
+was necessary to zero-pad the image to the dimensions
 $165\times200\times200$ voxels, which subsequently increased the total image size to 13MB.
 
-\subsubsection{Applications} % 1 page
+\subsection{Applications} % 1 page
 \begin{algorithm}\caption{Incrementation}\label{alg:incrementation}
-    \begin{algorithmic}
-    \Require $x$, a sleep delay in float
-    \Require $fs$, filesystem to save image to (i.e. mem, tmpfs, local disk, lustre). 
-    \ForEach{$i \in iterations$}
-    \ForEach{$chunk \in image$}
+    \begin{algorithmic}[1]
+    \Input
+    \Desc{$x$}{a sleep delay in seconds}
+    \Desc{$n$}{a number of iterations}
+    \Desc{$C$}{a set of image chunks}
+    \Desc{$fs$}{filesystem to write to (mem, tmpfs, local disk, Lustre).}
+    \EndInput
+    \ForEach{$chunk \in C$}
+    \State read $chunk$ from Lustre
+    \For{$i \in [1, n]$}
         \State $chunk\gets chunk+1$
         \State sleep $x$
+        \If{$i < n$}
         \State save $chunk$ to $fs$
+        \EndIf
     \EndFor
+    \State write $chunk$ to Lustre
     \EndFor
@@
 \end{algorithm}
-
-\begin{figure*}
-\begin{subfigure}{0.33\linewidth}
-\centering
-    \includegraphics[width=\textwidth]{results/figures/tmpfs-incrementation.pdf}
-\caption{}
-\label{fig:sub1}
-\end{subfigure}%
-    \begin{subfigure}{0.33\linewidth}
-\centering
-    \includegraphics[width=\textwidth]{results/figures/local-incrementation.pdf}
-\caption{}
-\label{fig:sub2}
-    \end{subfigure}%
-\begin{subfigure}{0.33\linewidth}
-\centering
-    \includegraphics[width=\textwidth]{results/figures/lustre-incrementation.pdf}
-\caption{}
-\label{fig:sub3}
-\end{subfigure}
-    \caption{Model evaluation on (a) tmpfs, (b) local and (c) Lustre. Grey 
-             regions denote areas that violate model predictions.}
-\label{fig:modeleval}
-\end{figure*}
 
@@
 process, we added a sleep delay to the incrementation stage in order to 
-increase task duration. The incremented chunks would be either 
+increase task duration as required. The incremented chunks would be either 
 maintained in-memory (Spark only) or saved to either tmpfs, local disk or 
@@
 
-\subsubsection{Setup}
+\subsection{Experiments}
+
+We conducted 4 experiments in which we varied (1) the number of 
+iterations in the application ($n$ in 
+Algorithm~\ref{alg:incrementation}), (2) the task duration ($x$), (3) 
+the chunk size for a constant image size, (4) the total image size. To 
+evaluate the page-cache model, experiment conditions fell in different 
+regions of Equations~\ref{eq:page-cache-inequality} 
+and~\ref{eq:page-cache-sharedfs}, as summarized in 
+Table~\ref{table:experiments}. Among the 16 nodes available, 1 was 
+dedicated to the Spark master and driver, and the remaining 15 were 
+used as compute nodes. Since data-locality is not normally preserved in 
+Nipype (a new SLURM allocation is requested for each processed task), 
+we instrumented Nipype to ensure data locality. That is, for each chunk 
+partition, we requested a SLURM allocation to process the entire 
+pipeline in parallel, using Nipype's MultiProc scheduler, on a given 
+node. This was possible as no communication was required between the 
+processed chunks.
+
+\begin{table*}
+\centering
+\begin{tabular}{c|ccc|cccc|cc}
+  \rowcolor{headcolor}
+  \multicolumn{10}{c}{Experiment 1: Number of Iterations}\\
+  \hline
+  \rowcolor{headcolor}
+ n &
+ D (GB) & C (s) & D/C (MB/s) &
+ $\gamma$ & $\delta/\gamma$ (MB/s) & $\Gamma$ & $\Delta/\Gamma$ (MB/s)&
+ \parbox{2cm}{Page-cache on local disk? (Eq.~\ref{eq:page-cache-inequality})} & \parbox{2cm}{Page cache on Lustre? (Eq.~\ref{eq:page-cache-sharedfs})}\\
+ \hline
+1   &75    & 430    & 178.6 & 9 & 21.5 & 125 & 4.0 & \cellcolor{red!20} No & \cellcolor{red!20} No\\
+10  & 750   & 4,300 & 178.6 & 9 & 21.5 & 125 & 4.0 & \cellcolor{red!20} No & \cellcolor{red!20} No\\
+100 & 7,500 & 43,000& 178.6 & 9 & 21.5 & 125 & 4.0 & \cellcolor{red!20} No & \cellcolor{red!20} No\\
+\hline
+  \multicolumn{10}{c}{}\\
+
+\rowcolor{headcolor}
+  \multicolumn{10}{c}{Experiment 2: Task Duration}\\
+  \hline
+  \rowcolor{headcolor}
+ x (s) &
+ D (GB) & C (s) & D/C (MB/s) &
+ $\gamma$ & $\delta/\gamma$ (MB/s) & $\Gamma$ & $\Delta/\Gamma$ (MB/s)&
+ \parbox{2cm}{Page-cache on local disk? (Eq.~\ref{eq:page-cache-inequality})} & \parbox{2cm}{Page cache on Lustre? (Eq.~\ref{eq:page-cache-sharedfs})}\\
+ \hline
+ 2.4  & 750 & 3,000   & 256   & 9 & 21.5 & 125 & 4.0 &  \cellcolor{red!20} No & \cellcolor{red!20} No   \\
+ 3.44 & 750 & 4,300   & 178.6 & 9 & 21.5 & 125 & 4.0 &  \cellcolor{red!20} No & \cellcolor{red!20} No   \\
+ 7.68 & 750 & 9,600   & 80    & 9 & 21.5 & 125 & 4.0 &  \cellcolor{red!20} No & \cellcolor{red!20} No   \\
+ 320  & 750 & 400,000 & 1.9   & 9 & 21.5 & 125 & 4.0 &  \cellcolor{green!20} Yes & \cellcolor{green!20} Yes  \\
+ \hline
+  \multicolumn{10}{c}{}\\
+
+ \rowcolor{headcolor}
+  \multicolumn{10}{c}{Experiment 3: Number of Chunks}\\
+  \hline
+  \rowcolor{headcolor}
+ chunks &
+ D (GB) & C (s) & D/C (MB/s) &
+ $\gamma$ & $\delta/\gamma$ (MB/s) & $\Gamma$ & $\Delta/\Gamma$ (MB/s)&
+ \parbox{2cm}{Page-cache on local disk? (Eq.~\ref{eq:page-cache-inequality})} & \parbox{2cm}{Page cache on Lustre? (Eq.~\ref{eq:page-cache-sharedfs})}\\
+ \hline
+ 30  & 750 & 4,400   & 174.6 & 2  & 96.8 & 30  & 16.8 &  \cellcolor{red!20} No & \cellcolor{red!20} No   \\
+ 125 & 750 & 4,400   & 174.6 & 9  & 21.5 & 125 & 4.0 &  \cellcolor{red!20} No & \cellcolor{red!20} No   \\
+ 750 & 750 & 4,400   & 174.6 & 25 & 7.7  & 375 & 1.3 &  \cellcolor{red!20} No & \cellcolor{red!20} No  \\
+
+\hline
+  \multicolumn{10}{c}{}\\ 
+ \rowcolor{headcolor}
+  \multicolumn{10}{c}{Experiment 4: Image Size}\\
+  \hline
+  \rowcolor{headcolor}
+ image  &
+ D (GB) & C (s) & D/C (MB/s) &
+ $\gamma$ & $\delta/\gamma$ (MB/s) & $\Gamma$ & $\Delta/\Gamma$ (MB/s)&
+ \parbox{2cm}{Page-cache on local disk? (Eq.~\ref{eq:page-cache-inequality})} & \parbox{2cm}{Page cache on Lustre? (Eq.~\ref{eq:page-cache-sharedfs})}\\
+ \hline
+ BigBrain      & 750   & 2,200   & 349.1     & 9  & 21.5  & 125 & 4.0 &  \cellcolor{red!20} No & \cellcolor{red!20} No   \\
+ Half BigBrain & 375   & 2,200   & 174.6     & 9  & 21.5  & 125 & 4.0 &  \cellcolor{red!20} No & \cellcolor{red!20} No   \\
+ MRI           & 0.127 & 2,200   & 0.06 & 9& 21.5  & 125 & 4.0 &  \cellcolor{green!20} Yes & \cellcolor{green!20} Yes  
+\end{tabular}
+\caption{Experiment conditions}
+\label{table:experiments}
+\end{table*}
 
 For our first incrementation experiment, we investigated the effects of 
-in-memory computing on varying total data size. To do this, we increased the 
-number of incrementation iterations from 1, 10 and 100 times. The total data 
-size would then increase from 75GB, at 1 iteration, to 7500GB, at 100 
-iterations. In order
-to validate the model, the D/C was varied to fall in different regions of 
-Equation~\ref{eq:page-cache-inequality}. The D/C selected were 256, 178.6, 80 
-and 1.9MB/s\todo{present this as a table? yes, I think it would be clearer (TG) but not mandatory.}. 80 and 1.9MB/s satisfied the 
-inequality for local disk, whereas only 1.9MB/s satisfied the inequality for 
-Lustre. These D/C ratios were applied to each iteration experiment, such that 
-we were able to get both the effect of increasing the total amount of data, in 
-addition to, increasing the task duration for a given data size.
+Big Data strategies on varying total data size. To do this, we 
+increased the number of incrementation iterations from 1, 10 and 100 
+times. The total data size would then increase from 75GB, at 1 
+iteration, to 7,500GB, at 100 iterations. The total number of chunks 
+was 125. Chunks all ran concurrently ($\Gamma$=125) and they were 
+equally balanced among 15 nodes, leading to 8 or 9 concurrent jobs per 
+node ($\gamma$=9). Task duration was fixed at 3.44 seconds.
 
-We combined the first experiment with our second experiment such that we could 
-evaluate the effects of in-memory computing on varying task duration. If the 
-page cache has sufficient time to flush, it should be expected that using 
-in-memory computing should have no increased speed-up over disk. We maintained 
-the D/C ratios as those mentioned above, but rather than applying only one to
-a different number of iterations, we applied all four D/C to all numbers of 
-iterations. That is 1 iteration, 10 iterations and 100 iterations were all 
-processed with using the aforementioned D/C ratios.
+In the second experiment, we evaluated the effects of Big Data 
+strategies on varying task duration. If the page cache has sufficient 
+time to flush, it should be expected that using in-memory computing 
+should have no increased speed-up over disk. We varied the task 
+duration between 2.4 seconds and 320 seconds so that the D/C falls into 
+different regions of Equations~\ref{eq:page-cache-inequality} 
+and~\ref{eq:page-cache-sharedfs}. The number of chunks was maintained 
+at 125, leading to $\Gamma$=125 and $\gamma$=9. The number of iterations was
+fixed to 10.
 
 As a third incrementation experiment, we were interested in the effects of 
-chunk size on in-memory computing. Naturally, greater chunk size signifies 
-a decrease in parallelization, however, it also means an increase in sequential
-I/O. For this experiment we incremented the 30, 125 and 750 chunk complete
-BigBrain image. The amount of parallelization for 30 chunks was 2 cores/node, 
-whereas it was 9 cores/node for 125 chunks and 25 cores/node for 750 chunks. 
-While Spark attempted to load-balance the data, it used up only 25 of the 40 
-cores for 750 chunks. In contrast, Nipype tried to use up as many cores as 
-possible. To ensure both engines were processing the same amount of data 
-concurrently, we fixed the amount of cores used to 25.
-Unlike the previous experiment, the D/C ratio was kept static at 174.4MB/s, 
-however, this ratio ensured that different regions of the inequality were 
-reached depending on amount of parallelization. A ratio of 174.4 MB/s satified
-the inequality for local disk at 30 chunks. All other ratios did not satisfy 
-the inequality for Lustre and local disk. Unfortunately, 30 chunks could no be 
-processed using in-memory Spark as its individual chunk size of 3GB exceeded 
-the 2GB partition size limit imposed by Spark.
+chunk size on Big Data strategies. Naturally, greater chunk size signifies 
+a decrease in parallelization, however, it also means an increase in 
+sequential I/O (increased $\Delta/\Gamma$ and $\delta/\gamma$). For 
+this experiment we partitioned the complete BigBrain image into 30, 125 
+and 750 chunks, corresponding to $\gamma$ values of 2, 9 and 25 
+respectively. While Spark attempted to load-balance the data, it used 
+up only 25 of the 40 cores for 750 chunks. In contrast, Nipype tried to 
+use up as many cores as possible. To ensure both engines were 
+processing the same amount of data concurrently (same $\gamma$ value), 
+we fixed the amount of cores used by Nipype to 25. Unlike the previous 
+experiment, the D/C ratio was kept static at 178.6MB/s, however, this 
+ratio ensured that different regions of the inequality were reached 
+depending on amount of parallelization. Unfortunately, 30 chunks could 
+no be processed using in-memory Spark as its individual chunk size of 
+3GB exceeded the 2GB partition size limit imposed by Spark. The number of iterations was
+fixed to 10, and the task duration was adjusted so that C remained constant at 4,400s.
 
@@
 38GB half BigBrain and the 13M T1W MRI image for this experiment. The number of
-chunks was fixed at 125. Similarly to the previous experiment, the total 
-sequential compute time was fixed, however, due to varying size in total data 
-processed, the D/C ratio varied. Once again, we ensured that the D/C ratio fell
-in multiple different regions of the inequality. The D/C ratio ranged from 
-348.8 MB/s for BigBrain and 174.4MB/s for half Bigbrain, to 0.1MB/s for the MRI 
-image. Only the 0.1MB/s MRI satisfied the inequality for both Lustre and local 
+chunks was fixed at 125. Similarly to the previous experiment, the 
+total sequential compute time was fixed (10 iterations, 1.76 seconds 
+per task), however, due to varying size in total data processed (D), 
+the D/C ratio varied. Once again, we ensured that the D/C ratio fell in 
+multiple different regions of the inequality. The D/C ratio ranged from 
+349.1 MB/s for BigBrain and 174.6MB/s for half Bigbrain, to 0.06MB/s for the MRI 
+image. Only the 0.06MB/s MRI satisfied the inequality for both Lustre and local 
 disk.
-
-Since data-locality is not normally preserved in Nipype (a new SLURM allocation 
-is requested for each processed task), we instrumented Nipype to ensure data 
-locality. That is, for each chunk partition, we requested a SLURM allocation to
-process the entire pipeline in parallel, using Nipype's MultiProc scheduler, on 
-a given node. This was possible as no communication was required between the 
-processed chunks.
 
@@
 
-\subsubsection{Execution modes} % perhaps find a better title, it's not informative
-
-We tuned job scheduling to balance the load among
-cluster nodes. This is the default behavior of Spark, but required
-specific instrumentation in Nipype.
-
 \section{Results} % 2 pages
@@
 
-\subsection{Model Evaluations}
-
-In order to evaluate our model, we compared the observed in-memory speedup 
-ratio to our model's expected speed-up ratio for the given 
-D/C / $\delta$/$\gamma$ used (Figure~\ref{fig:modeleval}). Experiments for 
-which there was no in-memory equivalent (i.e. BigBrain split into 30 chunks) 
-were not considered.
-
-%probably confusing to mention values, but they're here for now
-Results show that, overall, the model correctly predicted the effect of page 
-cache on processing times for local disk and Lustre. Points which violated 
-model predictions were found at 1 iteration, where page cache would not have 
-been saturated, due to the amount of data being written, and at a task duration
-of 320 seconds. However, in all cases, the ``1" boundary was never trespassed by 
-more than a factor of 0.19, and is therefore likely a result of system 
-variability.
-
-Lustre's regression slope is smallest of the three filesystems, possibly 
-lending to the fact that Lustre was nearly always in transient state. Out of 
-the thirty four benchmarks taken on Lustre, Lustre was only expected to be in steady 
-state for four of the benchmarks (MRI and all benchmarks with a steady state of
-320). This was validated by our experiments.
-
-The next smallest slope is that of the local disk. The local disk was \todo{continue after disk benchmarks}
-
-Unlike local disk and Lustre, tmpfs exists within the page cache and should not 
-be affected by it in the same way. However, when memory is low, tmpfs will 
-begin to write to the swap space, located on disk. As a result, it is possible 
-to see similar trends to local disk when space is low. 
-Figure~\ref{fig:modeleval} (a) shows that the speed-up between in-memory 
-computing and tmpfs increases as the D/C / $\delta$/$\gamma$ ratio increases. 
-This is unexpected as both tmpfs and in-mem reside in memory and should 
-therefore both have comparable makespans. This was found to occur when the 
-contention to RAM was highest. In other words, when 750 chunks were written to
-disk in 10 iterations with a task duration of 0.59s, Spark in-memory was 
-approximately 2x faster. 
-%why is this? will generate gantts to investigate further.
-
-\subsection{Iterations}
-\begin{figure}[h]
+\begin{figure*}
+\begin{subfigure}{\columnwidth}
     \centering
     \includegraphics[width=\linewidth]{results/figures/iterations.pdf}
-    \caption{Makespans of Spark and Nipype writing to memory, tmpfs, local 
-             disk and Lustre for a given number of iterations. All 
-             makespans use 125 chunks of the complete BigBrain and have a fixed
-             task duration of 3.44 seconds}\label{fig:iterations}
-\end{figure}
+    \caption{Experiment 1: complete BigBrain, 125 chunks, 3.44-second tasks.}\label{fig:iterations}
+\end{subfigure}
+\begin{subfigure}{\columnwidth}
+    \centering
+    \includegraphics[width=\linewidth]{results/figures/cputime.pdf}
+    \caption{Experiment 2: complete BigBrain, 125 chunks, 10 iterations.}\label{fig:cputime}
+\end{subfigure}
+\begin{subfigure}{\columnwidth}
+    \centering
+    \includegraphics[width=\linewidth]{results/figures/numchunks.pdf}
+    \caption{Experiment 3: complete BigBrain, 10 iterations, C= 4,400s.}\label{fig:numchunks}
+\end{subfigure}
+\begin{subfigure}{\columnwidth}
+    \centering
+    \includegraphics[width=\linewidth]{results/figures/datasize.pdf}
+    \caption{Experiment 4: 125 chunks, 10 iterations, 1.76-second tasks.}\label{fig:datasize}
+\end{subfigure}
+\caption{Experiment results: Makespans of Spark and Nipype writing to memory, tmpfs, local  disk and Lustre.}
+\end{figure*}
+\subsection{Experiment 1: Number of Iterations}
 
-Figure~\ref{fig:iterations} shows the difference between the different 
+Fig.~\ref{fig:iterations} shows the difference between the different 
 filesystem choices given the number of iterations. At 1 iteration, all 
-filesystems behave the same, as they are all writing to the page cache. As the 
-total amount of data written by the application increases to 750~GB, there is a 
-greater disparity between Lustre and in-memory (2.67~x slower, on average). 
-Local disk performance, however, 
+filesystems behave the same, although the application was writing faster than the page cache flushed.
+This is explained by the fact that the application wasn't running long enough to saturate the page 
+cache (transient phase). As the total amount of data written by the 
+application increases to 750~GB, there is a greater disparity between 
+Lustre and in-memory (2.67~x slower, on average). Local disk 
+performance, however, 
 is still comparable to memory (1.38~x slower, on average). Despite local disk and 
@@
 
-At 100 iterations, or 7500~GB, Lustre can be found to be, on average, 3.82x 
+At 100 iterations, or 7,500~GB, Lustre can be found to be, on average, 3.82x 
 slower than Spark in-memory. The page cache, on Zenith, occupies 20\% of 
 available memory. Assuming all memory (187~GB) was available, only 37.4~GB of 
-data could be held in a node's page cache at any given time. The slowdown 
-experienced can be explained by the much less data could reside in the page 
-cache at a given time, in contrast to 10 iterations, and therefore the effects 
-of disk bandwidth are more significant in this workflow
+data could be held in a node's page cache at any given time. The 
+slowdown experienced can be explained by the much less data could 
+reside in the page cache at a given time, in contrast to 10 iterations, 
+and therefore the effects of disk bandwidth are more significant in 
+this application. At 100 iterations, the application was writing 500GB 
+per node (7,500 GB / 15 nodes) and hence could not run on tmpfs.
 
-While there is some variability that can be seen in Figure~\ref{fig:iterations} 
+While there is some variability that can be seen in Fig.~\ref{fig:iterations} 
 between the two engines, this believed to be insignificant, and potentially due 
@@
 
-\subsection{CPU Time}
+\subsection{Experiment 2: Task Duration}
 % Comment on the general trend
@@
 %
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=\linewidth]{results/figures/cputime.pdf}
-    \caption{Makespans of Spark and Nipype writing to memory, tmpfs, local 
-             disk and Lustre with varying task durations. All 
-             makespans use 125 chunks of the complete BigBrain and are fixed at 
-             10 iterations}\label{fig:cputime}
-\end{figure}
 
-Increasing CPU time ensured that all filesystems had a comparable performance
-(Figure~\ref{fig:cputime}). Lustre, for instance, is approximately 1.01x slower
+Increasing task duration ensured that all file systems had a comparable performance
+(Fig.~\ref{fig:cputime}). Lustre, for instance, is approximately 1.01x slower
 than Spark in-memory at a task duration of 320 seconds, whereas it is 
 approximately 3.25x slower that Spark in-memory with 2.4 second tasks. This 
-pattern corroborated by our model which postulates that 
+pattern corroborated our page-cache model which postulates that 
 data movement costs will have little impact on compute-intensive tasks. The 
@@
 
-\subsection{Image Block size}
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=\linewidth]{results/figures/numchunks.pdf}
-    \caption{Makespans of Spark and Nipype writing to memory, tmpfs, local 
-             disk and Lustre with varying numbers of chunks. All 
-             makespans have the same overall sequential CPU time and 
-             are fixed at 
-             10 iterations}\label{fig:numchunks}
-\end{figure}
+\subsection{Experiment 3: Image Block Size}
+
 % Comment on the general trend
-As can be seen in Figure~\ref{fig:numchunks}, makespan decreases when the number 
-of chunks increase. This is due to the fact that parallelism increases. With 
+As can be seen in Fig.~\ref{fig:numchunks}, makespan decreases when the number 
+of chunks increases. This is due to the fact that parallelism increases. With 
 30 chunks, only 2 CPUs per node are actively working. At 125 chunks, this 
@@
 
-\subsection{Data Size}
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=\linewidth]{results/figures/datasize.pdf}
-    \caption{Makespans of Spark and Nipype writing to memory, tmpfs, local 
-             disk and Lustre with varying data sizes. All 
-             makespans use 125 chunks and are fixed at 
-             10 iterations with a task duration of 1.76 
-             seconds }\label{fig:datasize}
-\end{figure}
+\subsection{Experiment 4: Image Size}
+
 % Comment on the general trend
 Increasing overall data size decreases performance, as can be seen in 
-Figure~\ref{fig:datasize}. When the data size is very small, as is the case 
-with the MRI image, all filesystem makespans are comparable. This is due to the 
+Fig.~\ref{fig:datasize}. When the data size is very small, as is the case 
+with the MRI image, all file system makespans are comparable. This is due to the 
 fact that page cache can be leveraged fully regardless of file system. However, 
 this time, Spark in-memory performed significantly worse than all other 
-filesystems. Upon further inspection, it appeared that Spark in-memory executed
+filesystems, with a makespan of 2,211 seconds (not reported on the graph to preserve the y scale). Upon further inspection, it appeared that Spark in-memory executed
 in a sequential order, on a single worker node. Lack of parallelism for the MRI 
 image may be a result of Spark's max partition size, which is by default 128~MB
--- significantly larger than the 13~MB MRI image. Due to Spark in-memory's 
-sequential execution, it was omitted from the graph.
+-- significantly larger than the 13~MB MRI image.
 
 At half BigBrain, the makespan differences become apparent in both local disk 
-and Lustre, with Lustre, with Lustre becoming 2.4x slower than in-memory. This 
+and Lustre, with Lustre becoming 2.4x slower than in-memory. This 
 can be attributed to page cache saturation, as predicted by the model for both 
@@
 
+\subsection{Page Cache Model Evaluations}
+\begin{figure*}
+\begin{subfigure}{0.5\linewidth}
+\centering
+    \includegraphics[width=\textwidth]{results/figures/local-incrementation.pdf}
+\caption{Local Disk}
+\label{fig:sub2}
+    \end{subfigure}%
+\begin{subfigure}{0.5\linewidth}
+\centering
+    \includegraphics[width=\textwidth]{results/figures/lustre-incrementation.pdf}
+\caption{Lustre}
+\label{fig:sub3}
+\end{subfigure}
+    \caption{Model evaluation. Grey 
+             regions denote areas that violate model predictions.}
+\label{fig:modeleval}
+\end{figure*}
+
+In order to evaluate the page cache model, we compared the observed 
+speedup ratio provided by in-memory computing to the (D/C) / 
+($\delta$/$\gamma$) and (D/C) / 
+($\Delta$/$\Gamma$) ratios (Fig.~\ref{fig:modeleval}). Experiments for 
+which there was no in-memory equivalent (i.e. BigBrain split into 30 
+chunks) 
+were not considered.
+
+%probably confusing to mention values, but they're here for now
+Results show that, overall, the model correctly predicted the effect of 
+page cache on processing times for local disk and Lustre. That is, the 
+speed-up provided by in-memory computing was larger than 1 for D/C 
+ratios larger than $\delta/\gamma$ (local disk) or $\Delta/\Gamma$ 
+(Lustre). Conversely, the speed-up provided by in-memory computing 
+remained close to 1 for D/C ratios smaller than $\delta/\gamma$ (local 
+disk) or $\Delta/\Gamma$ (Lustre). The two points close to the origin 
+correspond to the sequential processing of the MRI image by Spark 
+mentioned previously.
+
+Points which violated 
+model predictions were found at 1 iteration, where page cache would not 
+have been saturated in spite of a high D/C (transient state). However, 
+in all cases, the ``1" boundary was never trespassed by more than a 
+factor of 0.19, and is therefore likely a result of system variability.
+
 \section{Discussion} % 2 pages
@@
 
-% Effect of Lazy Evaluation?
+\subsection{Effect of Lazy Evaluation}
+
+% Spark local vs nipype local: presumably increases cache hits
 

### 2018-12-13 results/code/d_c_graph.py

@@ -143,34 +143,34 @@ def d_c_fig(bench_dir, makespan_file, out_file):
     plt.figure()
     fit = polyfit(x_disk, y_disk,1)
     fit_fn = poly1d(fit)
-    plt.plot(x_disk, y_disk, 'b+', label="Local Disk")
-    plt.plot(x_disk, fit_fn(x_disk), '--k', label=fit_fn)
-    rect = plt.Rectangle([1, 0], 150, 1, color='gray', edgecolor=None)
+    plt.plot(x_disk, y_disk, 'b+', color="#20948B")
+    #plt.plot(x_disk, fit_fn(x_disk), '--k', label=fit_fn)
+    rect = plt.Rectangle([1, 0], 150, 1, color='#CCCCCC', edgecolor=None)
     plt.gca().add_patch(rect)
-    rect = plt.Rectangle([0, 1], 1, 5, color='gray', edgecolor=None)
+    rect = plt.Rectangle([0, 1], 1, 5, color='#CCCCCC', edgecolor=None)
     plt.gca().add_patch(rect)
-    plt.xlabel("(D/C) / (d(D)elta/g(G)amma)")
+    plt.xlabel("(D/C) / ($\delta/\gamma$)")
     plt.ylabel("Speed-up of Spark in-mem")
     plt.legend()
-    plt.ylim(0)
-    plt.xlim(0)
+    plt.ylim(0, 5.5)
+    plt.xlim(0, 130)
     #plt.show()
     plt.savefig(os.path.join(os.path.dirname(out_file), 
                 'local-{}'.format(os.path.basename(out_file))))
     plt.figure()
     fit = polyfit(x_sfs, y_sfs,1)
     fit_fn = poly1d(fit)
-    plt.plot(x_sfs, y_sfs, 'r+', label='Lustre')
-    plt.plot(x_sfs, fit_fn(x_sfs), '--k', label=fit_fn)
-    rect = plt.Rectangle([1, 0], 250, 1, color='gray', edgecolor=None)
+    plt.plot(x_sfs, y_sfs, 'r+', color="#1a1aff")
+    #plt.plot(x_sfs, fit_fn(x_sfs), '--k', label=fit_fn)
+    rect = plt.Rectangle([1, 0], 250, 1, color='#CCCCCC', edgecolor=None)
     plt.gca().add_patch(rect)
-    rect = plt.Rectangle([0, 1], 1, 5, color='gray', edgecolor=None)
[…]

### 2018-12-13 paper.tex
Started discussion

@@
 \algdef{S}[FOR]{ForEach}[1]{\algorithmicforeach\ #1\ \algorithmicdo}
-\usepackage[bookmarks=false]{hyperref}
+\usepackage{hyperref}
 \hypersetup{colorlinks=true,linkcolor=black,citecolor=blue,filecolor=black,urlcolor=blue}
@@
 \begin{abstract}
-
+We benchmark Big Data processing strategies, namely in-memory 
+computing, data locality, and lazy evaluation, on typical neuroimaging 
+processing use cases represented by the BigBrain dataset. We use Dell 
+EMC's Top-500 cluster, as well as Apache Spark and Nipype as 
+representative processing engines. We model Big Data thresholds by 
+comparing the data/compute ratio of the application to the disk and 
+shared file system bandwidths reported to the number of concurrent 
+processes in a node and across the cluster. This model acknowledges the 
+fact that page caching provided by the Linux kernel is critical to the 
+performance of Big Data applications. Results show that in-memory 
+computing alone speeds-up executions by a factor of up to 1.6, while 
+combined with data locality this factor reaches 5.3. Such important 
+speed-up values are likely to be observed on typical image processing 
+operations performed on images of size larger than 75GB. A ballpark 
+speculation from our model showed that in-memory computing alone would 
+not speed-up current 
+functional MRI analyses, but that it would when coupled with data 
+locality, for a number of concurrently  processed subjects in the order 
+of 200. In addition, we observe that emulating in-memory computing 
+using in-memory file systems (tmpfs) does not reach the performance of 
+an in-memory engine, presumably due to swapping to disk and the lack of data cleanup.
+We conclude that Big Data processing strategies are 
+worth developing for neuroimaging 
+applications. 
 \end{abstract}
@@
 
+\todo{Remove if not included}
+
 % Even if we end up not using simulation, here we should explain
@@
 Processing large images is only considered to be part of the Big Data problem 
-in neuroscience. The other problem is processing large MRI datasets, that is,
-datasets consisting of many small brain images belonging to various different 
-subjects. Although we have not explored explicitly the processing of large MRI
-datasets, the 75GB BigBrain is within the size ballpark~\cite{van2014human} of 
+in neuroscience. The other problem is processing large MRI datasets, 
+that is, datasets consisting of many small brain images belonging to 
+various different subjects. This situation is commonly observed in 
+functional MRI (fMRI), where it is becoming increasingly common to 
+process data from hundreds of subjects. Although we have not explored 
+explicitly the processing of large MRI datasets, the 75GB BigBrain is 
+within the size ballpark~\cite{van2014human} of 
 MRI datasets commonly processed in today's studies.
@@
  $\gamma$ & $\delta/\gamma$ (MB/s) & $\Gamma$ & $\Delta/\Gamma$ (MB/s)&
- \parbox{2cm}{Page-cache on local disk? (Eq.~\ref{eq:page-cache-inequality})} & \parbox{2cm}{Page cache on Lustre? (Eq.~\ref{eq:page-cache-sharedfs})}\\
- \hline
-1   &75    & 430    & 178.6 & 9 & 21.5 & 125 & 4.0 & \cellcolor{red!20} No & \cellcolor{red!20} No\\
-10  & 750   & 4,300 & 178.6 & 9 & 21.5 & 125 & 4.0 & \cellcolor{red!20} No & \cellcolor{red!20} No\\
-100 & 7,500 & 43,000& 178.6 & 9 & 21.5 & 125 & 4.0 & \cellcolor{red!20} No & \cellcolor{red!20} No\\
+ (D/C)/($\delta/\gamma$) & (D/C)/($\Delta/\Gamma$)\\
+ \hline
+1   &75    & 430    & 178.6 & 9 & 21.5 & 125 & 4.0 & \cellcolor{red!20} 8.3 & \cellcolor{red!20} 44.7\\
+10  & 750   & 4,300 & 178.6 & 9 & 21.5 & 125 & 4.0 & \cellcolor{red!20} 8.3 & \cellcolor{red!20} 44.7\\
+100 & 7,500 & 43,000& 178.6 & 9 & 21.5 & 125 & 4.0 & \cellcolor{red!20} 8.3 & \cellcolor{red!20} 44.7\\
 \hline
@@
  $\gamma$ & $\delta/\gamma$ (MB/s) & $\Gamma$ & $\Delta/\Gamma$ (MB/s)&
- \parbox{2cm}{Page-cache on local disk? (Eq.~\ref{eq:page-cache-inequality})} & \parbox{2cm}{Page cache on Lustre? (Eq.~\ref{eq:page-cache-sharedfs})}\\
- \hline
- 2.4  & 750 & 3,000   & 256   & 9 & 21.5 & 125 & 4.0 &  \cellcolor{red!20} No & \cellcolor{red!20} No   \\
- 3.44 & 750 & 4,300   & 178.6 & 9 & 21.5 & 125 & 4.0 &  \cellcolor{red!20} No & \cellcolor{red!20} No   \\
- 7.68 & 750 & 9,600   & 80    & 9 & 21.5 & 125 & 4.0 &  \cellcolor{red!20} No & \cellcolor{red!20} No   \\
- 320  & 750 & 400,000 & 1.9   & 9 & 21.5 & 125 & 4.0 &  \cellcolor{green!20} Yes & \cellcolor{green!20} Yes  \\
+ (D/C)/($\delta/\gamma$) & (D/C)/($\Delta/\Gamma$)\\
+ \hline
+ 2.4  & 750 & 3,000   & 256   & 9 & 21.5 & 125 & 4.0 &  \cellcolor{red!20} 11.9 & \cellcolor{red!20} 64   \\
+ 3.44 & 750 & 4,300   & 178.6 & 9 & 21.5 & 125 & 4.0 &  \cellcolor{red!20} 8.3 & \cellcolor{red!20} 44.7   \\
+ 7.68 & 750 & 9,600   & 80    & 9 & 21.5 & 125 & 4.0 &  \cellcolor{red!20} 3.7 & \cellcolor{red!20} 20   \\
+ 320  & 750 & 400,000 & 1.9   & 9 & 21.5 & 125 & 4.0 &  \cellcolor{green!20} 0.09 & \cellcolor{green!20} 0.48  \\
  \hline
@@
  $\gamma$ & $\delta/\gamma$ (MB/s) & $\Gamma$ & $\Delta/\Gamma$ (MB/s)&
- \parbox{2cm}{Page-cache on local disk? (Eq.~\ref{eq:page-cache-inequality})} & \parbox{2cm}{Page cache on Lustre? (Eq.~\ref{eq:page-cache-sharedfs})}\\
- \hline
- 30  & 750 & 4,400   & 174.6 & 2  & 96.8 & 30  & 16.8 &  \cellcolor{red!20} No & \cellcolor{red!20} No   \\
- 125 & 750 & 4,400   & 174.6 & 9  & 21.5 & 125 & 4.0 &  \cellcolor{red!20} No & \cellcolor{red!20} No   \\
- 750 & 750 & 4,400   & 174.6 & 25 & 7.7  & 375 & 1.3 &  \cellcolor{red!20} No & \cellcolor{red!20} No  \\
+ (D/C)/($\delta/\gamma$) & (D/C)/($\Delta/\Gamma$)\\
+ \hline
+ 30  & 750 & 4,400   & 174.6 & 2  & 96.8 & 30  & 16.8 &  \cellcolor{red!20} 1.8 & \cellcolor{red!20} 10.4   \\
+ 125 & 750 & 4,400   & 174.6 & 9  & 21.5 & 125 & 4.0 &  \cellcolor{red!20} 8.1 & \cellcolor{red!20} 43.7  \\
+ 750 & 750 & 4,400   & 174.6 & 25 & 7.7  & 375 & 1.3 &  \cellcolor{red!20} 22.7 & \cellcolor{red!20} 134.3  \\
 
@@
  $\gamma$ & $\delta/\gamma$ (MB/s) & $\Gamma$ & $\Delta/\Gamma$ (MB/s)&
- \parbox{2cm}{Page-cache on local disk? (Eq.~\ref{eq:page-cache-inequality})} & \parbox{2cm}{Page cache on Lustre? (Eq.~\ref{eq:page-cache-sharedfs})}\\
- \hline
- BigBrain      & 750   & 2,200   & 349.1     & 9  & 21.5  & 125 & 4.0 &  \cellcolor{red!20} No & \cellcolor{red!20} No   \\
- Half BigBrain & 375   & 2,200   & 174.6     & 9  & 21.5  & 125 & 4.0 &  \cellcolor{red!20} No & \cellcolor{red!20} No   \\
- MRI           & 0.127 & 2,200   & 0.06 & 9& 21.5  & 125 & 4.0 &  \cellcolor{green!20} Yes & \cellcolor{green!20} Yes  
+(D/C)/($\delta/\gamma$) & (D/C)/($\Delta/\Gamma$)\\
+ \hline
+ BigBrain      & 750   & 2,200   & 349.1     & 9  & 21.5  & 125 & 4.0 &  \cellcolor{red!20} 16.2 & \cellcolor{red!20} 87.3   \\
+ Half BigBrain & 375   & 2,200   & 174.6     & 9  & 21.5  & 125 & 4.0 &  \cellcolor{red!20} 8.1 & \cellcolor{red!20} 43.7   \\
+ MRI           & 0.127 & 2,200   & 0.06      & 9  & 21.5  & 125 & 4.0 &  \cellcolor{green!20} 0.003 & \cellcolor{green!20} 0.015  
 \end{tabular}
-\caption{Experiment conditions}
+\caption{Experiment conditions. Red cells denote the conditions where 
+the inequalities in Equations~\ref{eq:page-cache-inequality} 
+and~\ref{eq:page-cache-sharedfs} do not hold, i.e., the page cache is 
+asymptotically useless. Green cells show the conditions where the page cache covers all data writes.}
 \label{table:experiments}
@@
 \caption{Experiment results: Makespans of Spark and Nipype writing to memory, tmpfs, local  disk and Lustre.}
+\label{fig:results}
 \end{figure*}
@@
 \caption{Local Disk}
-\label{fig:sub2}
+\label{fig:modeleval-local}
     \end{subfigure}%
@@
 \caption{Lustre}
-\label{fig:sub3}
+\label{fig:modeleval-lustre}
 \end{subfigure}
@@
 speedup ratio provided by in-memory computing to the (D/C) / 
-($\delta$/$\gamma$) and (D/C) / 
-($\Delta$/$\Gamma$) ratios (Fig.~\ref{fig:modeleval}). Experiments for 
-which there was no in-memory equivalent (i.e. BigBrain split into 30 
+($\delta$/$\gamma$) and (D/C) / ($\Delta$/$\Gamma$) ratios 
+(Fig.~\ref{fig:modeleval}). Speed-up ratios were computed as the ratio 
+between the makespan obtained with Spark on local disk or Lustre, and 
+the makespan obtained with Spark for in-memory computing. Experiments 
+for which there was no in-memory equivalent (i.e. BigBrain split into 
+30 
 chunks) 
@@
 \label{sec:discussion}
+
 \subsection{Effect of In-Memory Computing}
 % Shared fs vs local disk vs in-memory: what do we gain?
-
-\subsection{Effect of Data Locality}
-% It should be important, in particular when there is contention and when
-% tasks are small.
+We measure the effect of in-memory computing by comparing the runs 
+of Spark in-memory (yellow bars in Fig.~\ref{fig:results}) to the ones 
+of Spark on local disk (non-hatched green bars). The speed-up provided 
+by in-memory computing is also reported in 
+Fig.~\ref{fig:modeleval-local}. The speed-up provided by in-memory 
+computing increases with (D/C) / ($\delta$/$\gamma$), as expected from 
+the model. In our experiments, it peaked at 1.6, for a ratio of 16.2. 
+This correspond to the processing of the BigBrain with 125 chunks and 
+1.76-second tasks in experiment 4 (total computing time C=2,200s), 
+which is typically encountered in common image processing tasks such as 
+denoising, intensity normalization, etc. The speed up of 1.6 is also 
+reached with a ratio of 22.7 in experiment 3, obtained by processing 
+the BigBrain with 750 chunks.
+
+The results also allow us to speculate on the effect of in-memory 
+computing on the pre-processing of functional MRI, another typical use 
+case in neuroimaging. Assuming an average processing time of 20 minutes 
+per subject, which is a ballpark value commonly observed with the 
+popular SPM or FSL packages \todo{add refs}, an input data size of 100MB per subject, 
+and an output data size of 2GB (20-fold increase compared to input 
+size), the D/C ratio would be 1.8MB/s, which would reach the 
+$\delta/\gamma$ threshold measured on this cluster for $\gamma=108$, 
+that is, if 108 subjects were processed on the same node. This is very 
+unlikely as the number of CPUs per node is 40. We therefore conclude 
+that in-memory computing is likely to be useless for fMRI analysis.
+
+%~ \subsection{Effect of Data Locality}
+%~ % It should be important, in particular when there is contention and when
+%~ % tasks are small. 
+%~ We measure the effect of data locality by comparing the runs of Spark 
+%~ on local disk (non-hatched green bars in Fig.~\ref{fig:results}) to the 
+%~ ones of Spark on Lustre (non-hatched blue bars). The speed-up provided by 
+%~ local execution peaks at 3.2, reached at 750 chunks in experiment 3. \todo{This needs to be finished or removed.}
+
+\subsection{Combined Effect of In-Memory and Data Locality}
+We measure the combined effect of data locality and in-memory computing 
+by comparing the runs of Spark in-memory (yellow bars in 
+Fig.~\ref{fig:results}) to the ones of Spark on Lustre (non-hatched 
+blue bars). The speed-up provided by the combine use of data locality 
+and in-memory computing is also reported in 
+Fig.~\ref{fig:modeleval-lustre}. The provided speed-up increases with 
+(D/C) / ($\Delta$/$\Gamma$), as expected from the model. In our 
+experiments, it peaked around 5, for ratios of 120.4 and 64. Again, 
+this configuration is likely to happen in typical image processing 
+tasks performed on the BigBrain.
+
+As for the fMRI speculation, the D/C 
+ratio of 1.8MB/s would reach the $\Delta/\Gamma$ threshold for 
+$\Gamma=280$, which is a realistic number of subjects to process on a 
+complete cluster. Naturally, this estimate is highly dependent on the observed bandwidth
+of the shared file system ($\Delta$).
 

### 2018-12-13 results/code/d_c_graph.py

@@ -52,9 +52,14 @@ def d_c_fig(bench_dir, makespan_file, out_file):
         750: 350
     }
     
+    file_names = []
     benches = {}
     for file_name in os.listdir(bench_dir):
         it, data, fs, wf, delay, blocks = parse_bench_file_name(file_name)
+        if wf != 'sp': # remove nipype runs from the comparison
+            print("Skipping {}".format(file_name))
+            continue
+        file_names.append(file_name)
         benches[file_name] = {
             'iterations': it,
             'data_file': data,
@@ -75,11 +80,12 @@ def d_c_fig(bench_dir, makespan_file, out_file):
             splits = line.split(' ')
             file_name = splits[0].split(':')[0]
             makespan = float(splits[2])
-            benches[file_name]['makespan'] = makespan
+            if file_name in file_names:
+                benches[file_name]['makespan'] = makespan
     
     
     # Set memory speed-ups
-    for file_name in os.listdir(bench_dir):
+    for file_name in file_names:
         b = benches[file_name]
         in_mem_file = "sp_mem_{}it{}{}_{}delay.out".format(b['iterations'],
                                                            b['blocks'],
@@ -99,7 +105,7 @@ def d_c_fig(bench_dir, makespan_file, out_file):
     y_disk = []
     x_sfs = []
     y_sfs = []
-    for file_name in os.listdir(sys.argv[1]):
+    for file_name in file_names:
         if benches[file_name].get('memory-speed-up') == None:
             continue # file isn't in benchmark
         fs = benches[file_name]['file_system']
@@ -159,7 +165,7 @@ def d_c_fig(bench_dir, makespan_file, out_file):
                 'local-{}'.format(os.path.basename(out_file))))
     plt.figure()
     fit = polyfit(x_sfs, y_sfs,1)
-    fit_fn = poly1d(fit)
+    fit_fn = poly1d(fit)   
     plt.plot(x_sfs, y_sfs, 'r+', color="#1a1aff")
     #plt.plot(x_sfs, fit_fn(x_sfs), '--k', label=fit_fn)
     rect = plt.Rectangle([1, 0], 250, 1, color='#CCCCCC', edgecolor=None)

### 2018-12-13 paper.tex
Added @shots47s

@@
   \IEEEauthorblockN{
-    Val\'erie Hayot-Sasson and 
-    Tristan Glatard
+    Val\'erie Hayot-Sasson$^1$, Shawn T Brown$^2$ and 
+    Tristan Glatard$^1$
   }\\
   \IEEEauthorblockA{
-    Department of Computer Science and Software Engineering, Concordia University, Montreal, Qu\'ebec, Canada
+    $^1$Department of Computer Science and Software Engineering, Concordia University, Montreal, Canada\\
+    $^2$Montreal Neurological Institute, McGill University, Montreal, Canada
   }

### 2018-12-13 paper.tex
Made several edits to the manuscript for grammar and language.  Overall looks pretty good.

@@
 \begin{abstract}
-We benchmark Big Data processing strategies, namely in-memory 
-computing, data locality, and lazy evaluation, on typical neuroimaging 
-processing use cases represented by the BigBrain dataset. We use Dell 
+We benchmark in-memory, computing, data locality and lazy Big Data processing 
+strategies on typical neuroimaging processing use cases represented by the BigBrain dataset. We use Dell 
 EMC's Top-500 cluster, as well as Apache Spark and Nipype as 
@@
 
-Big Data processing engines reduce data-related overheads, through 
+Big Data processing engines reduce data-related overheads through 
 use of concepts such as data locality, in-memory computing, and lazy evaluation.
@@
 Lazy evaluation, also used in Spark, means that results are computed only 
-when necessary, enabling further performance optimizations. Frameworks 
+when necessary to enable further performance optimizations. Frameworks 
 such as
@@
 scientific engines. While Big Data applications mostly target text 
-processing (e.g. Web search, frequent pattern mining, recommender 
+processing (e.g. web search, frequent pattern mining, recommender 
 systems~\cite{leskovec2014mining}) implemented in consistent software 
@@
 
-Neuroimaging, our primary field of interest, is no exception to the 
-generalized rise of data volumes in science, due to the joint increase 
+Our primary field of interest, Neuroimaging, is no exception to the 
+generalized rise of data volumes in science due to the joint increase 
 of image resolution and subject cohort sizes~\cite{van2014human}. 
@@
 locality. On the other hand, neuroimaging data processing engines are 
-still being developed, and it is legitimate to wonder whether these 
+still being developed, and the question remains whether these 
 projects should just be migrated to Spark, Dask, or other Big Data 
@@
 
-Our study focuses on performance. We intentionally do not intend to 
+Our study focuses on performance. We intentionally do not 
 compare Big Data and scientific data processing engines on the grounds 
 of workflow language expressivity, fault-tolerance, provenance capture 
-and representation, portability or reproducibility, which are otherwise 
+and representation, portability or reproducibility which are otherwise 
 critical concerns, addressed for instance in~\cite{samba}. Besides, our 
@@
 scheduler, and that a file system shared among the compute nodes is 
-available. We intentionally leave distributed, 
-HDFS-like file system out, as initiatives to deploy them in HPC 
+available. We intentionally did not consider distributed, 
+HDFS-like file system, as initiatives to deploy them in HPC 
 centers, for instance Hadoop on-demand~\cite{krishnan2011myhadoop}, have not 
@@
 Our methods, including performance models, processing engines, 
-applications and infrastructure used are described in 
+applications and infrastructure used, are described in 
 Section~\ref{sec:methods}. Section~\ref{sec:results} presents our 
@@
 Two types of operations can be performed on RDDs:
-transformations and actions. Applying a transformation to an RDD produces a new,
-child, RDD through a narrow or wide dependency. A narrow dependency signifies 
+transformations and actions. Applying a transformation to an RDD produces a new
+child RDD through a narrow or wide dependency. A narrow dependency signifies 
 that the child is only dependent on a single parent partition, whereas a child 
@@
 scheduler is the default scheduler. YARN was a scheduler designed for Hadoop~\cite{white2012hadoop}
-clusters. It is prepackaged with Hadoop installations. In contrast to YARN, 
+clusters and is prepackaged with Hadoop installations. In contrast to YARN, 
 Mesos was designed to be used in multitenant cluster environments. In our experiments,
@@
 a child variant of the Node, can execute on multiple inputs. All tasks in
-Nipype execute in their own uniquely named subdirectory which facilitates the provenance 
-tracking of inputs and outputs. This also enables checkpointing of the workflow.
+Nipype execute in their own uniquely named subdirectory which facilitates provenance 
+tracking of inputs and outputs, and also enables checkpointing of the workflow.
 In the case of Node failure or application modification, only the nodes who have
@@
  and simply writing to an in-memory file system, such as tmpfs, is 
- related to what happens when the processed data fills up the available 
+ what happens when the processed data fills up the available 
  memory. In case engine memory is used, the engine must cleanup 
@@
 application, but only on the max number of concurrent processes 
-executing on the same node ($\gamma$). We realize that these 
-assumptions would usually be violated in practice, but this simple 
+executing on the same node ($\gamma$). While these 
+assumptions would usually be violated in practice, this simple 
 model already provides interesting insights on the performance of disk 
@@
 asymptotically useless. It should be noted that the transient phase 
-during the page cache fills up might last a significant amount of time, 
-in particular when $\dot d$ is positive but small. We intentionally do not model the 
-transient phase as it requires detailed knowledge on the size of the 
-page cache and the initial amount 
-of data in it ($d_0$), which are difficult to estimate.
+during which the page cache fills up might last a significant amount of time, 
+in particular when $\dot d$ is positive and small. We intentionally do not model the 
+transient phase as it requires detailed knowledge of difficult to estimate parameters  
+such as the page cache and the initial amount 
+of data in it ($d_0$).
 
@@
  SLURM. For the Spark experiments, a Spark cluster was started atop our 
- SLURM allocation. We had access to 16 dedicated nodes. Each compute 
+ SLURM allocation through access to 16 dedicated nodes. Each compute 
  node ran on RHEL 7.4, was 
@@
 histological image of a 65-year-old male's brain. The BigBrain was selected due 
-to being the only one of its kind, as there does not yet exist a 
+to being the only one of its kind, as there does not yet exist any other
 higher-resolution image of a human brain. Moreover, there currently exists a 
 lack of standardized tools for processing the BigBrain as a consequence 
-of its size. To examine the effects processing BigBrain has on the Big Data 
+of its size. To examine the effects processing the BigBrain has on Big Data 
 strategies, we partitioned the full $3845\times3015\times3470$ voxel image into 30 ($5\times3\times2$)

### 2018-12-14 paper.tex

@@
 \begin{abstract}
-We benchmark in-memory, computing, data locality and lazy Big Data processing 
-strategies on typical neuroimaging processing use cases represented by the BigBrain dataset. We use Dell 
-EMC's Top-500 cluster, as well as Apache Spark and Nipype as 
-representative processing engines. We model Big Data thresholds by 
-comparing the data/compute ratio of the application to the disk and 
-shared file system bandwidths reported to the number of concurrent 
-processes in a node and across the cluster. This model acknowledges the 
+    Neuroimaging datasets are rapidly growing in size as a result of 
+    advancements in image acquisition methods, open-science and data sharing. 
+    However, it remains that current neuroimaging data processing engines do 
+    not employ strategies adopted by Big Data processing engines to improve 
+    Big Data processing time. Here, we 
+    evaluate three Big Data processing stategies (in-memory computing, 
+    data-locality and lazy-evaluation) on typical neuroimaging use 
+    cases, represented by the BigBrain dataset. We contrast the various 
+    processing strategies by using Apache Spark and Nipype, as our 
+    representative Big Data and neuroimaging processing engines, on Dell EMC's 
+    Top-500 cluster. 
+    Big Data thresholds were modeled by comparing the data/compute ratio of the 
+    application to the filesystem/concurrent processes ratio. 
+    This model acknowledges the 
 fact that page caching provided by the Linux kernel is critical to the 
 performance of Big Data applications. Results show that in-memory 
-computing alone speeds-up executions by a factor of up to 1.6, while 
-combined with data locality this factor reaches 5.3. Such important 
+computing alone speeds-up executions by a factor of up to 1.6, whereas 
+when combined with data locality, this factor reaches 5.3. Such important 
 speed-up values are likely to be observed on typical image processing 
 operations performed on images of size larger than 75GB. A ballpark 
-speculation from our model showed that in-memory computing alone would 
+speculation from our model showed that in-memory computing alone will
 not speed-up current 
-functional MRI analyses, but that it would when coupled with data 
-locality, for a number of concurrently  processed subjects in the order 
-of 200. In addition, we observe that emulating in-memory computing 
+functional MRI analyses unless coupled with data 
+locality, with a number of concurrently processed subject in the  200 ballpark. 
+    In addition, we observe that emulating in-memory computing 
 using in-memory file systems (tmpfs) does not reach the performance of 
@@
 Data locality-aware scheduling, popularized by the MapReduce~\cite{dean2008mapreduce} 
-framework, runs tasks to the nodes holding the data rather than 
+framework, executes tasks in the nodes holding the data rather than 
 performing costly data transfers across the network. In-memory 
@@
 possible, reducing costly disk I/Os between pipeline steps. 
-Lazy evaluation, also used in Spark, means that results are computed only 
-when necessary to enable further performance optimizations. Frameworks 
+Lazy evaluation, also used in Spark, enables further Big Data performance 
+optimizations, by evaluating results only when necessary. Frameworks 
 such as
@@
 the Big Data era due to increasing data volumes and variety. 
-However, the adoption of Big Data engines for scientific data analysis 
+Nevertheless, the adoption of Big Data engines for scientific data analysis 
 remains limited, perhaps due to the widespread availability of 
@@
 
-Scientific applications differ from the typical Big Data use 
+Scientific applications differ from typical Big Data use 
 cases, which might explain the remaining gap between Big Data and 
@@
 libraries, scientific applications often involve 
-binary data such as images and signals, processed by command-line tools 
+binary data such as images and signals, processed by a sequence of 
+command-line/containerized tools
 using a mix of programming languages (C, Fortran, Python, shell 
-scripts). Infrastructure-wise, Big Data applications commonly run on 
+scripts), referred to as workflows or pipelines. Infrastructure-wise, 
+Big Data applications commonly run on 
 clouds or dedicated commodity clusters with locality-aware file systems 
@@
 explain the lack of adoption of Big Data engines for scientific 
-applications.
-In-memory computing would then be a feature provided by 
+applications. In such configurations, in-memory computing would be a feature 
+provided by 
 the operating system rather than by the engine itself. The frontier 
@@
 data writes and movement. Task scheduling will be part of our 
-discussion though.
+discussion, however.
 
@@
 
-\todo{Perhaps talk about modeling and simulation here.}
+%\todo{Perhaps talk about modeling and simulation here.}
 
@@
 fine-grained transformations, only coarse-grained transformations (applying to
-many elements in the RDD), can be applied, thereby making it simple to keep a 
-log of how the data was modified. This log, known as the lineage, is used
+many elements in the RDD), can be applied, thereby making it simple to maintain a 
+log of data modifications. This log, known as the lineage, is used
 to reproduce any lost data.
@@
 Spark transformations include map, filter and join. To materialize an RDD, an
-action must be performed, such as a reduce or a collect. \todo{Lazy evaluation should be mentioned here, or elsewhere in the paragraph.
-It is currently not described.}
+action must be performed, such as a reduce or a collect. Lazy evaluation is 
+represented in Spark through the use of transformations and actions. A series of
+transformations may be defined without the data ever being materialized. Using 
+this strategy, Spark can make optimizations to ensure optimal processing speeds 
+throughout the application.
+
+%\todo{Lazy evaluation should be mentioned here, or elsewhere in the paragraph.
+%It is currently not described.}
 
@@
 operation. Every shuffle begins with each map task saving its data to local
-files to be fetched by the shuffle operation. The shuffle operation then 
+files for fault tolerance. The shuffle operation then 
 redistributes the data across the partitions as requested. A shuffle marks a 
-stage boundary in Spark. In other words, a shuffle will not begin until all 
-dependent map tasks \todo{shouldn't it be 'actions'?} have completed.
+stage boundary in Spark. In other words, an reduce operation  will not begin 
+until all 
+dependent map tasks have completed.
 
@@
 policy. If an evicted RDD needs to be reused, Spark will recompute it using the
-lineage data collected. It therefore becomes of great importance to persist to
-memory and/or 
-disk RDDs which will be reused throughout the application, particularly if the
-recomputation is costly. 
+lineage data collected. As demonstrated in~\cite{freeman2014mapping
+}, caching significantly improves processing times of iterative algorithms where 
+RDDs are reused. It can be of even greater importance if the RDD is costly to 
+recompute.
 %mention results in thunder paper
 
-Data locality in Spark is achieved through the  scheduling of task to partitions which have the
+Data locality in Spark is achieved through the scheduling of tasks to partitions 
+which have the
 data loaded in memory. If the data is instead stored on HDFS, the scheduler will
@@
 nodes to the Spark cluster as they are allocated by the underlying 
-schedulers may speedup allocation and overall processing time 
-\todo{Cite Shantenu's paper here}. This, however, is not studied in the 
+schedulers may speedup allocation and overall processing time~\cite{paraskevakos2018pilot}. This, however, is not studied in the 
 current paper.
@@
 
+% i use 'popular' for both...
 Nipype is a popular Python neuroimaging processing engine. It aims at being a 
@@
 plugins to schedulers found in most clusters readily available to researchers, 
-such as SGE, PBS, SLURM and HTCondor\todo{expand acronyms}. It also includes its own scheduler, MultiProc,
+such as the SunGrid Engine (SGE), Portable Batch System (PBS), Slurm and HTCondor\todo{cite}. It also includes its own scheduler, MultiProc,
 for parallel processing on single nodes. Nipype also includes many built-in 
@@
 is required. However, it is still possible to save to a non-shared local filesystem, 
-but it may come at the expense of fault-tolerance \todo{explain why}. Moreover, the user will need to
+but it may come at the expense of fault-tolerance, as a results of being stored in a disk-local fs. Moreover, the user will need to
 ensure that the files are appropriately directed to the nodes that require them as
@@
 
-\subsection{Simulation} % is it included? 0.5 page
-
-\todo{Remove if not included}
+%\subsection{Simulation} % is it included? 0.5 page
+
+% \todo{Remove if not included}
 
@@
  cluster}, a Top-500 machine in the Dell EMC HPC Innovation Lab, running 
- SLURM. For the Spark experiments, a Spark cluster was started atop our 
- SLURM allocation through access to 16 dedicated nodes. Each compute 
+ Slurm. For the Spark experiments, a Spark cluster was started on our 
+ Slurm allocation through access to 16 dedicated nodes. Each compute 
  node ran on RHEL 7.4, was 
@@
 application). Each partitioned chunk was incremented by 1, in parallel, by
-the processing engines. As incrementing the images is a relatively quick 
+the applications. As incrementing the images is a relatively quick 
 process, we added a sleep delay to the incrementation stage in order to 
@@
 Lustre (Spark and Nipype). Should more than a single iteration be requested, 
-the increment chunks would be incremented again and saved to the same storage. 
+the incremented chunks would be incremented again and saved to the same storage. 
 This would repeat until the number of requested iterations had elapsed. In all 
-conditions, the first input chunks would be reads from Lustre and the final 
-output chunks would be saved to Lustre. We chose to perform our initial 
+conditions, the first input chunks and final output chunks would be read from / written to  Lustre and the final. 
+We chose to perform our initial 
 reads and final writes on Lustre as a multi-tenant clusters' compute node's 
@@
 
-We conducted 4 experiments in which we varied (1) the number of 
+We conducted four experiments in which we varied (1) the number of 
 iterations in the application ($n$ in 
@@
 used as compute nodes. Since data-locality is not normally preserved in 
-Nipype (a new SLURM allocation is requested for each processed task), 
-we instrumented Nipype to ensure data locality. That is, for each chunk 
-partition, we requested a SLURM allocation to process the entire 
+Nipype (a new Slurm allocation is requested for each processed task), 
+we instrumented Nipype to ensure data locality. That is, the chunk were split into partitions, and for each partition of 
+chunks, we requested a Slurm allocation to process the entire 
 pipeline in parallel, using Nipype's MultiProc scheduler, on a given 
@@
 chunk size on Big Data strategies. Naturally, greater chunk size signifies 
-a decrease in parallelization, however, it also means an increase in 
+a decrease in parallelization, however, it also signifies an increase in 
 sequential I/O (increased $\Delta/\Gamma$ and $\delta/\gamma$). For 
@@
 up only 25 of the 40 cores for 750 chunks. In contrast, Nipype tried to 
-use up as many cores as possible. To ensure both engines were 
-processing the same amount of data concurrently (same $\gamma$ value), 
-we fixed the amount of cores used by Nipype to 25. Unlike the previous 
+use up as many cores as possible. Unlike the previous 
 experiment, the D/C ratio was kept static at 178.6MB/s, however, this 
@@
 filesystems behave the same, although the application was writing faster than the page cache flushed.
-This is explained by the fact that the application wasn't running long enough to saturate the page 
-cache (transient phase). As the total amount of data written by the 
+This is explained by the fact that the application data was not large enough to saturate the page  
+cache fully (transient phase). As the total amount of data written by the 
 application increases to 750~GB, there is a greater disparity between 
@@
 
-At 100 iterations, or 7,500~GB, Lustre can be found to be, on average, 3.82x 
+At 100 iterations, or 7,500~GB, Lustre can be found to be, on average, 3.82~x 
 slower than Spark in-memory. The page cache, on Zenith, occupies 20\% of 
-available memory. Assuming all memory (187~GB) was available, only 37.4~GB of 
-data could be held in a node's page cache at any given time. The 
-slowdown experienced can be explained by the much less data could 
-reside in the page cache at a given time, in contrast to 10 iterations, 
-and therefore the effects of disk bandwidth are more significant in 
+total memory. Since there was 187~GB of RAM on each node, only 37.4~GB of 
+dirty data could be held in a node's page cache at any given time. The 
+slowdown experienced can be explained by the smaller percentage of total data 
+residing in the page cache at a given time, compared to 10 iterations. 
+Therefore, the effects of Lustre bandwidth are more significant in 
 this application. At 100 iterations, the application was writing 500GB 
-per node (7,500 GB / 15 nodes) and hence could not run on tmpfs.
+per node (7,500 GB / 15 nodes) and hence could not run on tmpfs or local disk.
 
@@
 data movement costs will have little impact on compute-intensive tasks. The 
-reasoning for this is that longer tasks give the page cache more time to flush 
+reasoning behind this is that longer tasks give the page cache more time to flush 
 between write to disk. The longer the delay, the less likely the page cache is 
@@
 As can be seen in Fig.~\ref{fig:numchunks}, makespan decreases when the number 
-of chunks increases. This is due to the fact that parallelism increases. With 
+of chunks increases. This is due to the fact that parallelism increases with a increase in chunk. With 
 30 chunks, only 2 CPUs per node are actively working. At 125 chunks, this 
@@
 node, only 40 can be processed concurrently, thus requiring two sequential 
-incrementation batches, and resulting in similar makespans as 125 
+incrementation batches and resulting in similar makespans as 125 
 chunks.
@@
 
-% Spark local vs nipype local: presumably increases cache hits
+The effects of lazy-evaluation can be seen throughout the experiments. Nipype 
+was found to generally be slower than Spark in all experiments. While a Nipype DAG is 
+generated prior to workflow execution, there are no such optimizations, as in 
+Spark, to ensure that the least amount of work is performed to produce the 
+results.
+
+During the processing of Experiment 3, it was found that Spark split up the data
+into an processed it in two batches for the 750 chunk input. That is to say, 
+each node in the cluster processed 50 chunks, and since there was only a maximum
+of 40 cores available on each node, Spark opted to run to batches of 25 to ensure
+a balanced workload. Rather than running each iteration on the full dataset, as with Nipype, 
+Spark opted to perform all the iterations on the first batch (load, increment, save),
+and then proceeds to process the second batch. Such an optimization is important, 
+even when processing data on disk, as it would presumably increase the occurrance
+of cache hits. This may partially explain the speedup seen at 750 chunks in
+Figure~\ref{fig:numchunks}.% Spark local vs nipype local: presumably increases cache hits
 
@@
 % Can we emulate in-memory computing using write buffers or tmpfs?
+
+Although tmpfs and page cache do improve performance in general, as seen in
+Figure~\ref{fig:results}, they do not always perform equivalently to in-memory. 
+Tmpfs's main limitation is that data residing on it may be swapped to disk if 
+the system's memory usage is high. When it reaches this point, its performance 
+slows down to swap disk bandwidth, as observed in Figure~\ref{fig:datasize}. Page cache suffers a similar dilemma. As detailed
+in~\href{https://access.redhat.com/documentation/en-us/red_hat_enterprise_linux/6/html/performance_tuning_guide/s-memory-tunables}{RedHatDoc}\todo{add to references},
+I/O blocking write to disk occurs when a giving percentage (e.g. 20~\%) of total memory is occupied by dirty data, once this 
+threshold is exceeded, processes wanting to execute writes must wait until the dirty data
+in the cache is flushed to disk. Furthermore, like memory, tmpfs and page cache are shared resources on a node.
+If users on the node are heavily using memory to incite tmpfs to write to swap space, or are performing data-intensive operations
+that fill up swap space, this will limit the performance on tmpfs/page cache of other users. 
+However, with memory, it is possible to request, through the HPC scheduler, a required
+amount of available memory. Ultimately, in-memory data will also need to be spilled to
+disk if memory usage exceeds amount of available memory, although disk writes are 
+likely to occur in tmpfs and page cache before requested available memory is filled.
 
@@
 
+An alternative Nipype plugin designed for running on Slurm was not used in the experiments.
+The Slurm plugin requests a Slurm allocation for each processed data chunk. Such a
+scheduling strategy was not ideal in our environment where oversubscription of nodes was 
+not enabled.
+
+Unlike Spark, Nipype, by default, opts to use all available CPUs rather than to load balance
+data across the cluster. That is, given 50 chunks and 40 cores, Spark will only use up 25
+cores, and will break it up into two batches. Nipype, on the other hand, will also have no 
+choice but to split up the processing into two batches, but will first processed 40 chunks, 
+immediately followed by the remaining 10. While both are reasonable strategies for 
+data processing, Spark may end up benefiting more from the page cache, as less data is 
+written in parallel (25 vs 40), giving more time for the page cache to flush.
+
 \subsection{Other Comments}
 
+Writing to node-local storage in a cluster environment comes at a cost, for both
+Nipype and Spark without HDFS.
 % fault-tolerance: if you loose a node, you loose the data when it's written on disk or in memory.

### 2018-12-14 paper.tex
Minor fixes, added a few todos, and latex comments for conclusion and remaining sections

@@
 functional MRI analyses unless coupled with data 
-locality, with a number of concurrently processed subject in the  200 ballpark. 
+locality, with a number of concurrently processed subject in the  280 ballpark. 
     In addition, we observe that emulating in-memory computing 
@@
 scientific engines. While Big Data applications mostly target text 
-processing (e.g. web search, frequent pattern mining, recommender 
+processing (e.g. Web search, frequent pattern mining, recommender 
 systems~\cite{leskovec2014mining}) implemented in consistent software 
@@
 available. We intentionally did not consider distributed, 
-HDFS-like file system, as initiatives to deploy them in HPC 
+HDFS-like file systems, as initiatives to deploy them in HPC 
 centers, for instance Hadoop on-demand~\cite{krishnan2011myhadoop}, have not 
@@
 
-The pipelines, benchmarks, performance data, and analysis scripts used 
+The application pipelines, benchmarks, performance data, and analysis scripts used 
 to implement the methods described hereafter are all available at 
 \url{https://github.com/big-data-lab-team/paper-in-mem-locality} for 
-further inspection and reproducibility. Links to the pipeline engines 
+further inspection and reproducibility. Links to the processing engines 
 and processed data are provided in the text.
@@
 transformations may be defined without the data ever being materialized. Using 
-this strategy, Spark can make optimizations to ensure optimal processing speeds 
+this strategy, Spark can optimize data processing
 throughout the application.
@@
 policy. If an evicted RDD needs to be reused, Spark will recompute it using the
-lineage data collected. As demonstrated in~\cite{freeman2014mapping
-}, caching significantly improves processing times of iterative algorithms where 
+lineage data collected. As demonstrated in~\cite{freeman2014mapping}, caching significantly improves processing times of iterative algorithms where 
 RDDs are reused. It can be of even greater importance if the RDD is costly to 
@@
 our experiments using the PySpark API. This is at a cost to performance 
-as the PySpark code must undergo Python to Java serialization. We used 
+as the PySpark code must undergo Python to Java serialization. We used Spark
 version 2.3.2 installed from \url{https://spark.apache.org}.
@@
 In order for Nipype to operate as intended, a filesystem shared by all the nodes 
-is required. However, it is still possible to save to a non-shared local filesystem, 
-but it may come at the expense of fault-tolerance, as a results of being stored in a disk-local fs. Moreover, the user will need to
-ensure that the files are appropriately directed to the nodes that require them as
-there is no guarantee of data locality in Nipype.
+is required. However, it is still possible to save to a non-shared 
+local filesystem, but it may come at the expense of fault-tolerance, as 
+a result of being stored in a disk-local file system \todo{I don't get 'as a result of being stored in a disk-local file system'}. Moreover, the 
+user will need to ensure that the files are appropriately directed to 
+the nodes that require them as there is no guarantee of data locality 
+in Nipype.
 
@@
 this happens, the page cache rapidly fills up and writes are limited by 
-the disk write bandwidth -- about 500~MB/s for random writes to 
-contemporary SSDs  -- as if no page cache was involved.
+the disk write bandwidth as if no page cache was involved.
 
@@
 
-As in the local disk model, we note that page caching will only be asymptotically
+As in the local disk model, we note that page caching will only be 
 useful when the flush bandwidth is greater than the write throughput of 
@@
 \end{equation}
-where $\Gamma$ is the max number of concurrent processes in the cluster. 
+where $\Gamma$ is the max number of concurrent processes \emph{in the cluster}. 
 Note that $\frac{\Delta}{\Gamma}$ will usually be much lower than 
@@
  All experiments were executed on 
- \href{https://www.dellemc.com/resources/en-us/asset/sales-documents/products/storage/h16221-hpc-lab-brochure.pdf}{Dell EMC's Zenith 
- cluster}, a Top-500 machine in the Dell EMC HPC Innovation Lab, running 
- Slurm. For the Spark experiments, a Spark cluster was started on our 
+ \href{https://www.dellemc.com/resources/en-us/asset/sales-documents/products/storage/h16221-hpc-lab-brochure.pdf}{Dell 
+ EMC's Zenith cluster}, a Top-500 machine in the 
+ \href{https://www.dellemc.com/en-us/solutions/high-performance-computing/HPC-AI-Innovation-Lab.htm}{Dell 
+ EMC HPC and AI Innovation Lab}, running Slurm. For the Spark 
+ experiments, a Spark cluster was started on our 
  Slurm allocation through access to 16 dedicated nodes. Each compute 
@@
     \EndFor
-    \State write $chunk$ to Lustre
+    \State save $chunk$ to Lustre
     \EndFor
@@
 This would repeat until the number of requested iterations had elapsed. In all 
-conditions, the first input chunks and final output chunks would be read from / written to  Lustre and the final. 
+conditions, the first input chunks and final output chunks would be read from / written to  Lustre. 
 We chose to perform our initial 
@@
 Nipype (a new Slurm allocation is requested for each processed task), 
-we instrumented Nipype to ensure data locality. That is, the chunk were split into partitions, and for each partition of 
+we instrumented Nipype to ensure data locality \todo{You could refer to the script doing that in the paper repo.}. That is, the chunk were split into partitions, and for each partition of 
 chunks, we requested a Slurm allocation to process the entire 
@@
 no be processed using in-memory Spark as its individual chunk size of 
-3GB exceeded the 2GB partition size limit imposed by Spark. The number of iterations was
+3GB exceeded the 2GB partition size limit imposed by Spark \todo{You should move this comment to the results or the reader will wonder
+what happened to the yellow bar when looking at figure 1.c.}. The number of iterations was
 fixed to 10, and the task duration was adjusted so that C remained constant at 4,400s.
@@
 filesystem choices given the number of iterations. At 1 iteration, all 
-filesystems behave the same, although the application was writing faster than the page cache flushed.
+filesystems behave the same, although the application was writing faster than the page cache flushed. \todo{Update the graph to the same y scale and 'broken'
+bars, and check that this impression still holds...}
 This is explained by the fact that the application data was not large enough to saturate the page  
@@
 reasoning behind this is that longer tasks give the page cache more time to flush 
-between write to disk. The longer the delay, the less likely the page cache is 
+between disk writes. The longer the delay, the less likely the page cache is 
 to fill up when the concurrent data size written fits within the page cache.
@@
 % Comment on the general trend
-As can be seen in Fig.~\ref{fig:numchunks}, makespan decreases when the number 
-of chunks increases. This is due to the fact that parallelism increases with a increase in chunk. With 
-30 chunks, only 2 CPUs per node are actively working. At 125 chunks, this 
-changes to a maximum of 9 CPUs per node, and at 750 chunks all 40 CPUs can be 
+As can be seen in Fig.~\ref{fig:numchunks}, makespan decreases when the 
+number of chunks increases. This is due to the fact that parallelism 
+increases with an increase in the number of chunks. With 30 chunks, 
+only 2 CPUs per node are actively working. At 125 chunks, this changes 
+to a maximum of 9 CPUs per node, and at 750 chunks all 40 CPUs can be 
 active. It can be noted that no significant speedup can be seen between 125 
-chunks and 750 chunks. This is because out of the 50 chunks processed by each 
+chunks and 750 chunks \todo{This is not true for in-memory, Spark tmpfs and Spark local disk}. This is because out of the 50 chunks processed by each 
 node, only 40 can be processed concurrently, thus requiring two sequential 
@@
 this time, Spark in-memory performed significantly worse than all other 
-filesystems, with a makespan of 2,211 seconds (not reported on the graph to preserve the y scale). Upon further inspection, it appeared that Spark in-memory executed
+filesystems, with a makespan of 2,211 seconds (not reported on the graph to preserve the y scale) \todo{You may want to add it back when you have
+broken bars}. Upon further inspection, it appeared that Spark in-memory executed
 in a sequential order, on a single worker node. Lack of parallelism for the MRI 
@@
 \end{subfigure}
-    \caption{Model evaluation. Grey 
+    \caption{Page cache model evaluation. Grey 
              regions denote areas that violate model predictions.}
@@
 that in-memory computing is likely to be useless for fMRI analysis.
+Naturally, this estimate is strongly dependent on the characteristics of the cluster.
 
@@
 generated prior to workflow execution, there are no such optimizations, as in 
-Spark, to ensure that the least amount of work is performed to produce the 
+Spark, to ensure that the least amount of data transfer is performed to produce the 
 results.
@@
 each node in the cluster processed 50 chunks, and since there was only a maximum
-of 40 cores available on each node, Spark opted to run to batches of 25 to ensure
+of 40 cores available on each node, Spark opted to run \todo{a word is missing here} to batches of 25 to ensure
 a balanced workload. Rather than running each iteration on the full dataset, as with Nipype, 
@@
 and then proceeds to process the second batch. Such an optimization is important, 
-even when processing data on disk, as it would presumably increase the occurrance
+even when processing data on disk, as it would presumably increase the occurrence
 of cache hits. This may partially explain the speedup seen at 750 chunks in
 Figure~\ref{fig:numchunks}.% Spark local vs nipype local: presumably increases cache hits
+\todo{Add as sentence on this paragraph in the abstract.}
 
@@
 threshold is exceeded, processes wanting to execute writes must wait until the dirty data
-in the cache is flushed to disk. Furthermore, like memory, tmpfs and page cache are shared resources on a node.
+in the cache is flushed to disk \todo{Revise this sentence (grammar?)}.
+
+Furthermore, like memory, tmpfs and page cache are shared resources on a node.
 If users on the node are heavily using memory to incite tmpfs to write to swap space, or are performing data-intensive operations
-that fill up swap space, this will limit the performance on tmpfs/page cache of other users. 
+that fill up the page cache, this will limit the performance on tmpfs/page cache of other users. 
 However, with memory, it is possible to request, through the HPC scheduler, a required
@@
 executors to 5, to preserve a good I/O 
-throughput\footnote{\url{http://blog.cloudera.com/blog/2015/03/how-to-tune-your-apache-spark-jobs-part-2}}. 
+throughput\footnote{\url{http://blog.cloudera.com/blog/2015/03/how-to-tune-your-apache-spark-jobs-part-2}}, but the rationale
+for this recommendation is hardly explained. 
 We believe that throughput degradation observed with more than 5 cores 
@@
 cores, and will break it up into two batches. Nipype, on the other hand, will also have no 
-choice but to split up the processing into two batches, but will first processed 40 chunks, 
+choice but to split up the processing into two batches, but will first process 40 chunks, 
 immediately followed by the remaining 10. While both are reasonable strategies for 
@@
 
+% We haven't benchmarked shuffles but we know it's important.
+
 \subsection{Engine-Specific Comparisons}
@@
 \label{sec:conclusion}
-% When is it useful to use a Big Data engine?
-
+
+% Summary of the achievements.
+% When is it useful to use a Big Data engine for neuroinformatics applications?
+
+% Future work:
+% - now that we know that Spark brings performance to neuro applications, make it usable
+% - scheduling in a shared environment.
+% - benchmark a large scale fmri processing to confirm our speculations
+% - try in-memory distributed file systems such as Apache Ignite
+% - workflow-aware cache eviction strategies instead of LRU/n
+% - improve Spark to better use page caches
 %    "```Disk and network I/O, of course, play a part in Spark performance 
@@
 
-% Future work:
-% - scheduling in a shared environment.
-% - workflow-aware cache eviction strategies instead of LRU/n
-
 \section{Acknowledgments}
 
-\todo{Acknowledge Dell and ask them if they want to co-author the paper.}
+We are thankful to the Dell EMC HPC and AI Innovation Lab
+in Roundrock, TX, for providing the infrastructure and high-quality
+technical support that made possible the experiments reported in this paper.
 

### 2018-12-14 paper.tex
Minor edits
rest of discussion and conclusion

--- 2018-12-14 paper.tex
@@
 
-%~ \subsection{Effect of Data Locality}
+\subsection{Effect of Data Locality}
 %~ % It should be important, in particular when there is contention and when
 %~ % tasks are small. 
-%~ We measure the effect of data locality by comparing the runs of Spark 
-%~ on local disk (non-hatched green bars in Fig.~\ref{fig:results}) to the 
-%~ ones of Spark on Lustre (non-hatched blue bars). The speed-up provided by 
-%~ local execution peaks at 3.2, reached at 750 chunks in experiment 3. \todo{This needs to be finished or removed.}
+We measure the effect of data locality by comparing the runs of Spark
+on local disk (non-hatched green bars in Fig.~\ref{fig:results}) to the 
+ones of Spark on Lustre (non-hatched blue bars). The speed-up provided by 
+local execution peaks at 3.2, reached at 750 chunks in experiment 3. Overall, 
+writing locally was usually preferrable to writing to Lustre, as a result of the
+lower contention on local disk. Although it may be true, as stated in~\cite{Anantharayanan2011datalocality},
+network bandwidths generally exceed that of disks, locality remains important 
+as contention to a shared filesystem will be much higher than local disk.\todo{This needs to be finished or removed.}
+The only time in our experiments where it was found that writing locally did not
+have a huge impact over Lustre, was in experiments 1 and 4, at 1 iteration and 
+when processing the MRI image, respectively. In both these scenarios, the Lustre 
+writes did not impact performance as the data was able to be written to page cache
+and flushed to Lustre asynchronously.
 
@@
 
+With Nipype, MapNodes, which apply a given function to each element in a list, 
+were found to be slower than the Node, which apply a function to a single element,
+due to a blocking mechanism. For this reason, we selected to iterate through a 
+series of Nodes in our code.
+
 \subsection{Other Comments}
@@
 Writing to node-local storage in a cluster environment comes at a cost, for both
-Nipype and Spark without HDFS.
+Nipype and Spark without HDFS. When a node is lost, the node-local data is lost
+with it. While Spark will recompute the lost partition automatically using lineage, 
+Nipype will fail for all tasks requiring the data. Spark, however, will fail if 
+RDDs of node-local filenames are shuffled, as the data associated with the filenames
+will not be shuffled with the RDD
+
+Nipype, when executed on Lustre, will checkpoint itself, ensuring resumption from
+last checkpoint. This is particularly important in the case of compute-intensive 
+applications, such as those found in neuroimaging. Spark also provides a checkpointing
+mechanism, however, it requires HDFS.
+
+It is common in neuroimaging applications for users to want access to all intermediate
+data. Such a feature is currently only possible when writing to shared filesystem. 
+It would also not be an option with Spark in-memory. To enable this burst buffers or 
+heteorogeneous storage managers (e.g. Triple-H~\cite{tripleH}) could be used to ensure
+fast processing and that all outputs (including intermediate outputs) will be 
+sent asynchronously to the shared filesystem. \todo{double check triple-H}
+
 % fault-tolerance: if you loose a node, you loose the data when it's written on disk or in memory.
@@
 % i.e., to benefit from data locality while getting your results on Lustre. 
-
 \subsection{Engine-Specific Comparisons}
+
+It was expected that Spark would experience longer processing times, particularly
+with the small datasets, due to Java serializing. This was not found to be the case. 
+Unlike Spark, Nipype performs a more thorough provenance capture, potentially owing to longer
+processing times.
 
@@
 \label{sec:conclusion}
+
+Big Data performance optimization strategies help improve performance of typical 
+neuroimaging applications. Our experiments indicate that overall, in-memory computing 
+enables greater speedups than what can be obtained by using page cache and tmpfs. 
+While page cache and tmpfs do give memory-like performance, they are likely to fill
+up faster than memory, leading to increased performance penalties when compared
+to in-memory computing.
+
+Data-locality was also found to play an important role in 
+application performance. Local disk was found to perform between than the shared
+filesystem, despite having lower bandwidth, due to increased contention on the 
+shared filesystem. As local disk will generally have less contention than a 
+shared filesystem. It is recommended to have the data stored locally. Although, 
+storing data locally may come at a cost to fault tolerance.
+
+Although a more thorough analysis of lazy evaluation remains to be performed, it
+is speculated that this may be the cause of the general performance difference 
+between Spark and Nipype. Furthermore, it is found that lazy evaluation optimizations increase 
+the likelihood of cache hits, thus improving overall performance.
+
+Even though Big Data strategies are beneficial to the processing of Big 
+neuroimaging data, it is estimated that it would require running a functional MRI
+dataset with 200 concurrent subjects for any noticeable impact using our Lustre 
+bandwidth estimate.
+
+To further improve neuroimaging data processing times, scheduling strategies for 
+Big Data application in a multitenant environment need to be investigated. 
+Additionally, workflow-aware cache eviction policies that maximizes page cache 
+usage for the workflows should be investigated.
 % When is it useful to use a Big Data engine?

### 2018-12-14 paper.tex

--- 2018-12-14 paper.tex
@@
 
+It was expected that Spark would experience longer processing times, particularly
+with the small datasets, due to Java serializing. This was not found to be the case. 
+Unlike Spark, Nipype performs a more thorough provenance capture, potentially owing to longer
+processing times.
+
 % Spark vs Nipype
@@
 \label{sec:conclusion}
+
+Big Data performance optimization strategies help improve performance of typical 
+neuroimaging applications. Our experiments indicate that overall, in-memory computing 
+enables greater speedups than what can be obtained by using page cache and tmpfs. 
+While page cache and tmpfs do give memory-like performance, they are likely to fill
+up faster than memory, leading to increased performance penalties when compared
+to in-memory computing.
+
+Data-locality was also found to play an important role in 
+application performance. Local disk was found to perform between than the shared
+filesystem, despite having lower bandwidth, due to increased contention on the 
+shared filesystem. As local disk will generally have less contention than a 
+shared filesystem. It is recommended to have the data stored locally. Although, 
+storing data locally may come at a cost to fault tolerance.
+
+Although a more thorough analysis of lazy evaluation remains to be performed, it
+is speculated that this may be the cause of the general performance difference 
+between Spark and Nipype. Furthermore, it is found that lazy evaluation optimizations increase 
+the likelihood of cache hits, thus improving overall performance.
+
+Even though Big Data strategies are beneficial to the processing of Big 
+neuroimaging data, it is estimated that it would require running a functional MRI
+dataset with 200 concurrent subjects for any noticeable impact using our Lustre 
+bandwidth estimate.
+
+To further improve neuroimaging data processing times, scheduling strategies for 
+Big Data application in a multitenant environment need to be investigated. 
+Additionally, workflow-aware cache eviction policies that maximizes page cache 
+usage for the workflows should be investigated.
 

### 2018-12-14 paper.tex
Final read, added more comments, added a few sentences in conclusion

@@
 
-% Talk about containers somewhere?
-
 \section{Introduction} % 1 page with abstract
@@
 
-%\todo{Lazy evaluation should be mentioned here, or elsewhere in the paragraph.
-%It is currently not described.}
-
 All actions and wide dependencies require a shuffle -- Spark's most costly
@@
 recompute.
-%mention results in thunder paper
 
@@
 
-%% Main concepts
-% pipeline description
-% detailed description relevant to the experiment
-% how shuffle works
-% how data locality is implemented
-% spills to disk when dataset (and derived data) is too large for memory
-% Lazy evaluation
-% persistence (refer to Thunder)
-% multiple schedulers: we focus on Spark Standalone
-% overlay cluster: not considered here as this would be scheduling
-
-%% Technical details
-% data serialization to Java (we use Python as most neuroinformatics do)
-% limitation in size of RDD element, due to java implementation of binary data
-% memory management: don't give too much memory to executors or GC
-%  many places recommend limiting to 5 cores/executor, no one explains why
-% more on tuning?
-
 \subsubsection{Nipype}
 
-% i use 'popular' for both...
+% i use 'popular' for both... widely-used (for Spark)?
 Nipype is a popular Python neuroimaging processing engine. It aims at being a 
@@
 We used Nipype version 1.1.4 installed through the pip package manager.
-
-% pipeline description: map nodes and their limitations
-% general description relevant to the experiment
-% two plugins
-% provenance
-% file system caching
-% memory management (bad?)
-% Nipype only has data locality becausae we use multiproc. Data locality is emulated by manual decisions on scheduling.
 
@@
 Shared File System                & \cellcolor{orange!25}Page Caching  & \cellcolor{red!25}No                
-% data locality is somehow present in shared fs due to page caching again.
 \end{tabular}
@@
 
-% measured disk bandwidth with random I/O, very different from sequential
-
 Storing data on local disks obviously enables data locality as tasks 
@@
 single node that contains their input data.
-% this brings data locality but not at the level of HDFS because replication isn't there and 
-% there is only 1 single option to schedule a task while preserving data locality.
 
@@
 
-% In~\cite{saini2012performance} the Lustre file system is benchmarked on NASA applications.
+% KEEP THIS FOR WHEN WE LOOK AT SIMULATION
 
@@
 
-% Dedicated Compute nodes with SLURM
-% Lustre server
-
-%~ \begin{table}
-%~ \centering
-%~ \begin{tabular}{c|c}
-%~ $\delta/\gamma$ & $\Delta/\Gamma$\\
-%~ \hline
-%~ 164.5~MB/s         & 3.2~MB/s
-%~ \end{tabular}
-%~ \label{table:infrastructure}
-%~ \caption{Bandwidth thresholds on the target cluster using 15 worker nodes and a
-         %~ maximum of 9 CPUs on each node.}
-%~ \end{table}
-
 \subsection{Datasets} % 1/4 of a page (half a col)
 
-% BigBrain
-% Some BIDS dataset for fmriprep
 We used BigBrain~\cite{amunts2013bigbrain}, a 75GB 40$\mu$m isotropic 
@@
 
-% Incrementation (avoid cache effects in binarization)
 In order to investigate how the different strategies impact processing, we 
@@
 
-% Interactive gantt charts?
-
 \subsection{Page Cache Model Evaluations}
@@
 \subsection{Effect of Data Locality}
-%~ % It should be important, in particular when there is contention and when
-%~ % tasks are small. 
+
 We measure the effect of data locality by comparing the runs of Spark
@@
 network bandwidths generally exceed that of disks, locality remains important 
-as contention to a shared filesystem will be much higher than local disk.\todo{This needs to be finished or removed.}
+as contention to a shared filesystem will be much higher than local disk.
 The only time in our experiments where it was found that writing locally did not
@@
 \subsection{Scheduling Remarks}
-% Scheduling: load balance as much as you can.
-% (which isn't what nipype multiproc does)
 
@@
 executors to 5, to preserve a good I/O 
-throughput\footnote{\url{http://blog.cloudera.com/blog/2015/03/how-to-tune-your-apache-spark-jobs-part-2}}, but the rationale
-for this recommendation is hardly explained. 
-We believe that throughput degradation observed with more than 5 cores 
-per executor might be coming from full page caches.
-
-An alternative Nipype plugin designed for running on Slurm was not used in the experiments.
-The Slurm plugin requests a Slurm allocation for each processed data chunk. Such a
-scheduling strategy was not ideal in our environment where oversubscription of nodes was 
-not enabled.
-
-Unlike Spark, Nipype, by default, opts to use all available CPUs rather than to load balance
-data across the cluster. That is, given 50 chunks and 40 cores, Spark will only use up 25
-cores, and will break it up into two batches. Nipype, on the other hand, will also have no 
-choice but to split up the processing into two batches, but will first process 40 chunks, 
-immediately followed by the remaining 10. While both are reasonable strategies for 
-data processing, Spark may end up benefiting more from the page cache, as less data is 
-written in parallel (25 vs 40), giving more time for the page cache to flush.
-
-With Nipype, MapNodes, which apply a given function to each element in a list, 
-were found to be slower than the Node, which apply a function to a single element,
-due to a blocking mechanism. For this reason, we selected to iterate through a 
-series of Nodes in our code.
+throughput\footnote{\url{http://blog.cloudera.com/blog/2015/03/how-to-tune-your-apache-spark-jobs-part-2}}, 
+but the rationale for this recommendation is hardly explained. We 
+believe that throughput degradation observed with more than 5 cores per 
+executor might be coming from full page caches.
+
+Spark currently doesn't include any active management of disk I/Os or 
+page caches. We believe that it would be beneficial to extend it toward 
+this direction, to increase the performance of operations where local 
+storage has to be used, such as disk spills or shuffles. For instance, 
+workflow-aware cache eviction policies that maximizes page cache usage 
+for the workflows could be investigated.
+
+An alternative Nipype plugin designed for running on Slurm was not used 
+in the experiments. The Slurm plugin requests a Slurm allocation for 
+each processed data chunk. Such a scheduling strategy was not ideal in 
+our environment where oversubscription of nodes was not enabled.
+
+Unlike Spark, Nipype, by default, opts to use all available CPUs rather 
+than to load balance data across the cluster. That is, given 50 chunks 
+and 40 cores, Spark will only use up 25 cores, and will break it up 
+into two batches. Nipype, on the other hand, will also have no choice 
+but to split up the processing into two batches, but will first process 
+40 chunks, immediately followed by the remaining 10. While both are 
+reasonable strategies for data processing, Spark may end up benefiting 
+more from the page cache, as less data is written in parallel (25 vs 
+40), giving more time for the page cache to flush.
+
+With Nipype, MapNodes, which apply a given function to each element in 
+a list, were found to be slower than the Node, which apply a function 
+to a single element, due to a blocking mechanism. For this reason, we 
+selected to iterate through a series of Nodes in our code \todo{while 
+MapNodes might have been easier to use?}.
+
+\todo{If time permits, it would be useful to say that we checked that scheduling between 
+both engines was similar, and show two Gantt charts to justify that.}
 
@@
 data. Such a feature is currently only possible when writing to shared filesystem. 
-It would also not be an option with Spark in-memory. To enable this burst buffers or 
+It would also not be an option with Spark in-memory. To enable this, burst buffers or 
 heteorogeneous storage managers (e.g. Triple-H~\cite{tripleH}) could be used to ensure
@@
 sent asynchronously to the shared filesystem. \todo{double check triple-H}
-
-% fault-tolerance: if you loose a node, you loose the data when it's written on disk or in memory.
-%                  then the only solution is to recompute, which spark would do automatically.
-%                  when nipype is run on lustre, it would avoid recomputing existing files, which Spark doesn't do.
-%                   this niype feature is lost when not using lustre.
-
-% Comment on the practical implications of using tmpfs or local disk: we might not get all the intermediary files.
-
-% mention burst buffers or heterogeneous storage managers as potential solutions to write to local disk, 
-% i.e., to benefit from data locality while getting your results on Lustre.
-
-% We haven't benchmarked shuffles but we know it's important.
-
-\subsection{Engine-Specific Comparisons}
 
@@
 
-% Spark vs Nipype
-% Do we see the effect of Java serialization?
-
 \section{Conclusion} % 1 page with refs
@@
 neuroimaging applications. Our experiments indicate that overall, in-memory computing 
-enables greater speedups than what can be obtained by using page cache and tmpfs. 
-While page cache and tmpfs do give memory-like performance, they are likely to fill
-up faster than memory, leading to increased performance penalties when compared
-to in-memory computing.
-
-Data-locality was also found to play an important role in 
-application performance. Local disk was found to perform between than the shared
-filesystem, despite having lower bandwidth, due to increased contention on the 
-shared filesystem. As local disk will generally have less contention than a 
-shared filesystem. It is recommended to have the data stored locally. Although, 
-storing data locally may come at a cost to fault tolerance.
-
-Although a more thorough analysis of lazy evaluation remains to be performed, it
-is speculated that this may be the cause of the general performance difference 
-between Spark and Nipype. Furthermore, it is found that lazy evaluation optimizations increase 
-the likelihood of cache hits, thus improving overall performance.
-
-Even though Big Data strategies are beneficial to the processing of Big 
-neuroimaging data, it is estimated that it would require running a functional MRI
-dataset with 200 concurrent subjects for any noticeable impact using our Lustre 
-bandwidth estimate.
-
-To further improve neuroimaging data processing times, scheduling strategies for 
-Big Data application in a multitenant environment need to be investigated. 
-Additionally, workflow-aware cache eviction policies that maximizes page cache 
-usage for the workflows should be investigated.
-
-% Summary of the achievements.
-% When is it useful to use a Big Data engine for neuroinformatics applications?
-
-% Future work:
-% - now that we know that Spark brings performance to neuro applications, make it usable
-% - scheduling in a shared environment.
-% - benchmark a large scale fmri processing to confirm our speculations
-% - try in-memory distributed file systems such as Apache Ignite
-% - workflow-aware cache eviction strategies instead of LRU/n
-% - improve Spark to better use page caches
-%    "```Disk and network I/O, of course, play a part in Spark performance 
-%    as well, but neither Spark nor YARN currently do anything to actively 
-%    manage them.```"
-%    http://blog.cloudera.com/blog/2015/03/how-to-tune-your-apache-spark-jobs-part-2/
+enables greater speedups than what can be obtained by using page cache 
+and tmpfs. While page cache and tmpfs do give memory-like performance, 
+they are likely to fill up faster than memory, leading to increased 
+performance penalties when compared to in-memory computing. We conclude 
+that extending Big Data processing engines to better support 
+neuroimaging applications, including developing their provenance, 
+fault-tolerance, and reproducibility features, is a worthwhile project.
+
+Data-locality was also found to play an important role in application 
+performance. Local disk was found to perform between than the shared 
+filesystem \todo{between than the shared filesystem: what does it 
+mean?}, despite having lower bandwidth, due to increased contention on 
+the shared filesystem. As local disk will generally have less 
+contention than a shared filesystem. It is recommended to have the data 
+stored locally. Although, storing data locally may come at a cost to 
+fault tolerance.
+
+Although a more thorough analysis of lazy evaluation remains to be 
+performed, it is speculated that this may be the cause of the general 
+performance difference between Spark and Nipype. Furthermore, it is 
+found that lazy evaluation optimizations increase the likelihood of 
+cache hits, thus improving overall performance.
+
+Even though Big Data strategies are beneficial to the processing of 
+large images, it is estimated that it would require running a 
+functional MRI dataset with 280 concurrent subjects for any noticeable 
+impact using our Lustre bandwidth estimate. Benchmarking Spark and 
+Nipype using such a large fMRI dataset would be a relevant follow-up 
+experiment, to test this hypothesis. It would also be useful to evaluate
+other types of applications, such as ones containing data shuffling steps.
+
+Finally, we plan to extend this study by including task scheduling 
+strategies in a multi-tenant environment. We expect to observe 
+important differences between Spark and Nipype, due to Spark's use of 
+overlay scheduling. The impact of other Big Data technologies, such as 
+distributed in-memory file systems~\cite{ignite}, could also be 
+investigated.
 

### 2018-12-14 results/code/gen_figures.py
added broken axis to figures

@@ -85,33 +85,80 @@ def genfig(exp, makef, outf, cond):
             bench_res[key] = (asarray(index), y_labels)
 
         bar_width = 0.1
-        fig, ax = plt.subplots()
-
-        spim = ax.bar(bench_res[('sp', 'mem')][0] - bar_width * 3,
-                      bench_res[('sp', 'mem')][1],
-                      bar_width,color='#F4CC70', label='Spark in-memory')
+        fig, (ax, ax2) = plt.subplots(2, 1, sharex=True)
         
-        sptmpfs = ax.bar(bench_res[('sp', 'tmpfs')][0] - bar_width * 2,
-                         bench_res[('sp', 'tmpfs')][1],
-                         bar_width, color='#DE7A22', label='Spark tmpfs')
-        nptmpfs = ax.bar(bench_res[('np', 'tmpfs')][0] - bar_width,
-                         bench_res[('np', 'tmpfs')][1], bar_width, alpha=0.6,
-                         color='#DE7A22', label='Nipype tmpfs', hatch='/')
-        splocal = ax.bar(bench_res[('sp', 'local')][0],
-                         bench_res[('sp', 'local')][1],
-                         bar_width, color='#20948B', label='Spark local')
-        nplocal = ax.bar(bench_res[('np', 'local')][0] + bar_width,
-                         bench_res[('np', 'local')][1], bar_width, alpha=0.6,
-                         color='#20948B', label='Nipype local', hatch='/')
-        splustre = ax.bar(bench_res[('sp', 'lustre')][0] + bar_width * 2,
-                          bench_res[('sp', 'lustre')][1],
-                          bar_width, color='#1a1aff', label='Spark lustre')
-        nplustre = ax.bar(bench_res[('np', 'lustre')][0] + bar_width * 3,
-                          bench_res[('np', 'lustre')][1], bar_width, alpha=0.6,
-                          color='#1a1aff', label='Nipype lustre', hatch='/')
-        
-        ax.set_ylabel('Makespan (s)')
+        # Nipype bar's design
+        alpha = 0.6
+        hatch = '/'
 
+        spmem_data = (bench_res[('sp', 'mem')][0] - bar_width * 3,
+                      bench_res[('sp', 'mem')][1], bar_width)
+        col = '#F4CC70'
+        lbl = 'Spark - in-memory'
[…]

### 2018-12-14 pipelines/nipype_binarization.py
removed unused files and cleaned up nipype incrementation code

@@ -1,105 +1,0 @@
-from nipype import Workflow, MapNode, Node, Function
-from time import time
-import argparse, os
-import glob
-
-
-def binarize_chunk(chunk, threshold, benchmark, start, output_dir=None):
-    import nibabel as nib
-    import numpy as np
-    import os, socket, uuid
-    from time import time
-
-    start_time = time() - start
-
-    def write_bench(name, start_time, end_time, node, output_dir, filename):
-        
-        benchmark_dir = os.path.join(output_dir, 'benchmarks')
-        os.makedirs(benchmark_dir, exist_ok=True)
-
-        benchmark_file = os.path.join(benchmark_dir, "bench-{}.txt".format(str(uuid.uuid1())))
-        
-        with open(benchmark_file, 'a+') as f:
-            f.write('{0} {1} {2} {3} {4}\n'.format(name, start_time, end_time, node, filename))
-
-    im = nib.load(chunk)
-    data = im.get_data()
-
-    # to avoid returning a blank image
-    if data.max() == 1:
-        threshold = 0
-
-    data = np.where(data > threshold, 1, 0)
-    
-    bin_im = nib.Nifti1Image(data, im.affine)
-
-    bin_file = os.path.join(output_dir, os.path.basename(chunk)) 
-    nib.save(bin_im, bin_file)
-
-    end_time = time() - start
-
-    if benchmark:
-        write_bench('binarize_chunk', start_time, end_time, socket.gethostname(), 
-                output_dir, os.path.basename(chunk))
-
-    return os.path.abspath(bin_file)
-
-def main():
-    parser = argparse.ArgumentParser(description="BigBrain binarization")
-
-    parser.add_argument('bb_dir',type=str, help='The folder containing BigBrain NIfTI images (local fs only)')
-    parser.add_argument('output_dir', type=str, help='the folder to save binarized images to (local fs only)')
-    parser.add_argument('threshold', type=int, help='binarization threshold')
-    parser.add_argument('iterations', type=int, help='number of iterations')
-    parser.add_argument('--benchmark', action='store_true', help='benchmark pipeline')
-
-    args = parser.parse_args()
-
-    start = time()
-    wf = Workflow('bin_bb')
-    wf.base_dir = os.path.dirname(args.output_dir)
-
-    output_dir = os.path.abspath(args.output_dir)
-    os.makedirs(output_dir, exist_ok=True)
-
-    # get all files in directory
-    bb_files = glob.glob(os.path.join(os.path.abspath(args.bb_dir), '*'))
-
[…]

### 2018-12-14 pipelines/nipype_inc.py

@@ -2,7 +2,6 @@
 
 from nipype import Workflow, MapNode, Node, Function
 from time import time
-from math import ceil
 import argparse
 import os
 import shutil
@@ -16,114 +15,132 @@ except Exception as e:
     from thread import get_ident
 
 
-def increment_chunk(chunk, delay, benchmark, benchmark_dir=None, cli=False,
-                    avg=None):
+def write_bench(name, start_time, end_time, node, benchmark_dir, filename,
+                executor):
+    """Write benchmarks to a file.
+
+    Keyword arguments:
+    name -- benchmarked task's name
+    start_time -- task execution start time (seconds)
+    end_time -- task execution end time (seconds)
+    node -- the hostname id of the host that executed the task
+    benchmark_dir -- the directory to save the benchmark file to
+    filename -- the file that was processed by the task
+    executor -- the process id that executed the task
+    """
+    benchmark_file = op.join(benchmark_dir,
+                             "bench-{}.txt".format(str(uuid.uuid1())))
+
+    with open(benchmark_file, 'a+') as f:
+        f.write('{0} {1} {2} {3} {4} {5}\n'.format(name, start_time,
+                                                   end_time,
+                                                   node, filename,
+                                                   executor))
+
+
+def increment_chunk(chunk, delay, benchmark_dir=None, cli=False):
+    """Increment neuroimaging data by 1.
+
+    Keyword arguments:
+    chunk -- image to increment
+    delay -- task duration (sleep time) in seconds
+    benchmark_dir -- directory to save benchmarks (default None)
+    cli -- increment data using a command-line tool (default False)
+
+    Returns: incremented image filename
+    """
     import nibabel as nib
     import os
     import socket
     import subprocess
     import uuid
     from time import time
+    from nipype_inc import write_bench
     try:
         from threading import get_ident
     except Exception as e:
         from thread import get_ident
 
-    def write_bench(name, start_time, end_time, node, benchmark_dir,
-                    filename, executor):
-
-        benchmark_file = os.path.join(benchmark_dir,
-                                      "bench-{}.txt".format(str(uuid.uuid1())))
-
-        with open(benchmark_file, 'a+') as f:
-            f.write('{0} {1} {2} {3} {4} {5}\n'.format(name, start_time,
[…]

### 2018-12-14 pipelines/nipype_kmeans.py

@@ -1,472 +1,0 @@
-from nipype import Workflow, MapNode, Node, Function
-from numpy import iinfo
-from random import seed, sample
-from time import time
-from socket import gethostname
-from math import ceil
-import argparse
-import os
-import glob
-import uuid
-from benchmark import write_bench
-try:
-    from threading import get_ident
-except Exception as e:
-    from thread import get_ident
-
-
-def get_nearest_centroid(img, centroids):
-    import nibabel as nib
-    import pickle
-    from json import dump
-    from os import path as op
-
-    data = nib.load(img).get_data().flatten()
-
-    assignments = {}
-
-    for vox in data:
-        distance = None
-        nearest_c = None
-        nearest_cv = None
-
-        for c in centroids:
-            c_dist = abs(vox - c[1])
-
-            if (distance is None or c_dist < distance
-                    or (c_dist == distance
-                        and ((vox % 2 == 1 and nearest_cv < c[1])
-                             or (vox % 2 == 0 and nearest_cv > c[1])))):
-                distance = c_dist
-                nearest_c = c[0]
-                nearest_cv = c[1]
-
-        vox = str(vox)
-
-        if nearest_c not in assignments:
-            assignments[nearest_c] = { vox: 1 }
-        elif vox not in assignments[nearest_c]:
-            assignments[nearest_c][vox] = 1
-        else:
-            assignments[nearest_c][vox] += 1
-
-    outfiles = []
-
-    try:
-        for k, v in assignments.items():
-            out_name = 'centroid-{}.json'.format(k)
-            with open(out_name, 'a+') as f:
-                dump(v, f)
-
-            outfiles.append((k, op.abspath(out_name)))
-
-    except Exception as e:
-        for k, v in assignments.iteritems():
-            out_name = 'centroid-{}.json'.format(k)
-
-            with open(out_name, 'a+') as f:
-                dump(v, f)
-
-            outfiles.append((k, op.abspath(out_name)))
-
-    return outfiles
-
-def reduceFilesByCentroid(centroid, assignments):
-    from json import load, dump
-    from collections import Counter
-    from os.path import basename, abspath, isfile
-
-    try:
-        a_files = [t[1] for l in assignments for t in l if t[0] == centroid[0]]
-
-    except Exception as e:
-        a_files = [l[1] for l in assignments if l[0] == centroid[0]]
-
-    c_assignments = Counter({})
-    for fn in a_files:
-        if isfile(fn):
-            with open(fn, 'r') as f:
[…]

### 2018-12-14 pipelines/spark_binarization.py

@@ -1,130 +1,0 @@
-from pyspark import SparkContext, SparkConf
-from io import BytesIO
-from time import time
-import os
-import socket
-import uuid
-import numpy as np
-import nibabel as nib
-import argparse
-
-
-def write_bench(name, start_time, end_time, node, output_dir, filename):
-
-    benchmark_dir = os.path.join(output_dir, 'benchmarks')
-    os.makedirs(benchmark_dir, exist_ok=True)
-
-    benchmark_file = os.path.join(
-            benchmark_dir,
-            "bench-{}.txt".format(str(uuid.uuid1()))
-            )
-
-    with open(benchmark_file, 'a+') as f:
-        f.write('{0} {1} {2} {3} {4}\n'.format(name, start_time, end_time,
-                                               node, filename))
-
-
-def read_img(filename, data, benchmark, start, output_dir):
-
-    start_time = time() - start
-
-    # load binary data into Nibabel
-    fh = nib.FileHolder(fileobj=BytesIO(data))
-    im = nib.Nifti1Image.from_file_map({'header': fh, 'image': fh})
-
-    data = im.get_data()
-
-    end_time = time() - start
-
-    bn = os.path.basename(filename)
-
-    if benchmark:
-        write_bench('read_img', start_time, end_time,
-                    socket.gethostname(), output_dir, bn)
-
-    return (filename, data, (im.affine, im.header))
-
-
-def binarize_data(filename, data, metadata, threshold,
-                  benchmark, start, output_dir):
-
-    start_time = time() - start
-
-    # in order to keep original binarization instead
-    # of returning a blank image
-    if data.max() == 1:
-        threshold = 0
-
-    data = np.where(data > threshold, 1, 0)
-
-    end_time = time() - start
-
-    bn = os.path.basename(filename)
-    if benchmark:
-        write_bench('binarize_data', start_time, end_time,
-                    socket.gethostname(), output_dir, bn)
-
-    return (filename, data, metadata)
-
-
-def save_binarized(filename, data, metadata, benchmark, start, output_dir):
-
-    start_time = time() - start
-
-    bn = os.path.basename(filename)
-    im = nib.Nifti1Image(data, metadata[0], header=metadata[1])
-    out_fn = os.path.join(output_dir, 'bin-' + bn)
-    nib.save(im, out_fn)
-
-    end_time = time() - start
-
-    if benchmark:
-        write_bench('save_binarized', start_time, end_time,
[…]

### 2018-12-14 pipelines/spark_kmeans.py

@@ -1,181 +1,0 @@
-#!/usr/bin/env python
-
-from pyspark import SparkContext, SparkConf
-from io import BytesIO
-from os import path as op, makedirs
-from time import time
-import argparse
-import sys
-import nibabel as nib
-import numpy as np
-from socket import gethostname
-import benchmark as bench
-try:
-    from threading import get_ident
-except Exception as e:
-    from thread import get_ident
-
-def get_nearest_centroid(d, c):
-
-    distance = None
-    nearest_c = None
-
-    for centroid in c:
-        c_dist = abs(d[0]-centroid)
-
-        if (distance is None or c_dist < distance
-                or (c_dist == distance
-                    and ((d[0] % 2 == 1 and nearest_c < centroid)
-                         or (d[0] % 2 == 0 and nearest_c > centroid)))):
-            distance = c_dist
-            nearest_c = centroid
-
-    return (nearest_c, (d[0], d[1]))
-
-
-def update_centroids(d):
-    # format = list of tuples
-    # e.g. [(0, 1), (2, 4), (3, 2)]
-    sum_els = float(sum([i[0]*i[1] for i in d]))
-    num_els = sum(i[1] for i in d)
-    updated = sum_els/num_els
-
-    return updated
-
-
-def get_voxels(d):
-
-    # read data into nibabel
-    fh = nib.FileHolder(fileobj=BytesIO(d[1]))
-    im = nib.Nifti1Image.from_file_map({'header': fh, 'image': fh})
-
-    data = im.get_data()
-
-    return data.flatten('F')
-
-
-def save_segmented(d, assignments, out):
-
-    # read data into nibabel
-    fh = nib.FileHolder(fileobj=BytesIO(d[1]))
-    im = nib.Nifti1Image.from_file_map({'header': fh, 'image': fh})
-
-    data = im.get_data()
-
-    assigned_class = [c[0] for c in assignments]
-    
-    for i in range(0, len(assigned_class)):
-        assigned_voxels = [l[0] for c in assignments if c[0] == assigned_class[i] for l in c[1]]#list(set(assignments[i][1]))
-        data[np.where(np.isin(data, assigned_voxels))] = assigned_class[i]
-    im_seg = nib.Nifti1Image(data, im.affine)
-
-    # save segmented image
-    output_file = op.join(out, 'classified-' + op.basename(d[0]))
-    nib.save(im_seg, output_file)
-
[…]

### 2018-12-14 pipelines/sparkprep.py

@@ -1,1399 +1,0 @@
-from pyspark import SparkContext, SparkConf, StorageLevel
-from niworkflows.nipype.interfaces.ants import BrainExtraction, N4BiasFieldCorrection
-from niworkflows.interfaces.registration import RobustMNINormalizationRPT
-from niworkflows.nipype.interfaces import (
-    utility as niu,
-    freesurfer as fs,
-    c3,
-    base,
-    fsl
-)
-from niworkflows.interfaces.fixes import FixHeaderApplyTransforms as ApplyTransforms
-from niworkflows.interfaces.masks import ROIsPlot
-import niworkflows.data as nid
-from fmriprep.interfaces import(
-        DerivativesDataSink, MakeMidthickness, FSInjectBrainExtracted,
-        FSDetectInputs, NormalizeSurf, GiftiNameSource, TemplateDimensions, Conform, Reorient,
-        ConcatAffines, RefineBrainMask, BIDSDataGrabber, BIDSFreeSurferDir, BIDSInfo,
-        SubjectSummary, AboutSummary
-)
-from fmriprep.utils.misc import add_suffix, fix_multi_T1w_source_name
-from fmriprep.utils.bids import collect_participants, collect_data
-from fmriprep.workflows.anatomical import _seg2msks
-from pkg_resources import resource_filename as pkgr
-from collections import namedtuple
-from multiprocessing import cpu_count
-from time import time
-import os, bunch, socket, argparse
-from fmriprep.info import __version__
-import sys, json
-
-# helper functions
-
-def initialize_json(work_dir):
-    benchmark_file = os.path.join(work_dir, 'benchmarks.json')
-    tasks = {}
-    tasks['task'] = []
-
-    with open(benchmark_file, 'w') as f:
-        json.dump(tasks, f)
-
-def write_bench(name, start_time, end_time, node, work_dir, subject='all', run=''):
-    
-    benchmark_dir = os.path.join(work_dir, 'benchmarks')
-    os.makedirs(benchmark_dir, exist_ok=True)
-
-    benchmark_file = os.path.join(benchmark_dir, "bench-{}.txt".format(node))
-    
-    with open(benchmark_file, 'a+') as f:
-        f.write('{0} {1} {2} {3} {4} {5}\n'.format(name, start_time, end_time, node, subject, run))
-
-
-def get_runtime(interface_dir):
-
-    if not os.path.isdir(interface_dir):
-        os.makedirs(interface_dir)
-
[…]

### 2018-12-14 pipelines/nipype_inc.py
fixed pycodestyle

@@ -64,7 +64,7 @@ def increment_chunk(chunk, delay, benchmark_dir=None, cli=False):
     start_time = time()
 
     inc_chunk = os.path.basename(chunk)
-    
+
     # process data directly in code
     if not cli:
         im = nib.load(chunk)
@@ -134,7 +134,6 @@ def save_results(input_img, output_dir, it=0, benchmark_dir=None):
         from threading import get_ident
     except Exception as e:
         from thread import get_ident
-
 
     start = time()
 

### 2018-12-14 benchmark_scripts/nipype_kmeans_template.sh
some cleanup

@@ -1,14 +1,0 @@
-#!/usr/bin/env bash
-#SBATCH --partition requestq       # Queue (Partition in slurm)
-#SBATCH --reservation=spark
-#SBATCH --time=99:00:00
-#SBATCH --nodes=1
-#SBATCH --mem=183G
-#SBATCH --cpus-per-task=40
-#SBATCH --ntasks-per-node=1
-
-source /mnt/lustrefs/spark/env/exp0/bin/activate
-export PATH=/mnt/lustrefs/spark/SOEN691-project/pipelines/:$PATH
-export PYTHONPATH=/mnt/lustrefs/spark/SOEN691-project/pipelines:$PYTHONPATH
-
-

### 2018-12-14 benchmark_scripts/test_write_buffer.sh

@@ -1,23 +1,0 @@
-#!/bin/bash
-
-#rm /local/todelete/*
-
-cdir=/mnt/lustrefs/spark/data/9_chunks/
-chunks=(${cdir}*)
-odir=/local/todelete
-write_data() {
-    for i in `seq 1 10`;
-        do
-            /mnt/lustrefs/spark/SOEN691-project/increment.py ${c} ${odir}/${i} >> bench_wb/${x}_${i}.out
-	    if [ "$i" -le 1 ]; then
-                c=${odir}/${i}/inc-$(basename ${c})
-            else
-                c=${odir}/${i}/$(basename ${c})
-            fi
-        done
-}
-
-for x in `seq 0 8`;
-    do
-        c=${chunks[${x}]} write_data&
-    done

### 2018-12-14 pipelines/increment.py

@@ -12,23 +12,28 @@ except Exception as e:
     from thread import get_ident
 
 
+def increment(fn, outdir, delay, benchmark_file, start_time):
+    """Increment image data by 1
 
-def increment(fn, outdir, delay, benchmark_file, start_time, avg):
+    Keyword arguments:
+    fn -- filename containing image to increment
+    outdir -- output directory to save data to
+    delay -- task duration (sleep time) in seconds
+    benchmark_file -- file to save benchmarks to
+    start_time -- application driver start time
+    """
     print('Incrementing image: ', fn)
 
     start = time.time() - start_time
     im = nib.load(fn)
     end = time.time() - start_time
     print("read time", end)
-  
+
     if benchmark_file is not None:
         write_bench(benchmark_file, "read_file", start, end,
                     socket.gethostname(), op.basename(fn), get_ident())
 
     inc_data = im.get_data() + 1
-
-    if avg is not None:
-        inc_data += nib.load(avg).get_data().astype(inc_data.dtype, copy=False)
 
     im = nib.Nifti1Image(inc_data, affine=im.affine, header=im.header)
 
@@ -50,12 +55,14 @@ def increment(fn, outdir, delay, benchmark_file, start_time, avg):
     time.sleep(delay)
     print('Saved image to: ', out_fn)
 
+
 def write_bench(benchmark_file, name, start_time, end_time, node, filename,
                 executor):
 
     with open(benchmark_file, 'a+') as f:
         f.write('{0} {1} {2} {3} {4} {5}\n'.format(name, start_time, end_time,
                                                    node, filename, executor))
+
 
 def main():
 
@@ -71,8 +78,6 @@ def main():
                         help='start time of the application')
     parser.add_argument('--delay', type=float, default=0,
                         help='task duration time (in s)')
-    parser.add_argument('--avg', type=str,
-                        help='average chunk to be added to incremented chunk')
 
     args = parser.parse_args()
 
@@ -82,7 +87,7 @@ def main():
         pass
 
     increment(args.filename, args.output_dir, args.delay,
-              args.benchmark_file, args.delay, args.avg)
+              args.benchmark_file, args.delay)
 
 
 if __name__ == '__main__':

### 2018-12-14 pipelines/nipype_inc.py

@@ -81,7 +81,7 @@ def increment_chunk(chunk, delay, benchmark_dir=None, cli=False):
     else:
         program = 'increment.py'
 
-        if benchmark:
+        if benchmark_dir is not None:
             benchmark_file = os.path.join(benchmark_dir,
                                           "bench-{}.txt".format(
                                               str(uuid.uuid1())))

### 2018-12-14 pipelines/spark_inc.py

@@ -18,8 +18,23 @@ except Exception as e:
     from thread import get_ident
 
 
-def write_bench(name, start_time, end_time, node, output_dir,
-                filename, executor, benchmark_dir=None, benchmark_file=None):
+def write_bench(name, start_time, end_time, node, filename, executor,
+                benchmark_dir=None, benchmark_file=None):
+    """Write benchmarks to file.
+
+    Keyword arguments:
+    name -- name of task benchmarked
+    start_time -- task execution start time (in seconds)
+    end_time -- task execution end time (in seconds)
+    node -- hostname ID of node which executed task
+    filename -- filename of image task processed.
+    executor -- process ID
+    benchmark_dir -- directory to save benchmarks to. Not required if
+                     file provided
+    benchmark_file -- file to append benchmarks to.
+
+    Returns: filename of benchmark file
+    """
 
     if not benchmark_file:
         assert benchmark_dir, 'benchmark_dir parameter has not been defined.'
@@ -41,8 +56,20 @@ def write_bench(name, start_time, end_time, node, output_dir,
     return benchmark_file
 
 
-def read_img(filename, data, benchmark, start, output_dir, bench_dir=None):
+def read_img(filename, data, benchmark, start, bench_dir=None):
+    """Convert image in-memory binary data into Numpy array.
+    For in-memory Spark only.
 
+    Keyword arguments:
+    filename -- filename for which data belongs to
+    data -- binary image data
+    benchmark -- boolean value that specifies whether to benchmark the task
+    start -- start time of the driver (in seconds)
+    benchmark_dir -- benchmark directory required if benchmarking is requested.
+                     (default None)
+
+    Returns: Tuple containing image data and metadata
+    """
     start_time = time() - start
 
     print('Reading image')
@@ -59,7 +86,7 @@ def read_img(filename, data, benchmark, start, output_dir, bench_dir=None):
     bench_file = None
     if benchmark:
         bench_file = write_bench('read_img', start_time, end_time,
-                                 socket.gethostname(), output_dir, bn,
+                                 socket.gethostname(), bn,
                                  get_ident(), benchmark_dir=bench_dir)
 
     return (socket.gethostname(), filename, data, (im.affine, im.header),
@@ -67,9 +94,26 @@ def read_img(filename, data, benchmark, start, output_dir, bench_dir=None):
 
 
 def increment_data(idx, filename, data, metadata, delay, benchmark, start,
[…]

### 2018-12-14 results/code/makespan_to_json.py

@@ -1,29 +1,0 @@
-import json
-import sys
-
-
-with open(sys.argv[1], 'r') as f:
-    bench = {}
-    bench['tasks'] = []
-    count = 0
-
-    driver_start = None
-    for line in f:
-
-        experiment = line.split(':')[0]
-        l = line.split(' ')
-        l[0] = l[0].replace(experiment + ':', '')
-        print(l)
-        node = {}
-        node['name'] = l[0]
-        node['id'] = count
-        node['start-time'] = float(l[1])
-        node['end-time'] = float(l[2]) - node['start-time']
-        node['node'] = experiment
-
-        bench['tasks'].append(node)
-        count += 1
-
-    with open(sys.argv[2], 'w') as j:
-        json.dump(bench, j)
-

### 2018-12-14 tests/test_comp_avg.py

@@ -1,135 +1,0 @@
-import pytest
-import subprocess
-import hashlib
-import os
-import shutil
-from os import path as op
-import nipype_inc as ca
-import nibabel as nib
-import numpy as np
-
-
-chunk  = op.abspath('sample_data/dummy_1.nii')
-chunks = [chunk, op.abspath('sample_data/dummy_3.nii')]
-
-def test_increment():
-    delay = 0
-    benchmark = False
-    cli = True
-    im = ca.increment_chunk(chunk, delay, benchmark, cli=True)
-
-    assert op.isfile(im)
-
-    original = nib.load(chunk).get_data()
-    incremented = nib.load(im).get_data()
-
-    assert np.array_equal(incremented, original + 1)
-
-def test_increment_wf():
-    delay = 0
-    benchmark = False
-    benchmark_dir = None
-    cli = True
-    wf_name = 'test_incwf'
-    avg = None
-    work_dir = 'test_incwf_work'
-    inc_chunks = ca.increment_wf([chunks], delay, benchmark, benchmark_dir,
-                                  cli, wf_name, avg, work_dir)
-    for im in inc_chunks:
-        assert op.isfile(im)
-
-def test_compute_avg():
-    benchmark = False
-    benchmark_dir = None
-    avg = ca.compute_avg([chunks], benchmark, benchmark_dir)
-
-    images = [i for c in [chunks] for i in c]
-
-    data = None
-    print(images)
-
-    for im in images:
-        if data is None:
-            data = nib.load(im).get_data()
-        else:
-            data += nib.load(im).get_data().astype(data.dtype, copy=False)
-
-    data = data / len(images)
-
-    assert op.isfile(avg)
-    assert np.array_equal(nib.load(avg).get_data(), data)
-
-def test_compute_avg_wf():
-    from nipype import Workflow
-
-    nnodes = 1
-    work_dir = 'test_ca_wf_work'
-    chunk = chunks
-    delay = 0
-    benchmark_dir = None
-    benchmark = False
-    cli = True
-
-    wf = Workflow('test_ca_wf')
-    wf.base_dir = work_dir
-
-    inc_1, ca_1 = ca.computed_avg_node('ca_bb',
-                                       nnodes, work_dir,
-                                       chunk=chunk,
-                                       delay=delay,
-                                       benchmark_dir=benchmark_dir,
-                                       benchmark=benchmark,
[…]

### 2018-12-15 paper.tex
word change
final changes

@@
 \documentclass{IEEEtran}
-
+\usepackage[hyphens]{url}
 \usepackage{graphicx}
@@
     not employ strategies adopted by Big Data processing engines to improve 
-    Big Data processing time. Here, we 
+    processing time. Here, we 
     evaluate three Big Data processing stategies (in-memory computing, 
@@
     cases, represented by the BigBrain dataset. We contrast the various 
-    processing strategies by using Apache Spark and Nipype, as our 
+    processing strategies by using Apache Spark and Nipype as our 
     representative Big Data and neuroimaging processing engines, on Dell EMC's 
@@
 computing alone speeds-up executions by a factor of up to 1.6, whereas 
-when combined with data locality, this factor reaches 5.3. Such important 
+when combined with data locality, this factor reaches 5.3. Lazy evaluation strategies
+were found to increase the likelihood of cache hits, further improving processing time. Such important 
 speed-up values are likely to be observed on typical image processing 
@@
 functional MRI analyses unless coupled with data 
-locality, with a number of concurrently processed subject in the  280 ballpark. 
+locality and processing around 280 subjects concurrently. 
     In addition, we observe that emulating in-memory computing 
@@
 Data locality-aware scheduling, popularized by the MapReduce~\cite{dean2008mapreduce} 
-framework, executes tasks in the nodes holding the data rather than 
+framework, execute tasks within the nodes which store the data rather than 
 performing costly data transfers across the network. In-memory 
@@
 of workflow language expressivity, fault-tolerance, provenance capture 
-and representation, portability or reproducibility which are otherwise 
+and representation, portability or reproducibility, which are otherwise 
 critical concerns, addressed for instance in~\cite{samba}. Besides, our 
@@
 
-Apache Spark is a popular Scala-based processing framework for Big Data, with
+Apache Spark is a well-established Scala-based processing framework for Big Data, with
 APIs in Java, Python and R. Its 
@@
 redistributes the data across the partitions as requested. A shuffle marks a 
-stage boundary in Spark. In other words, an reduce operation  will not begin 
-until all 
-dependent map tasks have completed.
+stage boundary in Spark, where reduce-like operations will not begin 
+until all dependent map tasks have completed.
 
@@
 memory. Moreover, as Spark transformations generate new RDDs and numerous 
-transformations may occur in a single application, Spark implements a Least-Recently Used (LRU) eviction
+transformations may occur within a single application, Spark implements a
+Least-Recently Used (LRU) eviction
 policy. If an evicted RDD needs to be reused, Spark will recompute it using the
@@
 utilizes delay scheduling to optimize fairness and locality for all tasks.
-%improve paragraph
 
@@
 Standalone, 2) YARN~\cite{vavilapalli2013apache} and 3) Mesos~\cite{hindman2011mesos}. The Spark Standalone
-scheduler is the default scheduler. YARN was a scheduler designed for Hadoop~\cite{white2012hadoop}
-clusters and is prepackaged with Hadoop installations. In contrast to YARN, 
-Mesos was designed to be used in multitenant cluster environments. In our experiments,
+scheduler is the default scheduler, whereas YARN is scheduler designed for Hadoop~\cite{white2012hadoop}
+clusters and is prepackaged with Hadoop installations. In contrast to Standalone and YARN, 
+Mesos was designed to be used in multi-tenant cluster environments. In our experiments,
 we focus on the Spark Standalone cluster.
@@
 plugins to schedulers found in most clusters readily available to researchers, 
-such as the SunGrid Engine (SGE), Portable Batch System (PBS), Slurm and HTCondor\todo{cite}. It also includes its own scheduler, MultiProc,
-for parallel processing on single nodes. Nipype also includes many built-in 
-interfaces to commonly used neuroimaging tools to be incorporated within the 
+such as the Sun/Oracle Grid Engine (SGE/OGE), \href{http://www.adaptivecomputing.com/products/torque/}{TORQUE},
+\href{https://slurm.schedmd.com/}{Slurm} and \href{https://research.cs.wisc.edu/htcondor/}{HTCondor}. It also includes its own scheduler, MultiProc,
+for parallel processing on single nodes. Nipype additionally provides many built-in 
+interfaces to commonly used neuroimaging tools that can be incorporated within the 
 workflows. Nipype's ability to easily parallelize workflows in researcher-available 
@@
 and allow users to easily integrate existing neuroimaging tools, make it preferable
-to existing Big Data solutions, which would necessitate modification to achieve this.
+to existing Big Data solutions, which would necessitate modifications to achieve this.
 
@@
 tracking of inputs and outputs, and also enables checkpointing of the workflow.
-In the case of Node failure or application modification, only the nodes who have
-been modified (verified by hash) are re-executed.
+In the case of Node failure or application modification, only the nodes which have
+been modified (verified by hash) or have no successfully completed, are re-executed.
 
@@
 is required. However, it is still possible to save to a non-shared 
-local filesystem, but it may come at the expense of fault-tolerance, as 
-a result of being stored in a disk-local file system \todo{I don't get 'as a result of being stored in a disk-local file system'}. Moreover, the 
+local filesystem, but it may come at the expense of fault-tolerance, as data 
+located on failed nodes will be permanently lost. Moreover, the 
 user will need to ensure that the files are appropriately directed to 
@@
 Data may reside in the engine memory, on 
-a file system whose content reside in virtual memory (for instance 
+a file system whose contents reside in virtual memory (for instance 
 tmpfs), on disks local to the processing node, or on a shared 
@@
  what happens when the processed data fills up the available 
- memory. In case engine memory is used, the engine must cleanup 
- unused data with some strategy to avoid crashes. In case an in-memory 
+ memory. When engine memory is used, the engine must cleanup 
+ unused data with some strategy to avoid crashes. When an in-memory 
  file system is used, the kernel starts swapping memory to disk, and 
@@
 
-Storing data on local disks obviously enables data locality as tasks 
-executed on the node where their input files reside do not require to 
-transfer input or output data. However, in absence of a more specific 
+Storing data on local disks obviously enables data locality, as tasks 
+executed on the node where their input files reside do not require 
+transfer of input or output data. However, in absence of a more specific 
 file system such as HDFS to handle 
@@
  experiments, a Spark cluster was started on our 
- Slurm allocation through access to 16 dedicated nodes. Each compute 
- node ran on RHEL 7.4, was 
- equipped with 40 Intel Xeon Gold 6148 2.4GHz CPUs and had
- 187GB of memory, 94GB of tmpfs and a Dell 120GB M.2 SATA SSD as local disk. 
- A 698TB Lustre server was accessible on each node through a 100Gb/s 
- Infiniband network. The apparent write bandwidth of the local disk, 
+ Slurm allocation through access to 16 dedicated nodes.
+ Each Compute node has Red Hat Enterprise Linux Server release 7.4 (Maipo) 
+ as the base operating system with kernel version 3.10.0-693.17.1.el7.x86\_64
+ (patched for Spectre/Meltdown). Dell EMC PowerEdge C6420 with dual Intel Xeon
+ Gold 6148/F processors (40 cores per node) and 192GB ($12\times16$~GB), 2666 MHz
+ memory, serve as the compute nodes. Each compute has a 120GB M.2 SATA SSD as 
+ local disk. A Dell HPC Lustre Solution with a raw storage of 960TB is 
+ accesible on each compute node through a 100~GB/s Intel OmniPath network. All
+ the nodes connect to a director switch in a 1:1 non-blocking topology.
+ The apparent write bandwidth of the local disk, 
  Lustre file system and tmpfs were measured by sequentially writing various numbers
@@
 
-As it is possible that both small and large data may need to be processed using
+Since both small and large datasets may need to be processed using
 the same analysis pipeline, we 
-wanted to determine the effects of the data management strategies on small 
+examined the effects of the data management strategies on small 
 data as well. For this, we selected a 12MB T1W image belonging to 
@@
 application). Each partitioned chunk was incremented by 1, in parallel, by
-the applications. As incrementing the images is a relatively quick 
-process, we added a sleep delay to the incrementation stage in order to 
-increase task duration as required. The incremented chunks would be either 
+the applications. As incrementing the images is not a time consuming process,
+we added a sleep delay to the incrementation stage to study the effects 
+of tasks duration. The incremented chunks would be either 
 maintained in-memory (Spark only) or saved to either tmpfs, local disk or 
 Lustre (Spark and Nipype). Should more than a single iteration be requested, 
-the incremented chunks would be incremented again and saved to the same storage. 
+the incremented chunks would be incremented again and saved to the same file system. 
 This would repeat until the number of requested iterations had elapsed. In all 
-conditions, the first input chunks and final output chunks would be read from / written to  Lustre. 
-We chose to perform our initial 
-reads and final writes on Lustre as a multi-tenant clusters' compute node's 
-local storage, if 
-available, is generally temporary and only available for the duration of the 
-allocation.
+conditions, the first input chunks and final output chunks would reside on Lustre. 
+We chose to perform our initial input/final output on Lustre as local storage is
+typically only accessible to a user for the duration of the execution in HPC
+environments.
 
@@
 Nipype (a new Slurm allocation is requested for each processed task), 
-we instrumented Nipype to ensure data locality \todo{You could refer to the script doing that in the paper repo.}. That is, the chunk were split into partitions, and for each partition of 
-chunks, we requested a Slurm allocation to process the entire 
+we instrumented Nipype to ensure data locality (see: \href{https://github.com/big-data-lab-team/paper-in-mem-locality/blob/master/benchmark\_scripts/run\_benchmarks.py}{run\_benchmarks.py}).
+That is, the chunk were split into partitions, and for each partition,
+we requested a Slurm allocation to process the entire 
 pipeline in parallel, using Nipype's MultiProc scheduler, on a given 
@@
 For our first incrementation experiment, we investigated the effects of 
-Big Data strategies on varying total data size. To do this, we 
+Big Data strategies on varying total data size. To achieve this, we 
 increased the number of incrementation iterations from 1, 10 and 100 
@@
 strategies on varying task duration. If the page cache has sufficient 
-time to flush, it should be expected that using in-memory computing 
-should have no increased speed-up over disk. We varied the task 
-duration between 2.4 seconds and 320 seconds so that the D/C falls into 
+time to flush, it would be expected that in-memory computing 
+and local disk perform equivalently. We varied the task 
+duration between 2.4 and 320 seconds such that the D/C falls into 
 different regions of Equations~\ref{eq:page-cache-inequality} 
@@
 As a third incrementation experiment, we were interested in the effects of 
-chunk size on Big Data strategies. Naturally, greater chunk size signifies 
-a decrease in parallelization, however, it also signifies an increase in 
+chunk size on Big Data strategies. Naturally, a greater chunk size signifies 
+a decrease in parallelization. However, it also signifies an increase in 
 sequential I/O (increased $\Delta/\Gamma$ and $\delta/\gamma$). For 
@@
 ratio ensured that different regions of the inequality were reached 
-depending on amount of parallelization. Unfortunately, 30 chunks could 
-no be processed using in-memory Spark as its individual chunk size of 
-3GB exceeded the 2GB partition size limit imposed by Spark \todo{You should move this comment to the results or the reader will wonder
-what happened to the yellow bar when looking at figure 1.c.}. The number of iterations was
+depending on amount of parallelization. The number of iterations was
 fixed to 10, and the task duration was adjusted so that C remained constant at 4,400s.
@@
 38GB half BigBrain and the 13M T1W MRI image for this experiment. The number of
-chunks was fixed at 125. Similarly to the previous experiment, the 
+chunks was fixed to 125. Similarly to the previous experiment, the 
 total sequential compute time was fixed (10 iterations, 1.76 seconds 
@@
 filesystem choices given the number of iterations. At 1 iteration, all 
-filesystems behave the same, although the application was writing faster than the page cache flushed. \todo{Update the graph to the same y scale and 'broken'
-bars, and check that this impression still holds...}
-This is explained by the fact that the application data was not large enough to saturate the page  
-cache fully (transient phase). As the total amount of data written by the 
+filesystems behave the same, although the application was writing faster than the
+disk bandwidth. This is because application data was not saturating the Page Cache
+(transient phase). The page cache, on Zenith, occupies 40\% of 
+total memory. With 192~GB of RAM on each node, 76.8~GB of 
+dirty data could be held in a node's page cache at any given time.
+As the total amount of data written by the 
 application increases to 750~GB, there is a greater disparity between 
@@
 
-%sp_mem 1it = 34s, 10it = 64 100it = 381
-%3.44 * 1 = 3.44, 3.44*10 = 34, 3.44*100 = 344
-
-%sp_tmpfs 1it = 31, 10it = 69,
-
 At 100 iterations, or 7,500~GB, Lustre can be found to be, on average, 3.82~x 
-slower than Spark in-memory. The page cache, on Zenith, occupies 20\% of 
-total memory. Since there was 187~GB of RAM on each node, only 37.4~GB of 
-dirty data could be held in a node's page cache at any given time. The 
+slower than Spark in-memory. The 
 slowdown experienced can be explained by the smaller percentage of total data 
@@
 % Comment on the general trend
-% note 'on average' = average between spark and nipype. local disk spark is 
-% 1.22x slower and local disk nipype is 1.54x slower
-% spark lustre 2.66x slower; nipype lustre 2.68x slower
 
@@
 reasoning behind this is that longer tasks give the page cache more time to flush 
-between disk writes. The longer the delay, the less likely the page cache is 
-to fill up when the concurrent data size written fits within the page cache.
+between disk writes.
 
@@
 number of chunks increases. This is due to the fact that parallelism 
-increases with an increase in the number of chunks. With 30 chunks, 
+increases with an increase in number of chunks. At 30 chunks, 
 only 2 CPUs per node are actively working. At 125 chunks, this changes 
-to a maximum of 9 CPUs per node, and at 750 chunks all 40 CPUs can be 
-active. It can be noted that no significant speedup can be seen between 125 
-chunks and 750 chunks \todo{This is not true for in-memory, Spark tmpfs and Spark local disk}. This is because out of the 50 chunks processed by each 
-node, only 40 can be processed concurrently, thus requiring two sequential 
-incrementation batches and resulting in similar makespans as 125 
-chunks.
-
-In terms of filesystem selection, local and tmpfs perform comparably for all 
+to a maximum of 9 CPUs per node, and at 750 chunks, up to 40 CPUs can be 
+active.
+
+Due to a size limitation of 2~GB imposed on Spark partitions, Spark with in-memory
+computing processing 30 chunks was not performed.
+
+Local disk and tmpfs perform comparably for all 
 conditions, with Lustre being significantly slower. As with varying the number 
 of iterations, Lustre is slower due to increased filesystem contention, which 
-is, at minimum, 15x greater than contention on local disk. With an increase in 
+is, at minimum, 15~x greater than contention on local disk, due to the number of nodes
+used. With an increase in 
 number of chunks, local disk and tmpfs makespans begin to converge. A potential 
@@
 pipeline were cached and all data concurrently accessed could be mantained 
-in-memory, even for the 750 chunks which were processed in two batches, it did 
+in-memory, it did 
 not need to spill to disk.
@@
 Increasing overall data size decreases performance, as can be seen in 
-Fig.~\ref{fig:datasize}. When the data size is very small, as is the case 
-with the MRI image, all file system makespans are comparable. This is due to the 
+Fig.~\ref{fig:datasize}. When the data size is very small (e.g. MRI image) all 
+file system makespans are comparable. This is due to the 
 fact that page cache can be leveraged fully regardless of file system. However, 
 this time, Spark in-memory performed significantly worse than all other 
-filesystems, with a makespan of 2,211 seconds (not reported on the graph to preserve the y scale) \todo{You may want to add it back when you have
-broken bars}. Upon further inspection, it appeared that Spark in-memory executed
+filesystems, with a makespan of 2,211 seconds. Upon further inspection,
+it appeared that Spark in-memory executed
 in a sequential order, on a single worker node. Lack of parallelism for the MRI 
-image may be a result of Spark's max partition size, which is by default 128~MB
+image may be a result of Spark's maximum partition size, which is by default 128~MB
 -- significantly larger than the 13~MB MRI image.
@@
 filesystems becomes even greater. Lustre becomes 3.68x slower, whereas local 
-disk is now 1.68x slower. The reasoning for this is that the page cache fills 
-up faster than with half of BigBrain.
-
-\todo{change Spark's max partition size and test to see if it fixes the problem}
+disk becomes 1.68x slower. An explanation for this is that the page cache fills 
+up faster due to data size.
 
@@
 
-%probably confusing to mention values, but they're here for now
 Results show that, overall, the model correctly predicted the effect of 
@@
 per subject, which is a ballpark value commonly observed with the 
-popular SPM or FSL packages \todo{add refs}, an input data size of 100MB per subject, 
+popular \href{https://www.fil.ion.ucl.ac.uk/spm/}{SPM} or \href{https://fsl.fmrib.ox.ac.uk/fsl/fslwiki/}{FSL packages},
+an input data size of 100MB per subject, 
 and an output data size of 2GB (20-fold increase compared to input 
@@
 writing locally was usually preferrable to writing to Lustre, as a result of the
-lower contention on local disk. Although it may be true, as stated in~\cite{Anantharayanan2011datalocality},
-network bandwidths generally exceed that of disks, locality remains important 
-as contention to a shared filesystem will be much higher than local disk.
-The only time in our experiments where it was found that writing locally did not
-have a huge impact over Lustre, was in experiments 1 and 4, at 1 iteration and 
+lower contention on local disk. Although it may be true that,
+network bandwidths exceed that of disks~\cite{ananthanarayanan2011disk},
+locality remains important 
+as contention on a shared filesystem will tends to be much higher than on local disk.
+The only time writing locally did not
+have significant impact over Lustre was in experiments 1 and 4, at 1 iteration and 
 when processing the MRI image, respectively. In both these scenarios, the Lustre 
@@
 The effects of lazy-evaluation can be seen throughout the experiments. Nipype 
-was found to generally be slower than Spark in all experiments. While a Nipype DAG is 
-generated prior to workflow execution, there are no such optimizations, as in 
-Spark, to ensure that the least amount of data transfer is performed to produce the 
+was found to be slower than Spark in most experiments. While the Nipype execution
+graph is 
+generated prior to workflow execution, there are no optimizations
+to ensure that the least amount of work is performed to produce the 
 results.
 
-During the processing of Experiment 3, it was found that Spark split up the data
-into an processed it in two batches for the 750 chunk input. That is to say, 
-each node in the cluster processed 50 chunks, and since there was only a maximum
-of 40 cores available on each node, Spark opted to run \todo{a word is missing here} to batches of 25 to ensure
-a balanced workload. Rather than running each iteration on the full dataset, as with Nipype, 
+During the processing of Experiment 3, 750 chunks were processed in two batches
+for both Spark and Nipype due to CPU limitations.
+Rather than running each iteration on the full dataset, as with Nipype, 
 Spark opted to perform all the iterations on the first batch (load, increment, save),
-and then proceeds to process the second batch. Such an optimization is important, 
+and then proceeded to process the second batch. Such an optimization is important, 
 even when processing data on disk, as it would presumably increase the occurrence
@@
 Figure~\ref{fig:numchunks}.% Spark local vs nipype local: presumably increases cache hits
-\todo{Add as sentence on this paragraph in the abstract.}
 
@@
 
-Although tmpfs and page cache do improve performance in general, as seen in
+Although tmpfs and page cache do improve performance, as seen in
 Figure~\ref{fig:results}, they do not always perform equivalently to in-memory. 
@@
 the system's memory usage is high. When it reaches this point, its performance 
-slows down to swap disk bandwidth, as observed in Figure~\ref{fig:datasize}. Page cache suffers a similar dilemma. As detailed
-in~\href{https://access.redhat.com/documentation/en-us/red_hat_enterprise_linux/6/html/performance_tuning_guide/s-memory-tunables}{RedHatDoc}\todo{add to references},
-I/O blocking write to disk occurs when a giving percentage (e.g. 20~\%) of total memory is occupied by dirty data, once this 
-threshold is exceeded, processes wanting to execute writes must wait until the dirty data
-in the cache is flushed to disk \todo{Revise this sentence (grammar?)}.
+slows down to swap disk bandwidth, as observed in Figure~\ref{fig:datasize}. Page cache suffers a similar dilemma. I/O blocking writes to disk occur when a given percentage (e.g. 40~\%)
+of total memory is occupied by dirty data. When the threshold is exceeded,
+processes performing writes must wait for dirty data to be flushed to disk.
 
 Furthermore, like memory, tmpfs and page cache are shared resources on a node.
-If users on the node are heavily using memory to incite tmpfs to write to swap space, or are performing data-intensive operations
-that fill up the page cache, this will limit the performance on tmpfs/page cache of other users. 
-However, with memory, it is possible to request, through the HPC scheduler, a required
+If users on the node are heavily using memory to incite tmpfs to writes to swap space, or are performing data-intensive operations
+that fill up the page cache, tmpfs/page cache performance on will be limited for other users. 
+However, with memory, it is possible to request through the HPC scheduler, a certain
 amount of available memory. Ultimately, in-memory data will also need to be spilled to
@@
 than to load balance data across the cluster. That is, given 50 chunks 
-and 40 cores, Spark will only use up 25 cores, and will break it up 
-into two batches. Nipype, on the other hand, will also have no choice 
+and 40 cores, Spark will only use up 25 cores and process in two batches.
+Nipype will also have no choice 
 but to split up the processing into two batches, but will first process 
@@
 to a single element, due to a blocking mechanism. For this reason, we 
-selected to iterate through a series of Nodes in our code \todo{while 
-MapNodes might have been easier to use?}.
-
-\todo{If time permits, it would be useful to say that we checked that scheduling between 
-both engines was similar, and show two Gantt charts to justify that.}
+selected to iterate through a series of Nodes in our code despite MapNodes being
+easier to use.
+
+% \todo{If time permits, it would be useful to say that we checked that scheduling between 
+% both engines was similar, and show two Gantt charts to justify that.}
 
@@
 with it. While Spark will recompute the lost partition automatically using lineage, 
-Nipype will fail for all tasks requiring the data. Spark, however, will fail if 
-RDDs of node-local filenames are shuffled, as the data associated with the filenames
-will not be shuffled with the RDD
+Nipype will fail for all tasks requiring the data. Nevertheless, Spark will alse fail if 
+RDDs of node-local filenames are shuffled, as the data associated to the filenames
+will not be shuffled with the RDD and there will be no mechanism in place to fetch it.
 
@@
 It would also not be an option with Spark in-memory. To enable this, burst buffers or 
-heteorogeneous storage managers (e.g. Triple-H~\cite{tripleH}) could be used to ensure
+heteorogeneous storage managers (e.g. Triple-H~\cite{islam2015triple}) could be used to ensure
 fast processing and that all outputs (including intermediate outputs) will be 
-sent asynchronously to the shared filesystem. \todo{double check triple-H}
+sent asynchronously to the shared filesystem.
 
@@
 
-Data-locality was also found to play an important role in application 
-performance. Local disk was found to perform between than the shared 
-filesystem \todo{between than the shared filesystem: what does it 
-mean?}, despite having lower bandwidth, due to increased contention on 
-the shared filesystem. As local disk will generally have less 
-contention than a shared filesystem. It is recommended to have the data 
-stored locally. Although, storing data locally may come at a cost to 
-fault tolerance.
+Data-locality plays an important role in application 
+performance. Local disk was found to perform better than the shared 
+filesystem, despite having lower bandwidth, due to increased contention on 
+the shared filesystem. Since local disk will typically have less 
+contention than a shared filesystem, it is recommended to store data locally.
+However, using local storage without a distributed file system 
+may come at a cost to fault tolerance.
 
@@
 performed, it is speculated that this may be the cause of the general 
-performance difference between Spark and Nipype. Furthermore, it is 
+performance difference between Spark and Nipype. Furthermore, it was
 found that lazy evaluation optimizations increase the likelihood of 

### 2018-12-15 results/code/gen_figures.py

@@ -40,9 +40,6 @@ def genfig(exp, makef, outf, cond):
        
             iterations = int(iterations)
 
-            if exp == 3 and fs == 'mem' and im == 'MRI':
-                continue
-
             if exp == 1:
                 if delay != cond: continue
                 bench_res[(engine, fs)].append((iterations,
@@ -188,7 +185,7 @@ def genfig(exp, makef, outf, cond):
         ax2.plot((-d, +d), (1 - d, 1 + d), **kwargs)
         ax2.plot((1 - d, 1 + d), (1 - d, 1 + d), **kwargs)
 
-        ax.legend(loc=2)
+        ax.legend()
         plt.savefig(outf)
 
 

### 2018-12-15 paper.tex
final final edits

@@
 the applications. As incrementing the images is not a time consuming process,
-we added a sleep delay to the incrementation stage to study the effects 
+we added a sleep delay to the incrementation task to study the effects 
 of tasks duration. The incremented chunks would be either 
@@
 filesystems behave the same, although the application was writing faster than the
-disk bandwidth. This is because application data was not saturating the Page Cache
+disk bandwidth. This is because application data was not saturating the page cache
 (transient phase). The page cache, on Zenith, occupies 40\% of 
@@
 
-Spark currently doesn't include any active management of disk I/Os or 
+Spark does not currently include any active management of disk I/Os or 
 page caches. We believe that it would be beneficial to extend it toward 

### 2019-03-28 paper.tex
fixes to content, bugs in figure display

@@
     advancements in image acquisition methods, open-science and data sharing. 
-    However, it remains that current neuroimaging data processing engines do 
-    not employ strategies adopted by Big Data processing engines to improve 
-    processing time. Here, we 
-    evaluate three Big Data processing stategies (in-memory computing, 
-    data-locality and lazy-evaluation) on typical neuroimaging use 
-    cases, represented by the BigBrain dataset. We contrast the various 
-    processing strategies by using Apache Spark and Nipype as our 
+    However, current neuroimaging data engines have not adopted any strategies
+    to address the rise
+    in neuroimaging data. Here, we 
+    evaluate three Big Data processing strategies (in-memory computing, 
+    data locality and lazy evaluation) on typical neuroimaging use 
+    cases, represented by the BigBrain dataset. We contrast these various 
+    strategies using Apache Spark and Nipype as our 
     representative Big Data and neuroimaging processing engines, on Dell EMC's 
     Top-500 cluster. 
-    Big Data thresholds were modeled by comparing the data/compute ratio of the 
-    application to the filesystem/concurrent processes ratio. 
+    Big Data thresholds were modeled by comparing the data-write rate of the 
+    application to the filesystem-application concurrent write rate. 
     This model acknowledges the 
@@
 computing alone speeds-up executions by a factor of up to 1.6, whereas 
-when combined with data locality, this factor reaches 5.3. Lazy evaluation strategies
-were found to increase the likelihood of cache hits, further improving processing time. Such important 
+when combined with data locality, this factor reaches 5.3. Lazy evaluation
+strategies
+were found to increase the likelihood of cache hits, further improving
+    processing time. Such important 
 speed-up values are likely to be observed on typical image processing 
-operations performed on images of size larger than 75GB. A ballpark 
-speculation from our model showed that in-memory computing alone will
-not speed-up current 
-functional MRI analyses unless coupled with data 
-locality and processing around 280 subjects concurrently. 
-    In addition, we observe that emulating in-memory computing 
-using in-memory file systems (tmpfs) does not reach the performance of 
-an in-memory engine, presumably due to swapping to disk and the lack of data cleanup.
+operations performed on images of size larger than 75GB. 
 We conclude that Big Data processing strategies are 
@@
 
-Big Data processing engines reduce data-related overheads through 
-use of concepts such as data locality, in-memory computing, and lazy evaluation.
-Data locality-aware scheduling, popularized by the MapReduce~\cite{dean2008mapreduce} 
-framework, execute tasks within the nodes which store the data rather than 
-performing costly data transfers across the network. In-memory 
-computing, adopted by MapReduce's successor Apache 
-Spark~\cite{zaharia2016apache}, stores data in memory whenever 
-possible, reducing costly disk I/Os between pipeline steps. 
-Lazy evaluation, also used in Spark, enables further Big Data performance 
-optimizations, by evaluating results only when necessary. Frameworks 
-such as
-MapReduce and Spark have become mainstream tools for data analytics, 
-although many others, such as 
-Dask~\cite{rocklin2015dask}, are emerging. 
+In Big Data applications, there is a significant cost associated with data movement.
+Big Data processing engines have addressed these issues by diminishing the amount 
+of data movement that occurs during the execution of an application. The 
+MapReduce~\cite{dean2008mapreduce} framework popularized the concept of locality-aware
+scheduling, which ensured tasks were scheduled nearest to where the data the data was
+located (data locality), thereby reducing the number of network transfers in an application.
+However, it was found that performance in MapReduce could be improved upon by also
+reducing the disk I/O between tasks by keeping data in-memory whenever possible,
+and thus, in-memory computing was added to the Apache Spark~\cite{zaharia2016apache} framework.
+Frameworks like Apache Spark also included the lazy evaluation strategy, which performs computations
+only when invoked. This allows for additional optimizations such as the regrouping of tasks
+and only invoking what is absolutely necessary to obtain the desired result.
+
+Frameworks such as
+MapReduce and Spark have become mainstream tools for data analytics, however, many
+others, such as Dask~\cite{rocklin2015dask}, are emerging.
 Meanwhile, several scientific 
@@
 using a mix of programming languages (C, Fortran, Python, shell 
-scripts), referred to as workflows or pipelines. Infrastructure-wise, 
+scripts), referred to as workflows or pipelines. With respect to infrastructure, 
 Big Data applications commonly run on 
@@
 (HDFS~\cite{shvachko2010hadoop}), whereas scientific applications are 
-usually deployed on large, shared clusters where data is moved between
-data nodes and compute nodes using shared file systems such 
+usually deployed on large, shared clusters where data is transferred between
+data and compute nodes through shared file systems such 
 as Lustre~\cite{schwan2003lustre}. Such differences in applications and 
-infrastructures have important consequences. To 
+infrastructure have important consequences. To 
 mention only one, in-memory computing requires instrumentation to be 
@@
 or the Pipeline System for Octave and Matlab 
-(PSOM~\cite{bellec2012pipeline}). Big Data engines have been used for 
-neuroimaging applications too, including in the Thunder 
+(PSOM~\cite{bellec2012pipeline}). Big Data engines have also been used for 
+neuroimaging applications, including the Thunder 
 project~\cite{freeman2014mapping} and in more specific works such 
@@
 
-%\todo{Perhaps talk about modeling and simulation here.}
-
 Our methods, including performance models, processing engines, 
@@
 data streaming and graph processing. Spark's
-main features include data-locality, in-memory processing and lazy evaluation,
+main features include data locality, in-memory processing and lazy evaluation,
 which it achieves through its principle abstraction, the Resilient Distributed 
@@
 log of data modifications. This log, known as the lineage, is used
-to reproduce any lost data.
+to reproduce any lost data modifications.
 
@@
 RDD is dependent on all parent partitions in a wide dependency. Examples of 
-Spark transformations include map, filter and join. To materialize an RDD, an
+transformations include map, filter and join. To materialize an RDD, an
 action must be performed, such as a reduce or a collect. Lazy evaluation is 
@@
 
-Spark will spill RDD data to disk whenever the complete dataset does not fit in
-memory. Moreover, as Spark transformations generate new RDDs and numerous 
+Although Spark uses in-memory computing, it is not necessary for all the data to
+fit in memory. Spark will spill any RDD elements that do not fit in memory to disk.
+Moreover, as Spark transformations generate new RDDs and numerous 
 transformations may occur within a single application, Spark implements a
@@
 policy. If an evicted RDD needs to be reused, Spark will recompute it using the
-lineage data collected. As demonstrated in~\cite{freeman2014mapping}, caching significantly improves processing times of iterative algorithms where 
+lineage data collected. As demonstrated in~\cite{freeman2014mapping}, caching
+significantly improves processing times of iterative algorithms where 
 RDDs are reused. It can be of even greater importance if the RDD is costly to 
@@
 Standalone, 2) YARN~\cite{vavilapalli2013apache} and 3) Mesos~\cite{hindman2011mesos}. The Spark Standalone
-scheduler is the default scheduler, whereas YARN is scheduler designed for Hadoop~\cite{white2012hadoop}
-clusters and is prepackaged with Hadoop installations. In contrast to Standalone and YARN, 
+scheduler is the default scheduler. The YARN scheduler designed for Hadoop~\cite{white2012hadoop}
+clusters and is prepackaged with Hadoop installations, whereas in contrast, 
 Mesos was designed to be used in multi-tenant cluster environments. In our experiments,
@@
 
-% i use 'popular' for both... widely-used (for Spark)?
 Nipype is a popular Python neuroimaging processing engine. It aims at being a 
 solution for easily creating reproducible neuroimaging workflows. Although Nipype
-does not employ any Big Data processing strategies, it provides access to numerous
-plugins to schedulers found in most clusters readily available to researchers, 
+does not employ any Big Data processing strategies, it provides plugins to numerous
+schedulers found in most clusters readily available to researchers, 
 such as the Sun/Oracle Grid Engine (SGE/OGE), \href{http://www.adaptivecomputing.com/products/torque/}{TORQUE},
 \href{https://slurm.schedmd.com/}{Slurm} and \href{https://research.cs.wisc.edu/htcondor/}{HTCondor}. It also includes its own scheduler, MultiProc,
-for parallel processing on single nodes. Nipype additionally provides many built-in 
+for parallel processing on single nodes. Furthermore, Nipype provides many built-in 
 interfaces to commonly used neuroimaging tools that can be incorporated within the 
@@
 and allow users to easily integrate existing neuroimaging tools, make it preferable
-to existing Big Data solutions, which would necessitate modifications to achieve this.
-
-Jobs, or Interfaces, in Nipype, are encapsulated by a Node. A Node dictates that
+over existing Big Data solutions, which would necessitate modifications to achieve this.
+
+Jobs, or Interfaces in Nipype, are encapsulated by a Node. A Node dictates that
 the job will execute on a single input. However, the MapNode, 
@@
 Nipype execute in their own uniquely named subdirectory which facilitates provenance 
-tracking of inputs and outputs, and also enables checkpointing of the workflow.
+tracking of inputs and outputs and also enables checkpointing of the workflow.
 In the case of Node failure or application modification, only the nodes which have
-been modified (verified by hash) or have no successfully completed, are re-executed.
+been modified (verified by hash) or have not successfully completed, are re-executed.
 
@@
  memory. When engine memory is used, the engine must cleanup 
- unused data with some strategy to avoid crashes. When an in-memory 
- file system is used, the kernel starts swapping memory to disk, and 
- the performance becomes that of local disk writes. In our experiments, 
- we will explore the region  when the data consumed by the 
+ unused data to avoid crashes. When an in-memory 
+ file system is used, the user is responsible ensuring that the filesystem does not
+ reach capacity. Should it be necessary to free up additional memory, the kernel will swap filesystem memory to disk,
+ and the performance will become that of local disk writes. In our experiments, 
+ we will explore the configuration where the data consumed by the 
  application approaches the threshold of available memory.
@@
 
-Storing data on local disks obviously enables data locality, as tasks 
-executed on the node where their input files reside do not require 
-transfer of input or output data. However, in absence of a more specific 
-file system such as HDFS to handle 
+Storing data on local disks inevitably enables data locality, since data transfers are
+not necessary when tasks are executed on the nodes where the data resides.
+However, in absence of a more specific 
+filesystem such as HDFS to handle 
 file replication across computing nodes, data locality comes at the price
@@
 strategies currently cannot be controlled by the application, which 
-prevents processing engines to anticipate reads by preloading the 
+prevents processing engines from anticipating reads by preloading the 
 cache. Scheduling strategies might be designed 
-to maximize cache hits though. For instance, lazy 
+to maximize cache hits, however. For instance, lazy 
 evaluation could result in more cache hits by scheduling data-dependent 
@@
 \end{equation}
-This defines a data-compute ratio beyond which the page cache becomes 
+This defines a data-write rate beyond which the page cache becomes 
 asymptotically useless. It should be noted that the transient phase 
@@
 transient phase as it requires detailed knowledge of difficult to estimate parameters  
-such as the page cache and the initial amount 
+such as the page cache size and the initial amount 
 of data in it ($d_0$).
@@
 
-We model a shared file system using its global apparent bandwidth 
+We model a shared file system using its global realized bandwidth 
 $\Delta$, shared by all concurrent processes in the cluster. We are 
@@
 management, RPC protocol optimizations and storage optimizations are 
-all covered under the apparent bandwidth. We do, however, consider the 
-effect of page caching in shared file systems too, since in Linux 
+all covered under the realized bandwidth. We do, however, consider the 
+effect of page caching in shared file systems, since in Linux 
 writes to network-mounted volumes benefit from this feature too.
@@
  EMC HPC and AI Innovation Lab}, running Slurm. For the Spark 
- experiments, a Spark cluster was started on our 
- Slurm allocation through access to 16 dedicated nodes.
+ experiments, a Spark cluster was started on a
+ Slurm allocation comprised of 16 dedicated nodes.
  Each Compute node has Red Hat Enterprise Linux Server release 7.4 (Maipo) 
@@
  local disk. A Dell HPC Lustre Solution with a raw storage of 960TB is 
- accesible on each compute node through a 100~GB/s Intel OmniPath network. All
+ accesible on each compute node through a 100~Gb/s Intel OmniPath network. All
  the nodes connect to a director switch in a 1:1 non-blocking topology.
- The apparent write bandwidth of the local disk, 
+ The realized write bandwidth of the local disk, 
  Lustre file system and tmpfs were measured by sequentially writing various numbers
  of image blocks containing random intensities, to avoid 
- caching effects (measure\_bandwidth.py script in folder 
- benchmark\_scripts). They are reported in Table~\ref{table:bdwdths}.
+ caching effects (measure\_bandwidth.py script).
+ They are reported in Table~\ref{table:bdwdths}.
 
@@
 We used BigBrain~\cite{amunts2013bigbrain}, a 75GB 40$\mu$m isotropic 
-histological image of a 65-year-old male's brain. The BigBrain was selected due 
-to being the only one of its kind, as there does not yet exist any other
+histological image of a 65-year-old human brain. The BigBrain was selected due 
+to its uniqueness, as there does not yet exist a
 higher-resolution image of a human brain. Moreover, there currently exists a 
@@
 Processing large images is only considered to be part of the Big Data problem 
-in neuroscience. The other problem is processing large MRI datasets, 
+in neuroscience. The other being the processing large MRI datasets, 
 that is, datasets consisting of many small brain images belonging to 
@@
 \href{https://openneuro.org/datasets/ds000001/versions/00006/file-display/sub-01:anat:sub-01_T1w.nii.gz}{ds000001 
-dataset version 6}. In order to be able to split it into 125 chunks, it 
+dataset version 6}. In order to split the image into 125 equal-sized chunks, it 
 was necessary to zero-pad the image to the dimensions
@@
 
-In order to investigate how the different strategies impact processing, we 
+To effectively investigate how the different strategies impact processing, we 
 selected a simple incrementation pipeline (Algorithm~\ref{alg:incrementation}) 
-that consisted exclusively of map 
+that consists exclusively of map 
 stages. A series of map-only stages would enable us to evaluate the effects of
@@
 application). Each partitioned chunk was incremented by 1, in parallel, by
-the applications. As incrementing the images is not a time consuming process,
-we added a sleep delay to the incrementation task to study the effects 
+each task. As incrementing images is not a time consuming process,
+we added a sleep delay to the tasks to study the effects 
 of tasks duration. The incremented chunks would be either 
@@
 Algorithm~\ref{alg:incrementation}), (2) the task duration ($x$), (3) 
-the chunk size for a constant image size, (4) the total image size. To 
+the chunk size given a constant image size, (4) the total image size. To 
 evaluate the page-cache model, experiment conditions fell in different 
@@
 dedicated to the Spark master and driver, and the remaining 15 were 
-used as compute nodes. Since data-locality is not normally preserved in 
+used as compute nodes. Since data locality is not normally preserved in 
 Nipype (a new Slurm allocation is requested for each processed task), 
 we instrumented Nipype to ensure data locality (see: \href{https://github.com/big-data-lab-team/paper-in-mem-locality/blob/master/benchmark\_scripts/run\_benchmarks.py}{run\_benchmarks.py}).
-That is, the chunk were split into partitions, and for each partition,
+That is, the chunks were split into partitions, and for each partition,
 we requested a Slurm allocation to process the entire 
-pipeline in parallel, using Nipype's MultiProc scheduler, on a given 
+pipeline in parallel using Nipype's MultiProc scheduler on a given 
 node. This was possible as no communication was required between the 
@@
 iteration, to 7,500GB, at 100 iterations. The total number of chunks 
-was 125. Chunks all ran concurrently ($\Gamma$=125) and they were 
+was 125. Chunks all ran concurrently ($\Gamma$=125) and were 
 equally balanced among 15 nodes, leading to 8 or 9 concurrent jobs per 
@@
 chunk size on Big Data strategies. Naturally, a greater chunk size signifies 
-a decrease in parallelization. However, it also signifies an increase in 
+a decrease in parallelism. However, it also signifies an increase in 
 sequential I/O (increased $\Delta/\Gamma$ and $\delta/\gamma$). For 
@@
 use up as many cores as possible. Unlike the previous 
-experiment, the D/C ratio was kept static at 178.6MB/s, however, this 
+experiment, the data-write rate was kept static at 178.6MB/s, however, this 
 ratio ensured that different regions of the inequality were reached 
-depending on amount of parallelization. The number of iterations was
+depending on amount of parallelism. The number of iterations was
 fixed to 10, and the task duration was adjusted so that C remained constant at 4,400s.
@@
 per task), however, due to varying size in total data processed (D), 
-the D/C ratio varied. Once again, we ensured that the D/C ratio fell in 
-multiple different regions of the inequality. The D/C ratio ranged from 
-349.1 MB/s for BigBrain and 174.6MB/s for half Bigbrain, to 0.06MB/s for the MRI 
+the data-write rate varied. Once again, we ensured that the data-write rate fell in 
+multiple different regions of the inequality. The data-write rate ranged from 
+349.1 MB/s for BigBrain and 174.6MB/s for half BigBrain, to 0.06MB/s for the MRI 
 image. Only the 0.06MB/s MRI satisfied the inequality for both Lustre and local 
@@
     \centering
+    \begin{subfigure}{\columnwidth}
+        \includegraphics[clip,width=\columnwidth]{results/figures/s-iterations.pdf}%
+    \end{subfigure}
+    \begin{subfigure}{\columnwidth}
+        \includegraphics[clip,width=\columnwidth]{results/figures/s-iterations.pdf}%
+    \end{subfigure}
     \includegraphics[width=\linewidth]{results/figures/iterations.pdf}
@@
 Fig.~\ref{fig:iterations} shows the difference between the different 
-filesystem choices given the number of iterations. At 1 iteration, all 
+filesystem choices given varying number of iterations. At 1 iteration, all 
 filesystems behave the same, although the application was writing faster than the
@@
 % Comment on the general trend
-As can be seen in Fig.~\ref{fig:numchunks}, makespan decreases when the 
-number of chunks increases. This is due to the fact that parallelism 
+As can be seen in Fig.~\ref{fig:numchunks}, makespan decreases with increasing
+number of chunks. This is due to the fact that parallelism 
 increases with an increase in number of chunks. At 30 chunks, 
@@
 pipeline were cached and all data concurrently accessed could be mantained 
-in-memory, it did 
-not need to spill to disk.
+in-memory, spilling to disk was not necessary.
 
@@
 page cache on processing times for local disk and Lustre. That is, the 
-speed-up provided by in-memory computing was larger than 1 for D/C 
-ratios larger than $\delta/\gamma$ (local disk) or $\Delta/\Gamma$ 
+speed-up provided by in-memory computing was larger than 1 for data-write 
+rates larger than $\delta/\gamma$ (local disk) or $\Delta/\Gamma$ 
 (Lustre). Conversely, the speed-up provided by in-memory computing 
-remained close to 1 for D/C ratios smaller than $\delta/\gamma$ (local 
+remained close to 1 for data-write rates smaller than $\delta/\gamma$ (local 
 disk) or $\Delta/\Gamma$ (Lustre). The two points close to the origin 
@@
 % Shared fs vs local disk vs in-memory: what do we gain?
-We measure the effect of in-memory computing by comparing the runs 
+We measured the effect of in-memory computing by comparing the runs 
 of Spark in-memory (yellow bars in Fig.~\ref{fig:results}) to the ones 
@@
 per subject, which is a ballpark value commonly observed with the 
-popular \href{https://www.fil.ion.ucl.ac.uk/spm/}{SPM} or \href{https://fsl.fmrib.ox.ac.uk/fsl/fslwiki/}{FSL packages},
+popular \href{https://www.fil.ion.ucl.ac.uk/spm/}{SPM} or \href{https://fsl.fmrib.ox.ac.uk/fsl/fslwiki/}{FSL} packages,
 an input data size of 100MB per subject, 
 and an output data size of 2GB (20-fold increase compared to input 
-size), the D/C ratio would be 1.8MB/s, which would reach the 
+size), the data-write rate would be 1.8MB/s, which would reach the 
 $\delta/\gamma$ threshold measured on this cluster for $\gamma=108$, 
 that is, if 108 subjects were processed on the same node. This is very 
-unlikely as the number of CPUs per node is 40. We therefore conclude 
+unlikely as the number of CPUs per node was 40. We therefore conclude 
 that in-memory computing is likely to be useless for fMRI analysis.
@@
 ones of Spark on Lustre (non-hatched blue bars). The speed-up provided by 
-local execution peaks at 3.2, reached at 750 chunks in experiment 3. Overall, 
-writing locally was usually preferrable to writing to Lustre, as a result of the
-lower contention on local disk. Although it may be true that,
+local execution peaked at 3.2, for 750 chunks in experiment 3. Overall, 
+writing locally was usually preferable over writing to Lustre, as a result of the
+lower contention on local disk. Although it may be true that
 network bandwidths exceed that of disks~\cite{ananthanarayanan2011disk},
 locality remains important 
-as contention on a shared filesystem will tends to be much higher than on local disk.
+as contention on a shared filesystem tends to be much higher than on local disk.
 The only time writing locally did not
@@
 Fig.~\ref{fig:results}) to the ones of Spark on Lustre (non-hatched 
-blue bars). The speed-up provided by the combine use of data locality 
+blue bars). The speed-up provided by the combined use of data locality 
 and in-memory computing is also reported in 
@@
 
-As for the fMRI speculation, the D/C 
-ratio of 1.8MB/s would reach the $\Delta/\Gamma$ threshold for 
+As for the fMRI speculation, the data-write 
+rate of 1.8MB/s would reach the $\Delta/\Gamma$ threshold for 
 $\Gamma=280$, which is a realistic number of subjects to process on a 
@@
 
-The effects of lazy-evaluation can be seen throughout the experiments. Nipype 
+The effects of lazy evaluation can be seen throughout the experiments. Nipype 
 was found to be slower than Spark in most experiments. While the Nipype execution
@@
 generated prior to workflow execution, there are no optimizations
-to ensure that the least amount of work is performed to produce the 
+to ensure that the least amount of work is performed to produce the required 
 results.
@@
 that fill up the page cache, tmpfs/page cache performance on will be limited for other users. 
-However, with memory, it is possible to request through the HPC scheduler, a certain
+However, it is possible to request through the HPC scheduler a certain
 amount of available memory. Ultimately, in-memory data will also need to be spilled to
@@
 
-Unlike Spark, Nipype, by default, opts to use all available CPUs rather 
+Unlike Spark, Nipype by default opts to use all available CPUs rather 
 than to load balance data across the cluster. That is, given 50 chunks 
@@
 
-With Nipype, MapNodes, which apply a given function to each element in 
+Nipype's MapNodes, which apply a given function to each element in 
 a list, were found to be slower than the Node, which apply a function 
@@
 with it. While Spark will recompute the lost partition automatically using lineage, 
-Nipype will fail for all tasks requiring the data. Nevertheless, Spark will alse fail if 
+Nipype will fail for all tasks requiring the data. Nevertheless, Spark will also fail if 
 RDDs of node-local filenames are shuffled, as the data associated to the filenames
@@
 
-Nipype, when executed on Lustre, will checkpoint itself, ensuring resumption from
-last checkpoint. This is particularly important in the case of compute-intensive 
+When executed on Lustre, Nipype will checkpoint itself, ensuring resumption from
+last checkpoint during re-execution. This is particularly important in the case of compute-intensive 
 applications, such as those found in neuroimaging. Spark also provides a checkpointing
@@
 It would also not be an option with Spark in-memory. To enable this, burst buffers or 
-heteorogeneous storage managers (e.g. Triple-H~\cite{islam2015triple}) could be used to ensure
+heterogeneous storage managers (e.g. Triple-H~\cite{islam2015triple}) could be used to ensure
 fast processing and that all outputs (including intermediate outputs) will be 
@@
 
-Data-locality plays an important role in application 
+Data locality plays an important role in application 
 performance. Local disk was found to perform better than the shared 
-filesystem, despite having lower bandwidth, due to increased contention on 
+filesystem despite having lower bandwidth, due to increased contention on 
 the shared filesystem. Since local disk will typically have less 

### 2019-03-28 results/code/gen_all_figures.sh

@@ -3,7 +3,11 @@
 python results/code/get_makespan.py results/benchmarks results/code/incrementation_makespan.out
 
 python results/code/d_c_graph.py results/benchmarks results/code/incrementation_makespan.out results/figures/incrementation.pdf
-python results/code/gen_figures.py 1 results/code/incrementation_makespan.out results/figures/iterations.pdf
-python results/code/gen_figures.py 2 results/code/incrementation_makespan.out results/figures/numchunks.pdf
-python results/code/gen_figures.py 3 results/code/incrementation_makespan.out results/figures/datasize.pdf
-python results/code/gen_figures.py 4 results/code/incrementation_makespan.out results/figures/cputime.pdf
+python results/code/gen_figures.py 1 results/code/incrementation_makespan.out -f "spark" results/figures/s-iterations.pdf
+python results/code/gen_figures.py 2 results/code/incrementation_makespan.out -f "spark" results/figures/s-numchunks.pdf
+python results/code/gen_figures.py 3 results/code/incrementation_makespan.out -f "spark" results/figures/s-datasize.pdf
+python results/code/gen_figures.py 4 results/code/incrementation_makespan.out -f "spark" results/figures/s-cputime.pdf
+python results/code/gen_figures.py 1 results/code/incrementation_makespan.out -f "nipype" results/figures/n-iterations.pdf
+python results/code/gen_figures.py 2 results/code/incrementation_makespan.out -f "nipype" results/figures/n-numchunks.pdf
+python results/code/gen_figures.py 3 results/code/incrementation_makespan.out -f "nipype" results/figures/n-datasize.pdf
+python results/code/gen_figures.py 4 results/code/incrementation_makespan.out -f "nipype" results/figures/n-cputime.pdf

### 2019-03-28 results/code/gen_figures.py

@@ -8,7 +8,7 @@ exp14_delays = [2.4, 3.44, 7.68, 320]
 exp2_delays = [0.59, 3.52, 14.67]
 exp3_delays = 1.76
 
-def genfig(exp, makef, outf, cond):
+def genfig(exp, makef, outf, cond, framework="Spark"):
     bench_res = {('sp', 'mem'): [],
                  ('sp', 'tmpfs'): [],
                  ('sp', 'local'): [],
@@ -88,64 +88,70 @@ def genfig(exp, makef, outf, cond):
         alpha = 0.6
         hatch = '/'
 
-        spmem_data = (bench_res[('sp', 'mem')][0] - bar_width * 3,
-                      bench_res[('sp', 'mem')][1], bar_width)
-        col = '#F4CC70'
-        lbl = 'Spark - in-memory'
-        spim = ax.bar(*spmem_data, color=col, label=lbl)
-        spim_2 = ax2.bar(*spmem_data, color=col, label=lbl)
+        if framework == "spark":
+            spmem_data = (bench_res[('sp', 'mem')][0] - bar_width * 1,
+                          bench_res[('sp', 'mem')][1], bar_width)
+            col = '#F4CC70'
+            lbl = 'Spark - in-memory'
+            spim = ax.bar(*spmem_data, color=col, label=lbl)
+            spim_2 = ax2.bar(*spmem_data, color=col, label=lbl)
 
-        sptmpfs_data = (bench_res[('sp', 'tmpfs')][0] - bar_width * 2,
-                        bench_res[('sp', 'tmpfs')][1],
-                        bar_width)
-        col = '#DE7A22'
-        lbl = 'Spark - tmpfs'
-        
-        sptmpfs = ax.bar(*sptmpfs_data, color=col, label=lbl)
-        sptmpfs_2 = ax2.bar(*sptmpfs_data, color=col, label=lbl)
+            sptmpfs_data = (bench_res[('sp', 'tmpfs')][0],
+                            bench_res[('sp', 'tmpfs')][1],
+                            bar_width)
+            col = '#DE7A22'
+            lbl = 'Spark - tmpfs'
+            
+            sptmpfs = ax.bar(*sptmpfs_data, color=col, label=lbl)
+            sptmpfs_2 = ax2.bar(*sptmpfs_data, color=col, label=lbl)
 
-        nptmpfs_data = (bench_res[('np', 'tmpfs')][0] - bar_width,
[…]

### 2019-03-29 results/code/gen_all_figures.sh
final fixes

@@ -3,11 +3,7 @@
 python results/code/get_makespan.py results/benchmarks results/code/incrementation_makespan.out
 
 python results/code/d_c_graph.py results/benchmarks results/code/incrementation_makespan.out results/figures/incrementation.pdf
-python results/code/gen_figures.py 1 results/code/incrementation_makespan.out -f "spark" results/figures/s-iterations.pdf
-python results/code/gen_figures.py 2 results/code/incrementation_makespan.out -f "spark" results/figures/s-numchunks.pdf
-python results/code/gen_figures.py 3 results/code/incrementation_makespan.out -f "spark" results/figures/s-datasize.pdf
-python results/code/gen_figures.py 4 results/code/incrementation_makespan.out -f "spark" results/figures/s-cputime.pdf
-python results/code/gen_figures.py 1 results/code/incrementation_makespan.out -f "nipype" results/figures/n-iterations.pdf
-python results/code/gen_figures.py 2 results/code/incrementation_makespan.out -f "nipype" results/figures/n-numchunks.pdf
-python results/code/gen_figures.py 3 results/code/incrementation_makespan.out -f "nipype" results/figures/n-datasize.pdf
-python results/code/gen_figures.py 4 results/code/incrementation_makespan.out -f "nipype" results/figures/n-cputime.pdf
+python results/code/gen_figures.py 1 results/code/incrementation_makespan.out results/figures/iterations.pdf
+python results/code/gen_figures.py 2 results/code/incrementation_makespan.out results/figures/numchunks.pdf
+python results/code/gen_figures.py 3 results/code/incrementation_makespan.out results/figures/datasize.pdf
+python results/code/gen_figures.py 4 results/code/incrementation_makespan.out results/figures/cputime.pdf

### 2019-03-29 results/code/gen_figures.py

@@ -8,7 +8,7 @@ exp14_delays = [2.4, 3.44, 7.68, 320]
 exp2_delays = [0.59, 3.52, 14.67]
 exp3_delays = 1.76
 
-def genfig(exp, makef, outf, cond, framework="Spark"):
+def genfig(exp, makef, outf, cond):
     bench_res = {('sp', 'mem'): [],
                  ('sp', 'tmpfs'): [],
                  ('sp', 'local'): [],
@@ -88,70 +88,64 @@ def genfig(exp, makef, outf, cond, framework="Spark"):
         alpha = 0.6
         hatch = '/'
 
-        if framework == "spark":
-            spmem_data = (bench_res[('sp', 'mem')][0] - bar_width * 1,
-                          bench_res[('sp', 'mem')][1], bar_width)
-            col = '#F4CC70'
-            lbl = 'Spark - in-memory'
-            spim = ax.bar(*spmem_data, color=col, label=lbl)
-            spim_2 = ax2.bar(*spmem_data, color=col, label=lbl)
+        spmem_data = (bench_res[('sp', 'mem')][0] - bar_width * 3,
+                      bench_res[('sp', 'mem')][1], bar_width)
+        col = '#F4CC70'
+        lbl = 'Spark - in-memory'
+        spim = ax.bar(*spmem_data, color=col, label=lbl)
+        spim_2 = ax2.bar(*spmem_data, color=col, label=lbl)
 
-            sptmpfs_data = (bench_res[('sp', 'tmpfs')][0],
-                            bench_res[('sp', 'tmpfs')][1],
-                            bar_width)
-            col = '#DE7A22'
-            lbl = 'Spark - tmpfs'
-            
-            sptmpfs = ax.bar(*sptmpfs_data, color=col, label=lbl)
-            sptmpfs_2 = ax2.bar(*sptmpfs_data, color=col, label=lbl)
+        sptmpfs_data = (bench_res[('sp', 'tmpfs')][0] - bar_width * 2,
+                        bench_res[('sp', 'tmpfs')][1],
+                        bar_width)
+        col = '#DE7A22'
+        lbl = 'Spark - tmpfs'
+        
+        sptmpfs = ax.bar(*sptmpfs_data, color=col, label=lbl)
+        sptmpfs_2 = ax2.bar(*sptmpfs_data, color=col, label=lbl)
 
-            splocal_data = (bench_res[('sp', 'local')][0] + bar_width,
-                            bench_res[('sp', 'local')][1],
[…]

### 2019-04-02 paper.tex

@@
 \newcommand{\Desc}[2]{\State \makebox[2em][l]{#1}#2}
-
 \begin{document}
@@
 \title{Performance Evaluation of Big Data Processing Strategies for Neuroimaging}
+\pagenumbering{gobble}
 
@@
     advancements in image acquisition methods, open-science and data sharing. 
-    However, current neuroimaging data engines have not adopted any strategies
-    to address the rise
-    in neuroimaging data. Here, we 
+    However, the adoption of Big Data processing strategies by neuroimaging
+    processing engines remains limited. Here, we 
     evaluate three Big Data processing strategies (in-memory computing, 
@@
     Big Data thresholds were modeled by comparing the data-write rate of the 
-    application to the filesystem-application concurrent write rate. 
+    application to the filesystem bandwidth and number of concurrent processes. 
     This model acknowledges the 
@@
 speed-up values are likely to be observed on typical image processing 
-operations performed on images of size larger than 75GB. 
+operations performed on images of size larger than 75GB. A ballpark 
+speculation from our model showed that in-memory computing alone will
+not speed-up current 
+functional MRI analyses unless coupled with data 
+locality and processing around 280 subjects concurrently. Furthermore, we observe
+    that emulating in-memory computing 
+using in-memory file systems (tmpfs) does not reach the performance of an in-memory engine,
+    presumably due to swapping to disk and the lack of data cleanup.
 We conclude that Big Data processing strategies are 
@@
 
-In Big Data applications, there is a significant cost associated with data movement.
-Big Data processing engines have addressed these issues by diminishing the amount 
-of data movement that occurs during the execution of an application. The 
-MapReduce~\cite{dean2008mapreduce} framework popularized the concept of locality-aware
-scheduling, which ensured tasks were scheduled nearest to where the data the data was
-located (data locality), thereby reducing the number of network transfers in an application.
-However, it was found that performance in MapReduce could be improved upon by also
-reducing the disk I/O between tasks by keeping data in-memory whenever possible,
-and thus, in-memory computing was added to the Apache Spark~\cite{zaharia2016apache} framework.
-Frameworks like Apache Spark also included the lazy evaluation strategy, which performs computations
-only when invoked. This allows for additional optimizations such as the regrouping of tasks
-and only invoking what is absolutely necessary to obtain the desired result.
+Big Data processing engines have significantly improved the performance of Big Data
+applications by diminishing the amount 
+of data movement that occurs during the execution of an application. Locality-aware
+scheduling, introduced by the MapReduce~\cite{dean2008mapreduce} framework, reduced the overall
+costs associated to network transfer of data by scheduling tasks to the nodes located nearest to
+the data. Reduction of data movement was further improved upon through in-memory computing~\cite{zaharia2016apache},
+which ensures that data is maintained in memory between tasks whenever possible. To further
+reduce the cost of data movement, lazy evaluation, the process of performing computations only when invoked,
+was leveraged by Big Data frameworks to enable further optimizations such as regrouping of tasks and
+computing only what is necessary.
 
 Frameworks such as
-MapReduce and Spark have become mainstream tools for data analytics, however, many
+MapReduce and Spark~\cite{zaharia2016apache} have become mainstream tools for data analytics, although many
 others, such as Dask~\cite{rocklin2015dask}, are emerging.
@@
 
-Infrastructure-wise, we focus on the case of High-Performance Computing 
+In terms of infrastructure, we focus on the case of High-Performance Computing 
 (HPC) clusters that are typically available through University 
@@
 Although Spark uses in-memory computing, it is not necessary for all the data to
-fit in memory. Spark will spill any RDD elements that do not fit in memory to disk.
+fit in memory. Spark will spill any RDD elements that cannot be maintained in memory to disk.
 Moreover, as Spark transformations generate new RDDs and numerous 
@@
 \end{tabular}
+\setlength{\belowcaptionskip}{-10pt}
 \caption{Big Data strategies on a shared HPC cluster.}
@@
  file system is used, the user is responsible ensuring that the filesystem does not
- reach capacity. Should it be necessary to free up additional memory, the kernel will swap filesystem memory to disk,
+ reach capacity. Should it be necessary to free up additional memory, the kernel will swap filesystem memory to disk
  and the performance will become that of local disk writes. In our experiments, 
@@
 \end{equation}
-This defines a data-write rate beyond which the page cache becomes 
+This defines a D/C (data-write) rate beyond which the page cache becomes 
 asymptotically useless. It should be noted that the transient phase 
@@
  of image blocks containing random intensities, to avoid 
- caching effects (measure\_bandwidth.py script).
+ caching effects (see: \href{https://github.com/big-data-lab-team/paper-in-mem-locality/blob/master/benchmark_scripts/measure_bandwidth.py}{measure\_bandwidth.py}).
  They are reported in Table~\ref{table:bdwdths}.
@@
 \end{tabular}
+\setlength{\belowcaptionskip}{-10pt}
 \caption{Measured bandwidths}
@@
 Processing large images is only considered to be part of the Big Data problem 
-in neuroscience. The other being the processing large MRI datasets, 
+in neuroscience. The other problem being the processing large MRI datasets, 
 that is, datasets consisting of many small brain images belonging to 
@@
 \end{tabular}
+\setlength{\belowcaptionskip}{-10pt}
 \caption{Experiment conditions. Red cells denote the conditions where 
@@
 use up as many cores as possible. Unlike the previous 
-experiment, the data-write rate was kept static at 178.6MB/s, however, this 
+experiment, the D/C rate was kept static at 178.6MB/s, however, this 
 ratio ensured that different regions of the inequality were reached 
@@
 per task), however, due to varying size in total data processed (D), 
-the data-write rate varied. Once again, we ensured that the data-write rate fell in 
-multiple different regions of the inequality. The data-write rate ranged from 
+the D/C rate varied. Once again, we ensured that the D/C rate fell in 
+multiple different regions of the inequality. The D/C rate ranged from 
 349.1 MB/s for BigBrain and 174.6MB/s for half BigBrain, to 0.06MB/s for the MRI 
@@
     \centering
-    \begin{subfigure}{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{results/figures/s-iterations.pdf}%
-    \end{subfigure}
-    \begin{subfigure}{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{results/figures/s-iterations.pdf}%
-    \end{subfigure}
-    \includegraphics[width=\linewidth]{results/figures/iterations.pdf}
+    \includegraphics[width=\columnwidth]{results/figures/iterations.pdf}%
     \caption{Experiment 1: complete BigBrain, 125 chunks, 3.44-second tasks.}\label{fig:iterations}
@@
 \end{subfigure}
+\setlength{\belowcaptionskip}{-10pt}
 \caption{Experiment results: Makespans of Spark and Nipype writing to memory, tmpfs, local disk and Lustre.}
@@
 \end{subfigure}
+\setlength{\belowcaptionskip}{-10pt}
     \caption{Page cache model evaluation. Grey 
@@
 the makespan obtained with Spark for in-memory computing. Experiments 
-for which there was no in-memory equivalent (i.e. BigBrain split into 
+for which there were no in-memory equivalent (i.e. BigBrain split into 
 30 
@@
 page cache on processing times for local disk and Lustre. That is, the 
-speed-up provided by in-memory computing was larger than 1 for data-write 
+speed-up provided by in-memory computing was larger than 1 for D/C 
 rates larger than $\delta/\gamma$ (local disk) or $\Delta/\Gamma$ 
 (Lustre). Conversely, the speed-up provided by in-memory computing 
-remained close to 1 for data-write rates smaller than $\delta/\gamma$ (local 
+remained close to 1 for D/C rates smaller than $\delta/\gamma$ (local 
 disk) or $\Delta/\Gamma$ (Lustre). The two points close to the origin 
@@
 and an output data size of 2GB (20-fold increase compared to input 
-size), the data-write rate would be 1.8MB/s, which would reach the 
+size), the D/C rate would be 1.8MB/s, which would reach the 
 $\delta/\gamma$ threshold measured on this cluster for $\gamma=108$, 
@@
 
-As for the fMRI speculation, the data-write 
+As for the fMRI speculation, the D/C 
 rate of 1.8MB/s would reach the $\Delta/\Gamma$ threshold for 
@@
 executors to 5, to preserve a good I/O 
-throughput\footnote{\url{http://blog.cloudera.com/blog/2015/03/how-to-tune-your-apache-spark-jobs-part-2}}, 
+throughput (see \href{http://blog.cloudera.com/blog/2015/03/how-to-tune-your-apache-spark-jobs-part-2}{Cloudera blog}), 
 but the rationale for this recommendation is hardly explained. We 
@@
 It was expected that Spark would experience longer processing times, particularly
-with the small datasets, due to Java serializing. This was not found to be the case. 
+with the small datasets, due to Java serialization. This was not found to be the case. 
 Unlike Spark, Nipype performs a more thorough provenance capture, potentially owing to longer
 processing times.
+
+In this paper, we analyzed the effects of Big Data strategies on an map-only artificial
+neuroimaging pipeline. This allowed us to examine the effects of these strategies
+without being significantly obscured by other conditions. Studying the effects 
+of such strategies on map-reduce type
+workflows, in addition to real neuroimaging pipelines, would allows us to gain
+further insight on the added
+value of the Big Data performance stragies for neuroimaging use cases remains to be
+done.
 
@@
 neuroimaging applications, including developing their provenance, 
-fault-tolerance, and reproducibility features, is a worthwhile project.
+fault-tolerance, and reproducibility features, is worthwhile.
 
@@
 filesystem despite having lower bandwidth, due to increased contention on 
-the shared filesystem. Since local disk will typically have less 
-contention than a shared filesystem, it is recommended to store data locally.
+the shared filesystem. Since local disk typically has less 
+contention than shared filesystems, it is recommended to store data locally.
 However, using local storage without a distributed file system 
-may come at a cost to fault tolerance.
+may limit fault tolerance.
 
@@
 overlay scheduling. The impact of other Big Data technologies, such as 
-distributed in-memory file systems~\cite{ignite}, could also be 
-investigated.
+distributed in-memory file systems(e.g.~\href{https://ignite.apache.org}{Apache Ignite}) and Lustre scalability issues~\cite{sparkhpc},
+could also be investigated.
 
@@
 in Roundrock, TX, for providing the infrastructure and high-quality
-technical support that made possible the experiments reported in this paper.
+technical support.
 
